---
title: "Effects of sediment on the recruit survival of juvenile corals"
author: "Lillian J. Tuttle"
date: "7/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(mixmeta)
library(dosresmeta)
library(splines)

#will want to getwd() and setwd(), as appropriate
effectsize <- read.csv("/Volumes/GoogleDrive/My Drive/iScience/Code/thresholds/sedimentation/effectsize2.csv", header=T)

#keyboard shortcuts to remember: 
#command-option-I inserts new code chunk
#command-shift-M inserts pipeline symbols
#command-enter runs line of code where cursor is (or highlighted portion)
```
  
```{r, echo=FALSE, warning=FALSE}
#Looking at data
#head(effectsize) 

#a lot of variables (columns) are factors instead of numeric/doubles/integers
effectsize <- effectsize %>% #converting appropriate columns to numeric...
  mutate_at(vars(ES, Hedges_d, Var_d, SE_d, CI_lo_d, CI_up_d, Sed_level_standardized, Sed_exposure, Sed_exposure_mg_d_L), as.character) %>% 
  mutate_at(vars(ES, Hedges_d, Var_d, SE_d, CI_lo_d, CI_up_d, Sed_level_standardized, Sed_exposure, Sed_exposure_mg_d_L), as.numeric)
#head(effectsize) #"double"-check!

#Now create a new variable that has both Genus and species, remove sediment burial experiments, and remove irrelevant rows (non-controls, -treatments, or -baselines)
effectsize <- effectsize %>%
  mutate(Gsp = paste(Updated_Genus, Updated_species, sep = "_")) %>% 
  filter(Binary_sed_burial=="0") %>%
  filter(!Control=="-1")
#head(effectsize)

#Separate into deposited sediment (DS) and suspended sediment (SS) datasets, WITH CONTROLS
ES_DS <- effectsize %>% filter(DS_SS=="DS")
ES_SS <- effectsize %>% filter(DS_SS=="SS")

#Second dataset that that removes controls (Control == c("1","2"))
effectsize2 <- effectsize %>% filter(Control=="0")
#effectsize2 #"double"-check!

#Separate into deposited sediment (DS) and suspended sediment (SS) datasets, WITHOUT CONTROLS
ES_DS2 <- effectsize2 %>% filter(DS_SS=="DS")
ES_SS2 <- effectsize2 %>% filter(DS_SS=="SS")

#Now separate into independent dataset for each coral response for DEPOSITED SEDIMENT
ES_DS_recruit <- ES_DS %>% 
  filter(Response == "recruitment") %>% 
  filter(Surface_direction == "top")
#ES_DS_recruit %>% tally()
#ES_DS_recruit %>% count(Ref)
#ES_DS_recruit %>% count(Ref_name)
#ES_DS_recruit %>% count(Gsp)
#ES_DS_recruit %>% count(Updated_Genus)
``` 
  
These datasets contain information from multiple articles (**Ref_name**) and studies/experiments (**Ref**) within those articles about the effects of DEPOSITED sediment on the **recruit survival rate** of juvenile corals. I would like to explore the dose-response relationship between sediment 'dose' and the effect size, estimated here as the standardized mean difference between treatment and control, in terms of Hedges' *d*:

- Hedges' *d* is the standardized mean difference between the treatment and control groups (Hedges and Olkin 1985). Each study may have multiple Hedges' *d* calculations, one for each treatment-control comparison at each time-point.   

- 'Dose' can be quantified as sediment exposure concentration (mg/cm^2^/day), exposure duration (days), or as "cumulative exposure", for which I multiply concentration and duration (mg/cm^2^).

Each study/experiment (Ref) may contain multiple sets of controls and treatments (due to using multiple species or sediment types, for instance). Thus, I have added two categorical vectors, '**Comparison**' and 'Control', which group each set of control-treatment comparisons into a unique two-letter code, within which is one control (Control=="1") and at least one treatment (Control=="0"). **Therefore, 'Comparison' is nested within 'Ref', which is further nested within 'Ref_name'.**

```{r, echo=FALSE}
#Using dosresmeta to calculate Hedges' d and variance of d
covar_DS_recruit <- by(ES_DS_recruit, ES_DS_recruit$Comparison, function(x) 
  covar.smd(Tx_mean, Tx_sd, Tx_n, "smd", method="hedges", data = x))
ES_DS_recruit$smd <- unlist(lapply(covar_DS_recruit, function(x) x$y))
ES_DS_recruit$vmd <- unlist(lapply(covar_DS_recruit, function(x) x$v))

#Looking at data to make sure smd and vmd added as columns
#ES_DS_recruit

#Removing controls for modeling
ES_DS_recruit2 <- subset(ES_DS_recruit, Control=="0")

#Looking at data to make sure controls were removed
#ES_DS_recruit2
ES_DS_recruit2 %>% tally()
``` 
  
# I. Recruit Survival Rate  

## A. Deposited Sediment  
  
### 1. Plots  
I calculated the standardized mean difference in terms of Hedges' *d* and plot it here with respect to cumulative exposure. I did this first for a Hedges' *d* that I calculated before importing the data to R and again for a Hedges' *d* that is calculated by the dosresmeta::covar.smd function.

```{r, echo=FALSE, warning=FALSE}
#CUMULATIVE EXPOSURE vs. effect size, by study
#Using previously calculated Hedges' d and variance of d
ggplot(ES_DS_recruit2, 
       aes(x = Sed_exposure_mg_cm2,
           y = Hedges_d,
           color = Ref,
           ymin = Hedges_d-Var_d,
           ymax = Hedges_d+Var_d)) + 
  geom_pointrange() +
  ggtitle("Recruit survival rate") +
  labs(x = "Sediment cumulative exposure (mg/cm2)",
       y = "Effect size (Hedges' d +/- s.d., manual calculation)",
       color = "Study") +
  geom_abline(intercept=0, slope=0) +
  theme_classic() +
  scale_x_continuous(limits = c(0,1500))
#Using dosresmeta's calculation of Hedges' d and variance of d
ggplot(ES_DS_recruit2, 
       aes(x = Sed_exposure_mg_cm2,
           y = smd,
           color = Ref,
           ymin = smd-vmd,
           ymax = smd+vmd)) + 
  geom_pointrange() +
  ggtitle("Recruit survival rate") +
  labs(x = "Sediment cumulative exposure (mg/cm2)",
       y = "Effect size (Hedges' d +/- s.d., dosresmeta calculation)",
       color = "Study") +
  geom_abline(intercept=0, slope=0) +
  theme_classic() +
  scale_x_continuous(limits = c(0,1500))

#EXPOSURE CONCENTRATION vs. effect size, by study
#Using previously calculated Hedges' d and variance of d
ggplot(ES_DS_recruit2, 
       aes(x = Sed_level_standardized,
           y = Hedges_d,
           color = Ref,
           ymin = Hedges_d-Var_d,
           ymax = Hedges_d+Var_d)) + 
  geom_pointrange() +
  ggtitle("Recruit survival rate") +
  labs(x = "Sediment exposure concentration (mg/cm2/day)",
       y = "Effect size (Hedges' d +/- s.d., manual calculation)",
       color = "Study") +
  geom_abline(intercept=0, slope=0) +
  theme_classic() 
#Using dosresmeta's calculation of Hedges' d and variance of d
ggplot(ES_DS_recruit2, 
       aes(x = Sed_level_standardized,
           y = smd,
           color = Ref,
           ymin = smd-vmd,
           ymax = smd+vmd)) + 
  geom_pointrange() +
  ggtitle("Recruit survival rate") +
  labs(x = "Sediment exposure concentration (mg/cm2/day)",
       y = "Effect size (Hedges' d +/- s.d., dosresmeta calculation)",
       color = "Study") +
  geom_abline(intercept=0, slope=0) +
  theme_classic() 


#EXPOSURE CONCENTRATION vs. effect size, by SEDIMENT TYPE
#Using previously calculated Hedges' d and variance of d
ggplot(ES_DS_recruit2, 
       aes(x = Sed_level_standardized,
           y = smd,
           color = Sed_grain_category,
           shape = Sed_source,
           ymin = smd-vmd,
           ymax = smd+vmd)) + 
  geom_pointrange() +
  ggtitle("Recruit survival rate") +
  labs(x = "Sediment exposure concentration (mg/cm2/day)",
       y = "Effect size (Hedges' d +/- s.d., dosresmeta calculation)",
       color = "Sediment relative\ngrain size",
       shape = "Sediment source") +
  geom_abline(intercept=0, slope=0) +
  theme_classic()
ggplot(ES_DS_recruit2, 
       aes(x = Sed_level_standardized,
           y = smd,
           color = Sed_grain_class,
           shape = Sed_mineralogy,
           ymin = smd-vmd,
           ymax = smd+vmd)) + 
  geom_pointrange() +
  ggtitle("Recruit survival rate") +
  labs(x = "Sediment exposure concentration (mg/cm2/day)",
       y = "Effect size (Hedges' d +/- s.d., dosresmeta calculation)",
       color = "Sediment\ngrain class",
       shape = "Sediment mineralogy") +
  geom_abline(intercept=0, slope=0) +
  theme_classic()

#EXPOSURE CONCENTRATION vs. effect size, by PHYLOGENY
#Using previously calculated Hedges' d and variance of d
ggplot(ES_DS_recruit2, 
       aes(x = Sed_level_standardized,
           y = smd,
           color = Gsp,
           ymin = smd-vmd,
           ymax = smd+vmd)) + 
  geom_pointrange() +
  ggtitle("Recruit survival rate") +
  labs(x = "Sediment exposure concentration (mg/cm2/day)",
       y = "Effect size (Hedges' d +/- s.d., dosresmeta calculation)",
       color = "Species") +
  geom_abline(intercept=0, slope=0) +
  theme_classic()

#EXPOSURE CONCENTRATION vs. effect size, by OCEAN and REGION
#Using previously calculated Hedges' d and variance of d
ggplot(ES_DS_recruit2, 
       aes(x = Sed_level_standardized,
           y = smd,
           color = Region,
           shape = Ocean,
           ymin = smd-vmd,
           ymax = smd+vmd)) + 
  geom_pointrange() +
  ggtitle("Recruit survival rate") +
  labs(x = "Sediment exposure concentration (mg/cm2/day)",
       y = "Effect size (Hedges' d +/- s.d., dosresmeta calculation)",
       color = "Region",
       shape = "Ocean") +
  geom_abline(intercept=0, slope=0) +
  theme_classic()
```
  
It is interesting that there is a difference between how I calculated Hedges' *d* and how *dosresmeta* appears to calculate it. But in both cases, there does not appear to be an effect of sediment exposure on effect size, except for *Acropora hyacinthus*, for which there appears to be a strong effect of sediment on effect size. There is not a discernible effect of geography (ocean and region) because almost all of the data were from the Marianas Islands. I still need to investigate the effects of sediment type (source, origin, grain size).
  
### 2. Model Fitting 
The following set of functions model fixed and random (nested) effects WITHOUT accounting for within-comparison (or within-study) correlations. The output for each is a univariate, random-effects meta-regression.

```{r, echo=FALSE, warning=FALSE}
#?mixmeta

# Organizational Note: 
#All models with 'a' or 'b' at the end of its name have cumulative exposure as its predictor. 
#All models with 'c' or 'd' at the end of its name have exposure concentration as its predictor. 
#All models with 'a' or 'c' at the end of its name have manually calculated Hedges' d as its response.
#All models with 'b' or 'd' at the end of its name have covar.smd calculated Hedges' d as its response.

# LINEAR FIXED AND RANDOM EFFECTS NOT ACCOUNTING FOR WITHIN-STUDY CORRELATIONS
mod_DS_recruit_1a <- mixmeta(Hedges_d ~ Sed_exposure_mg_cm2, S = Var_d,
                       random =  ~ Sed_exposure_mg_cm2 | Comparison,
                       data = ES_DS_recruit2, method = "ml")
mod_DS_recruit_1b <- mixmeta(smd ~ Sed_exposure_mg_cm2, S = vmd,
                      random =  ~ Sed_exposure_mg_cm2 | Comparison,
                      data = ES_DS_recruit2, method = "ml")
mod_DS_recruit_1c <- mixmeta(Hedges_d ~ Sed_level_standardized, S = Var_d,
                       random =  ~ Sed_level_standardized | Comparison,
                       data = ES_DS_recruit2, method = "ml")
mod_DS_recruit_1d <- mixmeta(smd ~ Sed_level_standardized, S = vmd,
                      random =  ~ Sed_level_standardized | Comparison,
                      data = ES_DS_recruit2, method = "ml")

# LINEAR FIXED AND RANDOM, NESTED EFFECTS NOT ACCOUNTING FOR WITHIN-STUDY CORRELATIONS
mod_DS_recruit_2a <- mixmeta(Hedges_d ~ Sed_exposure_mg_cm2, S = Var_d,
                       random =  ~ Sed_exposure_mg_cm2 | Ref/Comparison,
                       data = ES_DS_recruit2, method = "ml")
mod_DS_recruit_2b <- mixmeta(smd ~ Sed_exposure_mg_cm2, S = vmd,
                      random =  ~ Sed_exposure_mg_cm2 | Ref/Comparison,
                      data = ES_DS_recruit2, method = "ml")
mod_DS_recruit_2c <- mixmeta(Hedges_d ~ Sed_level_standardized, S = Var_d,
                       random =  ~ Sed_level_standardized | Ref/Comparison,
                       data = ES_DS_recruit2, method = "ml")
mod_DS_recruit_2d <- mixmeta(smd ~ Sed_level_standardized, S = vmd,
                      random =  ~ Sed_level_standardized | Ref/Comparison,
                      data = ES_DS_recruit2, method = "ml")
```

To account for within-comparison correlations in the mixmeta model, I replace the 'S' argument with a 'control' argument, which requires me to provide the covariance matrices, as done with the covar.smd function. 

```{r, echo=FALSE, warning=FALSE}
# COMPUTE THE WITHIN-COMPARISON CORRELATIONS 
# already done above but code repeated here...
#covar_DS_recruit <- by(ES_DS_recruit, ES_DS_recruit$Comparison, function(x) 
#  covar.smd(Tx_mean, Tx_sd, Tx_n, "smd", method="hedges", data = x))

# LINEAR FIXED AND RANDOM EFFECTS ACCOUNTING FOR WITHIN-COMPARISON CORRELATIONS
#Code for covariance matrix to include for addSlist for Comparison random effect
newlist_REC <- list(NA)
for (i in seq(1,length(covar_DS_recruit))) {
newlist_REC[i] <- list(covar_DS_recruit[[i]]$S)
}
#Models with the Hedges' d I calculated (5a) and the one calculated by the package (5b)
mod_DS_recruit_5a <- mixmeta(Hedges_d ~ Sed_exposure_mg_cm2,
                       random = ~ Sed_exposure_mg_cm2 | Comparison,
                       data=ES_DS_recruit2, method="ml", #using dataset without controls
                       control=list(addSlist=newlist_REC))
mod_DS_recruit_5b <- mixmeta(smd ~ Sed_exposure_mg_cm2,
                       random = ~ Sed_exposure_mg_cm2 | Comparison,
                       data=ES_DS_recruit2, method="ml", #using dataset without controls
                       control=list(addSlist=newlist_REC))
mod_DS_recruit_5c <- mixmeta(Hedges_d ~ Sed_level_standardized,
                       random = ~ Sed_level_standardized | Comparison,
                       data=ES_DS_recruit2, method="ml", #using dataset without controls
                       control=list(addSlist=newlist_REC))
mod_DS_recruit_5d <- mixmeta(smd ~ Sed_level_standardized,
                       random = ~ Sed_level_standardized | Comparison,
                       data=ES_DS_recruit2, method="ml", #using dataset without controls
                       control=list(addSlist=newlist_REC))


# LINEAR FIXED AND RANDOM, NESTED EFFECTS ACCOUNTING FOR WITHIN-COMPARISON & WITHIN-STUDY CORRELATIONS
#Code for covariance matrix to include for addSlist for Ref/Comparison
newlist_REC2 <- list(NA) #collects the list of covariance matrices for the block diag matrix by reference for the nested hierarchial ref/comparison
templist_REC <- list(NA) #holds temp list of covariance matrices associated with each reference
templist_REC2 <-list(NA)
reflista <- ES_DS_recruit %>% distinct(Ref,Ref_name,Comparison)
reflist <- reflista[,1]
for (i in seq(1,length(unique(reflista$Ref))))  {
  #pull the elements from covar_DS_recruit that are all from the same reference [i]
  templist_REC[i] <-list(covar_DS_recruit[reflist==unique(reflista$Ref)[i]])
  for (j in seq(1,length(templist_REC[[i]]))) {
  #for each comparison in the reference, pull out the covar matrices (element $S) and put in into templist_REC2
     templist_REC2[j] <- list(templist_REC[[i]][[j]]$S)
  }
  #turn list of covars from all comparison in one reference into block diag matrix
  newlist_REC2[i] <- list(bdiagMat(templist_REC2))
  templist_REC2 <- list(NA)
}
#Models with the Hedges' d I calculated (6a) and the one calculated by the package (6b)
mod_DS_recruit_6a <- mixmeta(Hedges_d ~ Sed_exposure_mg_cm2,
                       random = ~ Sed_exposure_mg_cm2 | Ref/Comparison,
                       data=ES_DS_recruit2, method="ml", #using dataset without controls
                       control=list(addSlist=newlist_REC2))
mod_DS_recruit_6b <- mixmeta(smd ~ Sed_exposure_mg_cm2,
                       random = ~ Sed_exposure_mg_cm2 | Ref/Comparison,
                       data=ES_DS_recruit2, method="ml", #using dataset without controls
                       control=list(addSlist=newlist_REC2))
mod_DS_recruit_6c <- mixmeta(Hedges_d ~ Sed_level_standardized,
                       random = ~ Sed_level_standardized | Ref/Comparison,
                       data=ES_DS_recruit2, method="ml", #using dataset without controls
                       control=list(addSlist=newlist_REC2))
mod_DS_recruit_6d <- mixmeta(smd ~ Sed_level_standardized,
                       random = ~ Sed_level_standardized | Ref/Comparison,
                       data=ES_DS_recruit2, method="ml", #using dataset without controls
                       control=list(addSlist=newlist_REC2))
```
  
And now to compare relative fit of linear models...

```{r, echo=FALSE, warning=FALSE}
# Model Comparisons
AIC(mod_DS_recruit_1a, mod_DS_recruit_2a, mod_DS_recruit_5a, mod_DS_recruit_6a) 
#of these, model 5a has lowest AIC

AIC(mod_DS_recruit_1b, mod_DS_recruit_2b, mod_DS_recruit_5b, mod_DS_recruit_6b) 
#of these, model 5b has lowest AIC

AIC(mod_DS_recruit_1c, mod_DS_recruit_2c, mod_DS_recruit_5c, mod_DS_recruit_6c) 
#of these, model 5c has lowest AIC

AIC(mod_DS_recruit_1d, mod_DS_recruit_2d, mod_DS_recruit_5d, mod_DS_recruit_6d) 
#of these, model 5d has lowest AIC
```
  
And now to evaluate residuals of models...

```{r, echo=FALSE, warning=FALSE}
# Residuals vs. Fitted Plots
op <- par(mfrow = c(2,2), mar = c(2,2,4,1))

resid_a <- resid(mod_DS_recruit_5a)
fitted_a <- fitted(mod_DS_recruit_5a)
plot(fitted_a, resid_a, main = "A: Hedges_d ~ mg_cm2, control = addSlist,
     random =  ~ mg_cm2 | Comparison")
abline(0,0)
#mod_DS_recruit_1a and mod_DS_recruit_3a were within 2 AIC

resid_b <- resid(mod_DS_recruit_5b)
fitted_b <- fitted(mod_DS_recruit_5b)
plot(fitted_b, resid_b, main = "B: smd ~ mg_cm2, control = addSlist,
     random =  ~ mg_cm2 | Comparison")
abline(0,0)
#mod_DS_recruit_6b within 1 AIC

resid_c <- resid(mod_DS_recruit_5c)
fitted_c <- fitted(mod_DS_recruit_5c)
plot(fitted_c, resid_c, main = "C: Hedges_d ~ mg_cm2_d, control = addSlist,
     random =  ~ mg_cm2_d | Comparison")
abline(0,0)
#mod_DS_recruit_2c within 0.3 AIC

resid_d <- resid(mod_DS_recruit_5d)
fitted_d <- fitted(mod_DS_recruit_5d)
plot(fitted_d, resid_d, main = "D: smd ~ mg_cm2_d, control = addSlist,
     random =  ~ mg_cm2_d | Comparison")
abline(0,0)

par(op)


# Normal Q-Q Plot
op <- par(mfrow = c(2,2), mar = c(2,2,4,1))
qqnorm(resid_a)
qqline(resid_a)
qqnorm(resid_b)
qqline(resid_b)
qqnorm(resid_c)
qqline(resid_c)
qqnorm(resid_d)
qqline(resid_d)
par(op)


# Patterns:
# 1.  Residuals for models A and B (mg/cm2 predictor) are very similar, with reasonable magnitudes (<abs(3)).
# 2.  Residuals for models C and D (mg/cm2/day predictor) are very similar, with reasonable magnitudes (<abs(3)), but a lot of "clumping" along fitted values axis bc they're all from same exposure concentration in single study.
# 3.  No obvious need to try out log or non-linear models!
```
  
### 3. Best-Fit Model Summary
  
#### i. MG/CM2 IS PREDICTOR, Model using manual calculation of Hedges' *d*
Now I will predict the effect size along the exposure range and plot with confidence intervals.  

```{r, echo=FALSE, warning=FALSE}
summary(mod_DS_recruit_5a)

# PREDICT THE EFFECT SIZE FOR 13.2 mg/cm2 FROM TWO MODELS (LOAEL identified by binary analyses)
#predict(mod_DS_recruit_5a, newdata=data.frame(Sed_exposure_mg_cm2=13.2), ci=TRUE)

# PREDICT THE EFFECT SIZE AND PLOT WITH CONFIDENCE INTERVALS
pred_mod_DS_recruit_5a <- predict(mod_DS_recruit_5a, newdata=ES_DS_recruit2, ci=TRUE)
ES_DS_recruit2_CI <- cbind(ES_DS_recruit2, pred_mod_DS_recruit_5a)
#head(ES_DS_recruit2_CI)
min_exp_a <- ES_DS_recruit2_CI %>% 
  mutate(overlap0 = 0 >= ci.lb & 0 <= ci.ub) %>% 
  filter(overlap0==FALSE) %>% 
  summarize(min_exp_a=min(Sed_exposure_mg_cm2))
min_exp_a2 <- as.numeric(min_exp_a)

ggplot(ES_DS_recruit2_CI, 
       aes(x = Sed_exposure_mg_cm2,
           y = Hedges_d,
           color = Ref,
           ymin = Hedges_d-Var_d,
           ymax = Hedges_d+Var_d)) + 
  geom_pointrange() +
  labs(x = expression("Sediment exposure concentration (mg/cm"^"2"*")"),
       y = expression("Effect size (Hedges'"~italic(d)~"+/- variance of"~italic(d)~")"),
       color = "Study") +
  geom_abline(intercept=0, slope=0) +
  geom_vline(xintercept=min_exp_a2, linetype="dashed", color = "red") +
  geom_ribbon(aes(x = Sed_exposure_mg_cm2, y = Hedges_d,
                  ymin = ci.lb, ymax = ci.ub), 
              fill = "grey70", alpha = .15, 
              show.legend=FALSE, inherit.aes=FALSE) +
  geom_line(aes(x = Sed_exposure_mg_cm2, y = fit), inherit.aes=FALSE) +
  theme_classic() +
  scale_x_continuous(limits = c(0,1500), breaks=c(0,500,1000,1500,min_exp_a2), 
                label=c("0","500","1000","1500",round(min_exp_a2,digits=1)))
```
  
#### ii. MG/CM2 IS PREDICTOR, Model using dosresmeta calculation of Hedges' *d*
Now I will predict the effect size along the exposure range and plot with confidence intervals.  

```{r, echo=FALSE, warning=FALSE}
summary(mod_DS_recruit_5b)

# PREDICT THE EFFECT SIZE FOR 13.2 mg/cm2 FROM TWO MODELS (LOAEL identified by binary analyses)
#predict(mod_DS_recruit_5b, newdata=data.frame(Sed_exposure_mg_cm2=13.2), ci=TRUE)

# PREDICT THE EFFECT SIZE AND PLOT WITH CONFIDENCE INTERVALS
pred_mod_DS_recruit_5b <- predict(mod_DS_recruit_5b, newdata=ES_DS_recruit2, ci=TRUE)
ES_DS_recruit2_CI2 <- cbind(ES_DS_recruit2, pred_mod_DS_recruit_5b)
#head(ES_DS_recruit2_CI2)
min_exp_b <- ES_DS_recruit2_CI2 %>% 
  mutate(overlap0 = 0 >= ci.lb & 0 <= ci.ub) %>% 
  filter(overlap0==FALSE) %>% 
  summarize(min_exp_b=min(Sed_exposure_mg_cm2))
min_exp_b2 <- as.numeric(min_exp_b)

ggplot(ES_DS_recruit2_CI2, 
       aes(x = Sed_exposure_mg_cm2,
           y = smd,
           color = Ref,
           ymin = smd-vmd,
           ymax = smd+vmd)) + 
  geom_pointrange() +
  labs(x = expression("Sediment exposure concentration (mg/cm"^"2"*")"),
       y = expression("Effect size (Hedges'"~italic(d)~"+/- variance of"~italic(d)~")"),
       color = "Study") +
  geom_abline(intercept=0, slope=0) +
  geom_vline(xintercept=min_exp_b2, linetype="dashed", color = "red") +
  geom_ribbon(aes(x = Sed_exposure_mg_cm2, y = smd,
                  ymin = ci.lb, ymax = ci.ub), 
              fill = "grey70", alpha = .15, 
              show.legend=FALSE, inherit.aes=FALSE) +
  geom_line(aes(x = Sed_exposure_mg_cm2, y = fit), inherit.aes=FALSE) +
  theme_classic() +
  scale_x_continuous(limits = c(0,1500), breaks=c(0,500,1000,1500,min_exp_b2), 
                label=c("0","500","1000","1500",round(min_exp_b2,digits=1)))
```
  
There is no significant relationship between cumulative exposure to deposited sediment and the Hedges' d effect size for coral recruit survival rate (DRMA z = -1.833, p = 0.067), after accounting for variability by comparison (Fig. X, Table X).
  
#### iii. MG/CM2/DAY IS PREDICTOR, Model using manual calculation of Hedges' *d*
Now I will predict the effect size along the exposure range and plot with confidence intervals.  

```{r, echo=FALSE, warning=FALSE}
summary(mod_DS_recruit_5c)

# PREDICT THE EFFECT SIZE FOR 13.2 mg/cm2 FROM TWO MODELS (LOAEL identified by binary analyses)
#predict(mod_DS_recruit_5c, newdata=data.frame(Sed_level_standardized=13.2), ci=TRUE)

# PREDICT THE EFFECT SIZE AND PLOT WITH CONFIDENCE INTERVALS
pred_mod_DS_recruit_5c <- predict(mod_DS_recruit_5c, newdata=ES_DS_recruit2, ci=TRUE)
ES_DS_recruit2_CI3 <- cbind(ES_DS_recruit2, pred_mod_DS_recruit_5c)
#head(ES_DS_recruit2_CI3)
min_conc_c <- ES_DS_recruit2_CI3 %>% 
  mutate(overlap0 = 0 >= ci.lb & 0 <= ci.ub) %>% 
  filter(overlap0==FALSE) %>% 
  summarize(min_conc_c=min(Sed_level_standardized))
min_conc_c2 <- as.numeric(min_conc_c)

ggplot(ES_DS_recruit2_CI3, 
       aes(x = Sed_level_standardized,
           y = Hedges_d,
           color = Ref,
           ymin = Hedges_d-Var_d,
           ymax = Hedges_d+Var_d)) + 
  geom_pointrange() +
  labs(x = expression("Sediment exposure concentration (mg/cm"^"2"*"/day)"),
       y = expression("Effect size (Hedges'"~italic(d)~"+/- variance of"~italic(d)~")"),
       color = "Study") +
  geom_abline(intercept=0, slope=0) +
  geom_vline(xintercept=min_conc_c2, linetype="dashed", color = "red") +
  geom_ribbon(aes(x = Sed_level_standardized, y = Hedges_d,
                  ymin = ci.lb, ymax = ci.ub), 
              fill = "grey70", alpha = .15, 
              show.legend=FALSE, inherit.aes=FALSE) +
  geom_line(aes(x = Sed_level_standardized, y = fit), inherit.aes=FALSE) +
  theme_classic() +
  scale_x_continuous(limits = c(0,40), breaks=c(0,10,20,30,40,min_conc_c2), 
                label=c("0","10","20","30","40",round(min_conc_c2,digits=1)))
```
  
#### iv. MG/CM2/DAY IS PREDICTOR, Model using dosresmeta calculation of Hedges' *d*
Now I will predict the effect size along the exposure range and plot with confidence intervals.  

```{r, echo=FALSE, warning=FALSE}
summary(mod_DS_recruit_5d)

# PREDICT THE EFFECT SIZE FOR 13.2 mg/cm2 FROM TWO MODELS (LOAEL identified by binary analyses)
#predict(mod_DS_recruit_5d, newdata=data.frame(Sed_level_standardized=13.2), ci=TRUE)

# PREDICT THE EFFECT SIZE AND PLOT WITH CONFIDENCE INTERVALS
pred_mod_DS_recruit_5d <- predict(mod_DS_recruit_5d, newdata=ES_DS_recruit2, ci=TRUE)
ES_DS_recruit2_CI4 <- cbind(ES_DS_recruit2, pred_mod_DS_recruit_5d)
#head(ES_DS_recruit2_CI4)
min_conc_d <- ES_DS_recruit2_CI4 %>% 
  mutate(overlap0 = 0 >= ci.lb & 0 <= ci.ub) %>% 
  filter(overlap0==FALSE) %>% 
  summarize(min_conc_d=min(Sed_level_standardized))
min_conc_d2 <- as.numeric(min_conc_d)

ggplot(ES_DS_recruit2_CI4, 
       aes(x = Sed_level_standardized,
           y = smd,
           color = Ref,
           ymin = smd-vmd,
           ymax = smd+vmd)) + 
  geom_pointrange() +
  labs(x = expression("Sediment exposure concentration (mg/cm"^"2"*"/day)"),
       y = expression("Effect size (Hedges'"~italic(d)~"+/- variance of"~italic(d)~")"),
       color = "Study") +
  geom_abline(intercept=0, slope=0) +
  geom_vline(xintercept=min_conc_d2, linetype="dashed", color = "red") +
  geom_ribbon(aes(x = Sed_level_standardized, y = smd,
                  ymin = ci.lb, ymax = ci.ub), 
              fill = "grey70", alpha = .15, 
              show.legend=FALSE, inherit.aes=FALSE) +
  geom_line(aes(x = Sed_level_standardized, y = fit), inherit.aes=FALSE) +
  theme_classic() +
  scale_x_continuous(limits = c(0,40), breaks=c(0,10,20,30,40,min_conc_d2), 
                label=c("0","10","20","30","40",round(min_conc_d2,digits=1)))
```
  
For every 1-unit increase in exposure concentration of deposited sediment, the Hedges' d effect size for coral recruit survival rate declined by 0.02 (95% CI -0.04, 0.00; DRMA z = -2.2410, p = 0.025), after accounting for variability by comparison (Fig. X, Table X). However, the best-fit model’s I2 statistic was 47.1%, indicating some residual heterogeneity unaccounted for by the model (Table X), which could be the result of taxonomic, geographic, and/or mineralogical differences among (and within) studies.
  