---
title: "Binary effects of sediment on coral fertilization"
author: "Lillian J. Tuttle"
date: "9/24/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(plotly)
library(lme4)
library(pROC)
library(effects)
library(lattice)
library(MuMIn)
binary <- read.csv("data/binary_ALL.csv", header=T)

#keyboard shortcuts to remember: 
#command-option-I inserts new code chunk
#command-shift-M inserts pipeline symbols
#command-enter runs line of code where cursor is (or highlighted portion)
```

#### Here I will explore how SUSPENDED sediment affects the FERTILIZATION SUCCESS of tropical, scleractinean corals. The database to be used is comprised of data extracted from studies deemed relevant during the systematic literature review process.

# The Dataset
 
Specifically, we will explore the results of 110 datapoints from 6 studies conducted in 2 oceans on 4 species within 2 genera.  

```{r, echo=FALSE, warning=FALSE}
#SETTLEMENT
#head(fertSS)
fertSS <- binary %>%
  mutate(Gsp = paste(Updated_Genus, Updated_species, sep = "_")) %>% 
  filter(!is.na(Binary_reduced_fertilization))
fertSS %>% tally() #n=110 datapoints, tally() returns same result as count(fertSS)
fertSS %>% count(Ref, sort=T) #n=10 studies
fertSS %>% count(Ref_name, sort=T) #n=6 studies, count(Ref_name) = group_by(Ref_name) %>% tally()
fertSS %>% count(Ocean, sort=T) #n=2 oceans
fertSS %>% count(Gsp) #n=4 species
fertSS %>% count(Updated_Genus, sort=T) #n=2 genera
```
  
  
# Exploratory plots

First let's explore all data from all species for which there is data about 'fertilization' as a result of exposure to suspended sediment.  

##### DEFINITIONS:  
* '**Fertilization success**' is defined as successful fertilization of coral eggs (existence of gamete) when exposed to sediment, as compared to surfaces without sediment (or "control" conditions).  

```{r, echo=FALSE, warning=FALSE}
#SUSPENDED SEDIMENT - Fertilization
fertSS_plot2 <- ggplot(data = fertSS) +
    geom_point(mapping = aes(
      x = Sed_level_stand_mg, 
      y = Binary_reduced_fertilization,
      color = Gsp)) +
    labs(x = "Sediment exposure concentration (mg/L)", 
         y = "Reduced fertilization success\ndue to suspended sediment? (0/1)",
         color = "Species") +
    scale_x_log10(breaks=c(0.01,0.1,1,10,100,1000),
                  label=c("0.01","0.1","1","10","100","1000")) +
    scale_y_continuous(breaks=c(0,1)) +
    theme_bw() +
    scale_color_discrete(labels = c(expression(italic("Acropora digitifera")),expression(italic("Acropora millepora")),expression(italic("Acropora tenuis")),expression(italic("Pectina lactusa")))) + theme(legend.text.align = 0)

fertSS_plot3 <- ggplot(data = fertSS) +
    geom_point(mapping = aes(
      x = Sed_level_stand_mg, 
      y = Binary_reduced_fertilization,
      color = Gsp, size = 3)) +
    labs(x = "", y = "", color = "") +
    scale_x_log10(breaks=c(0.01,0.1,1,10,100,1000),
                  label=c("0.01","0.1","1","10","100","1000")) +
    scale_y_continuous(breaks=c(0,1)) +
    scale_color_discrete(labels = c("","","","")) + 
    guides(size = FALSE, color = guide_legend(override.aes = list(size=5))) +
    theme_bw() +
    theme(axis.text.x=element_text(size=rel(2)), 
          axis.text.y=element_text(size=rel(2)),
          legend.key.size = unit(4,"line"))

#plot for Fig. 2A
fertSS_plot4 <- ggplot(data = fertSS) +
    geom_point(mapping = aes(
      x = Sed_level_stand_mg, 
      y = Binary_reduced_fertilization, size = 5)) +
    labs(x = "", y = "") +
    scale_x_log10(limits = c(1,1000), breaks=c(0.01,0.1,1,10,100,1000),
                  label=c("0.01","0.1","1","10","100","1000")) + 
    guides(size = FALSE) +
    theme_bw() +
    theme(axis.text.x=element_text(size=rel(3)), 
          axis.text.y=element_text(size=rel(3)))
    

#PLOTS:
fertSS_plot2 + labs(title = "Fertilization Success and Suspended Sediment", 
                    subtitle = "n=110 datapoints, n=10 studies, n=2 genera")
fertSS_plot3
fertSS_plot4 + scale_y_continuous(breaks=c(0,1)) +
  geom_vline(xintercept=30.4, linetype="dashed", color = "red", size = 2)
```
  
  
# Threshold Analyses with Binary Data 
Now let's calculate the thresholds, based on the binary data explored above.  

##### DEFINITIONS:  
* '**LOAEL**', or the 'lowest observed adverse effect level' is defined as the lowest dose at which there was an observed adverse effect.  
* '**NOAEL**', or the 'no observed adverse effect level' is defined as the highest dose at which there was NOT an observed adverse effect.  
* '**Adverse effect**' is defined here as any response of a coral individual, colony, or experimental treatment group that *may* negatively affect the coral’s fitness and/or survival. These adverse effects may include sublethal physiological changes (e.g., significantly reduced growth or photosynthetic rates when compared to coral in ambient/control conditions), bleaching/tissue loss, and mortality.  This definition of adverse effect is independent of its magnitude, so while the affect may have potentially reduced fitness, its magnitude may be sufficiently small that the fitness effect is not measureable.  
  
  
```{r, echo=FALSE, warning=FALSE}
#Setting up for LOAEL by filtering for adverse effects (binary=1)
adverse_fertSS <- fertSS %>% 
  filter(Binary_reduced_fertilization=="1") %>%
  arrange(Sed_level_stand_mg)

#LOAEL
summarize(adverse_fertSS, LOAEL=min(Sed_level_stand_mg)) #LOAEL is 30.4 mg/L for 0.0625 days and comes from Humanes et al. 2017b: Acropora tenuis
loael <- adverse_fertSS %>% summarize(min(Sed_level_stand_mg))
loael <- as.numeric(loael)

#Setting up for NOAEL by filtering for non-adverse effects (binary=0), and then by any sediment level less than or equal to the LOAEL
non_adverse_fertSS <- fertSS %>%
  filter(Binary_reduced_fertilization=="0") %>%
  arrange(desc(Sed_level_stand_mg)) %>%
  filter(Sed_level_stand_mg<=30.4) #less than or equal to the LOAEL, which = 30.4

#NOAEL = maximum sediment level with no adverse effect
summarize(non_adverse_fertSS, NOAEL=max(Sed_level_stand_mg)) #NOAEL is 25

#PLOT
fertSS_plot2 + annotate("rect", xmin=30.4, xmax=Inf, 
                         ymin=-Inf, ymax=Inf, fill = "red", alpha=0.2) +
  labs(title = "Limited Fertilization from Suspended Sediment", 
       subtitle = "n = 110 datapoints, n = 10 studies from 6 articles, n = 4 spp. from 2 genera", 
       caption = "LOAEL concentration = 30.4\nNOAEL concentration = 25")

fertSS_plot2 + annotate("rect", xmin=30.4, xmax=Inf, 
                         ymin=-Inf, ymax=Inf, fill = "red", alpha=0.2) +
  labs(caption = "LOAEL concentration = 30.4, NOAEL concentration = 25.0\nn = 110 datapoints, n = 10 studies from 6 articles, n = 4 spp. from 2 genera")

fertSS_plot3 + annotate("rect", xmin=30.4, xmax=Inf, 
                         ymin=-Inf, ymax=Inf, fill = "red", alpha=0.2)
```
     
# Logistic Regression Model Fitting
```{r, echo=FALSE, warning=FALSE}
#Let's remove controls for which sediment level was zero, and add a log transformation of sediment level, in case it's needed
fertSS2 <- fertSS %>% 
  select(Ref, Ref_name, Binary_reduced_fertilization, Control_code, Sed_level_stand_mg, Sed_exposure_days, Gsp) %>% 
  mutate(log10_mg_L=log10(Sed_level_stand_mg))

#Models with fixed effect(s) only
mod1 <- glm(Binary_reduced_fertilization ~ Sed_level_stand_mg, 
            family = binomial(link = "logit"), data = fertSS2)
mod2 <- glm(Binary_reduced_fertilization ~ Sed_level_stand_mg + Sed_exposure_days, 
            family = binomial(link = "logit"), data = fertSS2)
mod3 <- glm(Binary_reduced_fertilization ~ Sed_level_stand_mg*Sed_exposure_days, 
            family = binomial(link = "logit"), data = fertSS2)

#Models with random effect for coral species or study
#In this case, exposure duration is not so important, but I've included models with this as a fixed effect, just to maintain consistency with other coral responses for which exposure duration may matter more
mod4 <- glmer(Binary_reduced_fertilization ~ Sed_level_stand_mg + (1 | Gsp), 
              family = binomial, data = fertSS2)
mod5 <- glmer(Binary_reduced_fertilization ~ Sed_level_stand_mg + Sed_exposure_days + (1 | Gsp), 
              family = binomial, data = fertSS2)
mod6 <- glmer(Binary_reduced_fertilization ~ Sed_level_stand_mg*Sed_exposure_days + (1 | Gsp), 
              family = binomial, data = fertSS2)
mod7 <- glmer(Binary_reduced_fertilization ~ Sed_level_stand_mg + (1 | Ref), 
              family = binomial, data = fertSS2)
mod8 <- glmer(Binary_reduced_fertilization ~ Sed_level_stand_mg + Sed_exposure_days + (1 | Ref), 
              family = binomial, data = fertSS2)
mod9 <- glmer(Binary_reduced_fertilization ~ Sed_level_stand_mg*Sed_exposure_days + (1 | Ref), 
              family = binomial, data = fertSS2)
mod10 <- glmer(Binary_reduced_fertilization ~ Sed_level_stand_mg + (1 | Gsp) + (1 | Ref), 
              family = binomial, data = fertSS2)
mod11 <- glmer(Binary_reduced_fertilization ~ 
                Sed_level_stand_mg + Sed_exposure_days + (1 | Gsp) + (1 | Ref), 
              family = binomial, data = fertSS2)
mod12 <- glmer(Binary_reduced_fertilization ~ 
                Sed_level_stand_mg*Sed_exposure_days + (1 | Gsp) + (1 | Ref), 
              family = binomial, data = fertSS2)
mod13 <- glmer(Binary_reduced_fertilization ~ Sed_level_stand_mg + (1 | Ref_name/Ref), 
              family = binomial, data = fertSS2)
mod14 <- glmer(Binary_reduced_fertilization ~ 
                 Sed_level_stand_mg + Sed_exposure_days + (1 | Ref_name/Ref), 
              family = binomial, data = fertSS2)
mod15 <- glmer(Binary_reduced_fertilization ~ 
                 Sed_level_stand_mg*Sed_exposure_days + (1 | Ref_name/Ref), 
              family = binomial, data = fertSS2)
mod16 <- glmer(Binary_reduced_fertilization ~ 
                 Sed_level_stand_mg + (1 | Gsp) + (1 | Ref_name), 
              family = binomial, data = fertSS2)
mod17 <- glmer(Binary_reduced_fertilization ~ 
                 Sed_level_stand_mg + Sed_exposure_days + (1 | Gsp) + (1 | Ref_name), 
              family = binomial, data = fertSS2)
mod18 <- glmer(Binary_reduced_fertilization ~ 
                 Sed_level_stand_mg*Sed_exposure_days + (1 | Gsp) + (1 | Ref_name), 
              family = binomial, data = fertSS2)
mod19 <- glmer(Binary_reduced_fertilization ~ 
                 Sed_level_stand_mg + (1 | Gsp) + (1 | Ref_name/Ref), 
              family = binomial, data = fertSS2)
mod20 <- glmer(Binary_reduced_fertilization ~ 
                 Sed_level_stand_mg + Sed_exposure_days + (1 | Gsp) + (1 | Ref_name/Ref), 
              family = binomial, data = fertSS2)
mod21 <- glmer(Binary_reduced_fertilization ~ 
                 Sed_level_stand_mg*Sed_exposure_days + (1 | Gsp) + (1 | Ref_name/Ref), 
              family = binomial, data = fertSS2)

#Model comparison and summary
anova(mod6,mod9,mod12,mod15,mod18,mod21) #compare full fixed effects models -- mod9 is best
anova(mod7,mod8,mod9) #No evidence against reduced model (mod7)
anova(mod4,mod7,mod10,mod13,mod16) #compare reduced fixed effects models -- mod7 is best
print(summary(mod7), correlation=FALSE) 
#Model is nearly unidentifiable: very large eigenvalue - Rescale variables?

#Trying out a log transformation of predictor...
fertSS3 <- fertSS2 %>% filter(Control_code=="0")
fertSS3 %>% tally() #n=86
mod7_log <- glmer(Binary_reduced_fertilization ~ log10_mg_L + (1 | Ref), 
                  family = binomial, data = fertSS3)
#print(summary(mod7_log), correlation=FALSE) #No errors!

#Changing to random slopes/intercept
mod7_log2 <- glmer(Binary_reduced_fertilization ~ log10_mg_L + (log10_mg_L | Ref), 
                   family = binomial, data = fertSS3)
anova(mod7_log,mod7_log2) #original (log) still best

print(summary(mod7_log), correlation=FALSE)
#The response is normally(?) distributed across studies with a standard deviation of 2.508. In other words, the typical difference between studies in the response is 2.508.

```
  
# Model Diagnostics, Random Effects, and Effect Estimates
```{r, echo=FALSE, warning=FALSE}
##Diagnostics
#predicted probabilities, converted to 0 or 1 respectively:
p <- as.numeric(predict(mod7_log, type="response")>0.5)
#proportion correct:
mean(p==fertSS3$Binary_reduced_fertilization) #[1] 0.8837209
#predicted-vs-observed table:
table(p,fertSS3$Binary_reduced_fertilization)
#ROC and area under the curve:
pred_mod7_log <- predict(mod7_log, newdata = fertSS3, type = "response")
auc(roc(fertSS3$Binary_reduced_fertilization, pred_mod7_log)) #AUC: 0.9534
#can also check out 'performance' package https://cran.r-project.org/web/packages/performance/performance.pdf
r.squaredGLMM(mod7_log) #The fixed effect of sediment level explains ~40% of variation of the total variation. If the random intercepts for studies are also included, then 80% of variation is explained.

##Exploration of random effect(s)
plot(allEffects(mod7_log))
dotplot(ranef(mod7_log, condVar = T)) #residual variation at the level of studies -- variation among studies is large
qqnorm(ranef(mod7_log)$Ref[,1])
qqline(ranef(mod7_log)$Ref[,1]) #distribution of random effects looks good
VarCorr(mod7_log) #SD=2.5075
exp(sqrt(VarCorr(mod7_log)$Ref[1])) #I think this means that two observations from different studies typically differ by a factor of exp(2.5) = 12.27421 (?)
mod7_log_noSed <- glmer(Binary_reduced_fertilization ~ 1 + (1 | Ref), 
                  family = binomial, data = fertSS3)
random.intercepts <- ranef(mod7_log)$Ref[["(Intercept)"]]
Ref.intercepts <- fixef(mod7_log)[1] + random.intercepts 
#plot with probability of binary response on y-axis, each line representing a study:
for (i in 1:length(random.intercepts)) {
  if (i == 1) curve(exp(Ref.intercepts[i] + fixef(mod7_log)["log10_mg_L"]*x)/(1 + exp(Ref.intercepts[i] + fixef(mod7_log)["log10_mg_L"]*x)), 
                    from = min(fertSS3$log10_mg_L), to = max(fertSS3$log10_mg_L), 
                    ylim = c(0, 1), 
                    ylab = "Probability of adverse effect", 
                    xlab = "log10(exposure concentration, mg/L)")
  if (i > 1) curve(exp(Ref.intercepts[i] + fixef(mod7_log)["log10_mg_L"]*x)/(1 + exp(Ref.intercepts[i] + fixef(mod7_log)["log10_mg_L"]*x)), add = T) 
}

##Effect Estimates
#https://stats.stackexchange.com/questions/8318/interpretation-of-log-transformed-predictors-in-logistic-regression
#"Case 2: base k transformed independent variable: Then the exponentiated coefficient, exp(Beta), can be interpreted as the proportionate increase in the odds from a k-fold increase in the independent variable." -- For k=10, exp(Beta) is the multiplicative change in the odds of success with a 10-fold increase in X1 (with X2 fixed, if relevant).
se <- sqrt(diag(vcov(mod7_log)))
tab <- cbind(Est = fixef(mod7_log), 
             LL = fixef(mod7_log) - 1.96 * se, 
             UL = fixef(mod7_log) + 1.96 * se)
exp(tab)
#                     Est           LL           UL
#(Intercept) 6.686841e-04 7.221468e-06   0.06191795
#log10_mg_L  5.471313e+01 6.205822e+00 482.37388230
```
For every 10-fold increase in sediment exposure, the odds of reduced fertilization success increase by 54.7 times (95% CI 6.2,482.4, GLMM p=0.0003), after accounting for the (random) effect of study. The fixed effect of sediment level explains 40.3% of variation of the total variation. If the random intercepts for studies are also included, then 79.5% of variation is explained.  
    
# Prediction figure  

```{r}
#Overlaying glmm results on figure, but prediction line is jagged...
pred_mod7_log <- predict(mod7_log, newdata=fertSS3, type="response")
fertSS4 <- cbind(fertSS3, pred_mod7_log)

fertSS_plot3 <- ggplot(data = fertSS4) +
    geom_point(mapping = aes(
      x = Sed_level_stand_mg, 
      y = Binary_reduced_fertilization,
      color = Gsp)) +
    labs(x = "Sediment exposure concentration (mg/L)", 
         y = "Reduced fertilization success\ndue to suspended sediment? (0/1)",
         color = "Species") +
    scale_x_log10(breaks=c(0.01,0.1,1,10,100,1000),
                  label=c("0.01","0.1","1","10","100","1000")) +
    scale_y_continuous(breaks=c(0,1)) +
    geom_line(aes(x = Sed_level_stand_mg, y = pred_mod7_log), inherit.aes=FALSE) +
    theme_bw() +
    scale_color_discrete(labels = c(expression(italic("Acropora digitifera")),expression(italic("Acropora millepora")),expression(italic("Acropora tenuis")),expression(italic("Pectina lactusa")))) + theme(legend.text.align = 0)

fertSS_plot3
```

That plot is confusing to interpret. Let's see if I can plot the average marginal probability, i.e., the average change in probability of the outcome across the range of the predictor of interest. This is described in some detail at the following, useful website: https://stats.idre.ucla.edu/r/dae/mixed-effects-logistic-regression/
  
```{r}
#summary(fertSS3$log10_mg_L)
#   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
# 0.6021  1.5051  1.9657  1.8956  2.3617  3.0103 

jvalues <- with(fertSS3, seq(from = min(log10_mg_L), to = max(log10_mg_L), length.out = 100))

# calculate predicted probabilities and store in a list
pp <- lapply(jvalues, function(j) {
    fertSS3$log10_mg_L <- j
    predict(mod7_log, newdata = fertSS3, type = "response")
})

# average marginal predicted probability across a few different sediment levels
#sapply(pp[c(1, 20, 40, 60, 80, 100)], mean)
## [1] 0.01857219 0.07270104 0.17399541 0.33310503 0.58820472 0.85682467
# get the means with lower and upper quartiles
plotdat <- t(sapply(pp, function(x) {
    c(M = mean(x), med = median(x), quantile(x, c(0.25, 0.75)))
}))

# add in log10_mg_L values and convert to data frame
plotdat <- as.data.frame(cbind(plotdat, jvalues))
plotdat <- plotdat %>% mutate(Sed_level_linear=10^jvalues)

# better names and show the first few rows
colnames(plotdat) <- c("PredictedProbability", "MedianProbability", 
                       "LowerQuantile", "UpperQuantile", "log10_mg_L", "mg_L")
#head(plotdat)

# plot average marginal predicted probabilities
ggplot(plotdat, aes(x = log10_mg_L)) + 
  geom_line(aes(y = PredictedProbability), size = 2) +
  geom_line(aes(y = MedianProbability), size = 0.5) +
  ylim(c(0, 1))
ggplot(plotdat, aes(x = log10_mg_L)) + 
  geom_ribbon(aes(ymin = LowerQuantile, ymax = UpperQuantile), alpha = 0.15) +
  geom_line(aes(y = PredictedProbability), size = 2) +
  geom_line(aes(y = MedianProbability), size = 0.5) + 
  ylim(c(0, 1))
ggplot(plotdat, aes(x = mg_L, y = PredictedProbability)) + 
  geom_ribbon(aes(ymin = LowerQuantile, ymax = UpperQuantile), alpha = 0.15) +
  geom_line(aes(y = PredictedProbability), size = 2) +
  geom_line(aes(y = MedianProbability), size = 0.5) + 
  ylim(c(0, 1)) +
  scale_x_log10()

#Overlaying average marginal predicted probabilities on figure
fertSS_plot3 <- ggplot() +
    geom_point(data = fertSS3, mapping = aes(
      x = Sed_level_stand_mg, 
      y = Binary_reduced_fertilization,
      color = Ref)) +
    labs(x = "Sediment exposure concentration (mg/L)", 
         y = "Predicted probability of reduced\nfertilization success due to sediment exposure",
         color = "Binary Data\nby Study",
         linetype = "Predicted\nProbability") +
    scale_x_log10(breaks=c(0.01,0.1,1,10,100,1000,loael),
                  label=c("0.01","0.1","1","10","100","1000",round(loael,digits=1))) +
    geom_ribbon(data = plotdat, aes(x = mg_L, y = PredictedProbability, 
                                    ymin = LowerQuantile, ymax = UpperQuantile), 
                alpha = 0.15) +
    geom_line(data = plotdat, aes(x = mg_L, y = PredictedProbability, linetype = "twodash")) +
    geom_line(data = plotdat, aes(x = mg_L, y = MedianProbability, linetype = "solid")) +
    geom_vline(xintercept=loael, linetype="dashed", color = "red") +
    theme_bw() +
    scale_linetype_manual(values=c("twodash", "solid"), labels = c("Median","Mean"))

#And now by Ref instead of species
fertSS_plot5 <- ggplot() +
    geom_point(data = fertSS3, mapping = aes(
      x = Sed_level_stand_mg, 
      y = Binary_reduced_fertilization,
      color = Ref,
      shape = Gsp)) +
    labs(x = "Sediment exposure concentration (mg/L)", 
         y = "Predicted probability of reduced\nfertilization success due to sediment exposure",
         color = "Study",
         shape = "Species") +
    scale_x_log10(breaks=c(0.01,0.1,1,10,100,1000),
                  label=c("0.01","0.1","1","10","100","1000")) +
    geom_ribbon(data = plotdat, aes(x = mg_L, y = PredictedProbability, 
                                    ymin = LowerQuantile, ymax = UpperQuantile), 
                alpha = 0.15) +
    geom_line(data = plotdat, aes(x = mg_L, y = PredictedProbability, linetype = "twodash")) +
    geom_line(data = plotdat, aes(x = mg_L, y = MedianProbability, linetype = "solid")) +
    theme_bw() +
    scale_shape_discrete(labels = c(expression(italic("Acropora digitifera")),expression(italic("Acropora millepora")),expression(italic("Acropora tenuis")),expression(italic("Pectina lactusa")))) + theme(legend.text.align = 0) + 
    guides(linetype = FALSE, color = guide_legend(order=1), shape = guide_legend(order=2))

fertSS_plot3
fertSS_plot5 + labs(caption = "LOAEL concentration = 30.4\nNOAEL concentration = 25.0")
#Plot for Fig. 2B
fertSS_plot4 + scale_y_continuous(breaks=c(0,0.25,0.5,0.75,1)) +
  geom_ribbon(data = plotdat, 
              aes(x = mg_L, y = PredictedProbability, 
                  ymin = LowerQuantile, ymax = UpperQuantile), 
              alpha = 0.15) +
  geom_line(data = plotdat, aes(x = mg_L, y = MedianProbability, linetype = "solid", size = 2)) +
  theme(legend.position = "none")
```

