---
title: "Effects of sediment on the growth of adult corals"
author: "Lillian J. Tuttle"
date: "7/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(mixmeta)
library(dosresmeta)
library(splines)

#will want to getwd() and setwd(), as appropriate
effectsize <- read.csv("/Volumes/GoogleDrive/My Drive/iScience/Code/thresholds/sedimentation/effectsize2.csv", header=T)

#keyboard shortcuts to remember: 
#command-option-I inserts new code chunk
#command-shift-M inserts pipeline symbols
#command-enter runs line of code where cursor is (or highlighted portion)
```
  
```{r, echo=FALSE, warning=FALSE}
#Looking at data
#head(effectsize) 

#a lot of variables (columns) are factors instead of numeric/doubles/integers
effectsize <- effectsize %>% #converting appropriate columns to numeric...
  mutate_at(vars(ES, Hedges_d, Var_d, SE_d, CI_lo_d, CI_up_d, Sed_level_standardized, Sed_exposure, Sed_exposure_mg_cm2), as.character) %>% 
  mutate_at(vars(ES, Hedges_d, Var_d, SE_d, CI_lo_d, CI_up_d, Sed_level_standardized, Sed_exposure, Sed_exposure_mg_cm2), as.numeric)
#head(effectsize) #"double"-check!

#Now create a new variable that has both Genus and species, remove sediment burial experiments, and remove irrelevant rows (non-controls, -treatments, or -baselines)
effectsize <- effectsize %>%
  mutate(Gsp = paste(Updated_Genus, Updated_species, sep = "_")) %>% 
  filter(Binary_sed_burial=="0") %>%
  filter(!Control=="-1")
#head(effectsize)

#Separate into deposited sediment (DS) and suspended sediment (SS) datasets, WITH CONTROLS
ES_DS <- effectsize %>% filter(DS_SS=="DS")
ES_SS <- effectsize %>% filter(DS_SS=="SS")

#Second dataset that that removes controls (Control == c("1","2"))
effectsize2 <- effectsize %>% filter(Control=="0")
#effectsize2 #"double"-check!

#Separate into deposited sediment (DS) and suspended sediment (SS) datasets, WITHOUT CONTROLS
ES_DS2 <- effectsize2 %>% filter(DS_SS=="DS")
ES_SS2 <- effectsize2 %>% filter(DS_SS=="SS")

#Now separate into independent dataset for each coral response for DEPOSITED SEDIMENT
ES_DS_growth <- ES_DS %>% filter(Response == "growth")
ES_DS_growth %>% tally()
ES_DS_growth %>% count(Ref)
ES_DS_growth %>% count(Ref_name)
ES_DS_growth %>% count(Gsp)
ES_DS_growth %>% count(Updated_Genus)
``` 
  
These datasets contain information from multiple articles (**Ref_name**) and studies/experiments (**Ref**) within those articles about the effects of DEPOSITED sediment on the **growth rate** of adult corals. I would like to explore the dose-response relationship between sediment 'dose' and the effect size, estimated here as the standardized mean difference between treatment and control, in terms of Hedges' *d*:

- Hedges' *d* is the standardized mean difference between the treatment and control groups (Hedges and Olkin 1985). Each study may have multiple Hedges' *d* calculations, one for each treatment-control comparison at each time-point.   

- 'Dose' can be quantified as sediment exposure concentration (mg/cm^2^/day), exposure duration (days), or as "cumulative exposure", for which I multiply concentration and duration (mg/cm^2^).

Each study/experiment (Ref) may contain multiple sets of controls and treatments (due to using multiple species or sediment types, for instance). Thus, I have added two categorical vectors, '**Comparison**' and 'Control', which group each set of control-treatment comparisons into a unique two-letter code, within which is one control (Control=="1") and at least one treatment (Control=="0"). **Therefore, 'Comparison' is nested within 'Ref', which is further nested within 'Ref_name'.**

```{r, echo=FALSE}
#Using dosresmeta to calculate Hedges' d and variance of d
covar_DS_growth <- by(ES_DS_growth, ES_DS_growth$Comparison, function(x) 
  covar.smd(Tx_mean, Tx_sd, Tx_n, "smd", method="hedges", data = x))
ES_DS_growth$smd <- unlist(lapply(covar_DS_growth, function(x) x$y))
ES_DS_growth$vmd <- unlist(lapply(covar_DS_growth, function(x) x$v))

#Looking at data to make sure smd and vmd added as columns
#ES_DS_growth

#Removing controls for modeling
ES_DS_growth2 <- subset(ES_DS_growth, Control=="0")

#Looking at data to make sure controls were removed
#ES_DS_growth2
ES_DS_growth2 %>% tally()
``` 
  
# I. Growth Rate  

## A. Deposited Sediment  
  
### 1. Plots  
I calculated the standardized mean difference in terms of Hedges' *d* and plot it here with respect to cumulative exposure. I did this first for a Hedges' *d* that I calculated before importing the data to R and again for a Hedges' *d* that is calculated by the dosresmeta::covar.smd function.

```{r, echo=FALSE, warning=FALSE}
#CUMULATIVE EXPOSURE vs. effect size, by study
#Using previously calculated Hedges' d and variance of d
ggplot(ES_DS_growth2, 
       aes(x = Sed_exposure_mg_cm2,
           y = Hedges_d,
           color = Ref,
           ymin = Hedges_d-Var_d,
           ymax = Hedges_d+Var_d)) + 
  geom_pointrange() +
  ggtitle("Growth rate") +
  labs(x = "Sediment cumulative exposure (mg/cm2)",
       y = "Effect size (Hedges' d +/- s.d., manual calculation)",
       color = "Study") +
  geom_abline(intercept=0, slope=0) +
  theme_classic() +
  scale_x_log10(limits = c(10,40000), breaks=c(10,100,1000,10000), 
                label=c("10","100","1000","10000"))
#Using dosresmeta's calculation of Hedges' d and variance of d
ggplot(ES_DS_growth2, 
       aes(x = Sed_exposure_mg_cm2,
           y = smd,
           color = Ref,
           ymin = smd-vmd,
           ymax = smd+vmd)) + 
  geom_pointrange() +
  ggtitle("Growth rate") +
  labs(x = "Sediment cumulative exposure (mg/cm2)",
       y = "Effect size (Hedges' d +/- s.d., dosresmeta calculation)",
       color = "Study") +
  geom_abline(intercept=0, slope=0) +
  theme_classic() +
  scale_x_log10(limits = c(10,40000), breaks=c(10,100,1000,10000), 
                label=c("10","100","1000","10000"))

#EXPOSURE CONCENTRATION vs. effect size, by study
#Using previously calculated Hedges' d and variance of d
ggplot(ES_DS_growth2, 
       aes(x = Sed_level_standardized,
           y = Hedges_d,
           color = Ref,
           ymin = Hedges_d-Var_d,
           ymax = Hedges_d+Var_d)) + 
  geom_pointrange() +
  ggtitle("Growth rate") +
  labs(x = "Sediment exposure concentration (mg/cm2/day)",
       y = "Effect size (Hedges' d +/- s.d., manual calculation)",
       color = "Study") +
  geom_abline(intercept=0, slope=0) +
  theme_classic() 
#Using dosresmeta's calculation of Hedges' d and variance of d
ggplot(ES_DS_growth2, 
       aes(x = Sed_level_standardized,
           y = smd,
           color = Ref,
           ymin = smd-vmd,
           ymax = smd+vmd)) + 
  geom_pointrange() +
  ggtitle("Growth rate") +
  labs(x = "Sediment exposure concentration (mg/cm2/day)",
       y = "Effect size (Hedges' d +/- s.d., dosresmeta calculation)",
       color = "Study") +
  geom_abline(intercept=0, slope=0) +
  theme_classic() 


#EXPOSURE CONCENTRATION vs. effect size, by SEDIMENT TYPE
#Using previously calculated Hedges' d and variance of d
ggplot(ES_DS_growth2, 
       aes(x = Sed_level_standardized,
           y = smd,
           color = Sed_grain_category,
           shape = Sed_source,
           ymin = smd-vmd,
           ymax = smd+vmd)) + 
  geom_pointrange() +
  ggtitle("Growth rate") +
  labs(x = "Sediment exposure concentration (mg/cm2/day)",
       y = "Effect size (Hedges' d +/- s.d., dosresmeta calculation)",
       color = "Sediment relative\ngrain size",
       shape = "Sediment source") +
  geom_abline(intercept=0, slope=0) +
  theme_classic()
ggplot(ES_DS_growth2, 
       aes(x = Sed_level_standardized,
           y = smd,
           color = Sed_grain_class,
           shape = Sed_mineralogy,
           ymin = smd-vmd,
           ymax = smd+vmd)) + 
  geom_pointrange() +
  ggtitle("Growth rate") +
  labs(x = "Sediment exposure concentration (mg/cm2/day)",
       y = "Effect size (Hedges' d +/- s.d., dosresmeta calculation)",
       color = "Sediment\ngrain class",
       shape = "Sediment mineralogy") +
  geom_abline(intercept=0, slope=0) +
  theme_classic()

#EXPOSURE CONCENTRATION vs. effect size, by PHYLOGENY
#Using previously calculated Hedges' d and variance of d
ggplot(ES_DS_growth2, 
       aes(x = Sed_level_standardized,
           y = smd,
           color = Updated_Genus,
           ymin = smd-vmd,
           ymax = smd+vmd)) + 
  geom_pointrange() +
  ggtitle("Growth rate") +
  labs(x = "Sediment exposure concentration (mg/cm2/day)",
       y = "Effect size (Hedges' d +/- s.d., dosresmeta calculation)",
       color = "Genus") +
  geom_abline(intercept=0, slope=0) +
  theme_classic()

#EXPOSURE CONCENTRATION vs. effect size, by OCEAN and REGION
#Using previously calculated Hedges' d and variance of d
ggplot(ES_DS_growth2, 
       aes(x = Sed_level_standardized,
           y = smd,
           color = Region,
           shape = Ocean,
           ymin = smd-vmd,
           ymax = smd+vmd)) + 
  geom_pointrange() +
  ggtitle("Growth rate") +
  labs(x = "Sediment exposure concentration (mg/cm2/day)",
       y = "Effect size (Hedges' d +/- s.d., dosresmeta calculation)",
       color = "Region",
       shape = "Ocean") +
  geom_abline(intercept=0, slope=0) +
  theme_classic()
```
  
It is interesting that there is a difference between how I calculated Hedges' *d* and how *dosresmeta* appears to calculate it. But in both cases, there may be a negative relationship between sediment exposure and effect size. There does not appear to be strong effects of phylogeny (genus) or geography (ocean and region). There may be effects of sediment type (source, origin, grain size), but I do not think these will be discernible given the variance.
  
### 2. Model Fitting  
The following set of functions model fixed and random (nested) effects WITHOUT accounting for within-comparison (or within-study) correlations. The output for each is a univariate, random-effects meta-regression.

```{r, echo=FALSE, warning=FALSE}
#?mixmeta

# Organizational Note: 
#All models with 'a' or 'b' at the end of its name have cumulative exposure as its predictor. 
#All models with 'c' or 'd' at the end of its name have exposure concentration as its predictor. 
#All models with 'a' or 'c' at the end of its name have manually calculated Hedges' d as its response.
#All models with 'b' or 'd' at the end of its name have covar.smd calculated Hedges' d as its response.

# LINEAR FIXED AND RANDOM EFFECTS NOT ACCOUNTING FOR WITHIN-STUDY CORRELATIONS
mod_DS_growth_1a <- mixmeta(Hedges_d ~ Sed_exposure_mg_cm2, S = Var_d,
                       random =  ~ Sed_exposure_mg_cm2 | Comparison,
                       data = ES_DS_growth2, method = "ml")
mod_DS_growth_1b <- mixmeta(smd ~ Sed_exposure_mg_cm2, S = vmd,
                      random =  ~ Sed_exposure_mg_cm2 | Comparison,
                      data = ES_DS_growth2, method = "ml")
mod_DS_growth_1c <- mixmeta(Hedges_d ~ Sed_level_standardized, S = Var_d,
                       random =  ~ Sed_level_standardized | Comparison,
                       data = ES_DS_growth2, method = "ml")
mod_DS_growth_1d <- mixmeta(smd ~ Sed_level_standardized, S = vmd,
                      random =  ~ Sed_level_standardized | Comparison,
                      data = ES_DS_growth2, method = "ml")

# LINEAR FIXED AND RANDOM, NESTED EFFECTS NOT ACCOUNTING FOR WITHIN-STUDY CORRELATIONS
mod_DS_growth_2a <- mixmeta(Hedges_d ~ Sed_exposure_mg_cm2, S = Var_d,
                       random =  ~ Sed_exposure_mg_cm2 | Ref/Comparison,
                       data = ES_DS_growth2, method = "ml")
mod_DS_growth_2b <- mixmeta(smd ~ Sed_exposure_mg_cm2, S = vmd,
                      random =  ~ Sed_exposure_mg_cm2 | Ref/Comparison,
                      data = ES_DS_growth2, method = "ml")
mod_DS_growth_2c <- mixmeta(Hedges_d ~ Sed_level_standardized, S = Var_d,
                       random =  ~ Sed_level_standardized | Ref/Comparison,
                       data = ES_DS_growth2, method = "ml")
mod_DS_growth_2d <- mixmeta(smd ~ Sed_level_standardized, S = vmd,
                      random =  ~ Sed_level_standardized | Ref/Comparison,
                      data = ES_DS_growth2, method = "ml")
```

To account for within-comparison correlations in the mixmeta model, I replace the 'S' argument with a 'control' argument, which requires me to provide the covariance matrices, as done with the covar.smd function. 

```{r, echo=FALSE, warning=FALSE}
# COMPUTE THE WITHIN-COMPARISON CORRELATIONS 
# already done above but code repeated here...
#covar_DS_growth <- by(ES_DS_growth, ES_DS_growth$Comparison, function(x) 
#  covar.smd(Tx_mean, Tx_sd, Tx_n, "smd", method="hedges", data = x))

# LINEAR FIXED AND RANDOM EFFECTS ACCOUNTING FOR WITHIN-COMPARISON CORRELATIONS
#Code for covariance matrix to include for addSlist for Comparison random effect
newlist_GRO <- list(NA)
for (i in seq(1,length(covar_DS_growth))) {
newlist_GRO[i] <- list(covar_DS_growth[[i]]$S)
}
#Models with the Hedges' d I calculated (5a) and the one calculated by the package (5b)
mod_DS_growth_5a <- mixmeta(Hedges_d ~ Sed_exposure_mg_cm2,
                       random = ~ Sed_exposure_mg_cm2 | Comparison,
                       data=ES_DS_growth2, method="ml", #using dataset without controls
                       control=list(addSlist=newlist_GRO))
mod_DS_growth_5b <- mixmeta(smd ~ Sed_exposure_mg_cm2,
                       random = ~ Sed_exposure_mg_cm2 | Comparison,
                       data=ES_DS_growth2, method="ml", #using dataset without controls
                       control=list(addSlist=newlist_GRO))
mod_DS_growth_5c <- mixmeta(Hedges_d ~ Sed_level_standardized,
                       random = ~ Sed_level_standardized | Comparison,
                       data=ES_DS_growth2, method="ml", #using dataset without controls
                       control=list(addSlist=newlist_GRO))
mod_DS_growth_5d <- mixmeta(smd ~ Sed_level_standardized,
                       random = ~ Sed_level_standardized | Comparison,
                       data=ES_DS_growth2, method="ml", #using dataset without controls
                       control=list(addSlist=newlist_GRO))


# LINEAR FIXED AND RANDOM, NESTED EFFECTS ACCOUNTING FOR WITHIN-COMPARISON & WITHIN-STUDY CORRELATIONS
#Code for covariance matrix to include for addSlist for Ref/Comparison
newlist_GRO2 <- list(NA) #collects the list of covariance matrices for the block diag matrix by reference for the nested hierarchial ref/comparison
templist_GRO <- list(NA) #holds temp list of covariance matrices associated with each reference
templist_GRO2 <-list(NA)
reflista <- ES_DS_growth %>% distinct(Ref,Ref_name,Comparison)
reflist <- reflista[,1]
for (i in seq(1,length(unique(reflista$Ref))))  {
  #pull the elements from covar_DS_growth that are all from the same reference [i]
  templist_GRO[i] <-list(covar_DS_growth[reflist==unique(reflista$Ref)[i]])
  for (j in seq(1,length(templist_GRO[[i]]))) {
  #for each comparison in the reference, pull out the covar matrices (element $S) and put in into templist_GRO2
     templist_GRO2[j] <- list(templist_GRO[[i]][[j]]$S)
  }
  #turn list of covars from all comparison in one reference into block diag matrix
  newlist_GRO2[i] <- list(bdiagMat(templist_GRO2))
  templist_GRO2 <- list(NA)
}
#Error in getSlist(S, nay, groups, m, k, control$addSlist, control$checkPD) : wrong dimensions in 'addSlist'. See help(mixmetaControl)
#Models with the Hedges' d I calculated (6a) and the one calculated by the package (6b)
#mod_DS_growth_6a <- mixmeta(Hedges_d ~ Sed_exposure_mg_cm2,
#                       random = ~ Sed_exposure_mg_cm2 | Ref/Comparison,
#                       data=ES_DS_growth2, method="ml", #using dataset without controls
#                       control=list(addSlist=newlist_GRO2))
#mod_DS_growth_6b <- mixmeta(smd ~ Sed_exposure_mg_cm2,
#                       random = ~ Sed_exposure_mg_cm2 | Ref/Comparison,
#                       data=ES_DS_growth2, method="ml", #using dataset without controls
#                       control=list(addSlist=newlist_GRO2))
#mod_DS_growth_6c <- mixmeta(Hedges_d ~ Sed_level_standardized,
#                       random = ~ Sed_level_standardized | Ref/Comparison,
#                       data=ES_DS_growth2, method="ml", #using dataset without controls
#                       control=list(addSlist=newlist_GRO2))
#mod_DS_growth_6d <- mixmeta(smd ~ Sed_level_standardized,
#                       random = ~ Sed_level_standardized | Ref/Comparison,
#                       data=ES_DS_growth2, method="ml", #using dataset without controls
#                       control=list(addSlist=newlist_GRO2))
```
  
And now to compare relative fit of linear models...

```{r, echo=FALSE, warning=FALSE}
# Model Comparisons
AIC(mod_DS_growth_1a, mod_DS_growth_2a, mod_DS_growth_5a) 
#of these, model 5a has lowest AIC (but 1a within 2 AIC)

AIC(mod_DS_growth_1b, mod_DS_growth_2b, mod_DS_growth_5b) 
#of these, model 5b has lowest AIC

AIC(mod_DS_growth_1c, mod_DS_growth_2c, mod_DS_growth_5c) 
#of these, model 5c has lowest AIC (but 1c within 2 AIC)

AIC(mod_DS_growth_1d, mod_DS_growth_2d, mod_DS_growth_5d) 
#of these, model 5d has lowest AIC (but 1d within 2 AIC)
```
  
And now to evaluate residuals of models...

```{r, echo=FALSE, warning=FALSE}
# Residuals vs. Fitted Plots
op <- par(mfrow = c(2,2), mar = c(2,2,4,1))

resid_a <- resid(mod_DS_growth_5a)
fitted_a <- fitted(mod_DS_growth_5a)
plot(fitted_a, resid_a, main = "A: Hedges_d ~ mg_cm2, control = addSlist,
     random =  ~ mg_cm2 | Comparison")
abline(0,0)
#mod_DS_growth_1a and mod_DS_growth_3a were within 2 AIC

resid_b <- resid(mod_DS_growth_5b)
fitted_b <- fitted(mod_DS_growth_5b)
plot(fitted_b, resid_b, main = "B: smd ~ mg_cm2, control = addSlist,
     random =  ~ mg_cm2 | Comparison")
abline(0,0)
#mod_DS_growth_6b within 1 AIC

resid_c <- resid(mod_DS_growth_5c)
fitted_c <- fitted(mod_DS_growth_5c)
plot(fitted_c, resid_c, main = "C: Hedges_d ~ mg_cm2_d, control = addSlist,
     random =  ~ mg_cm2_d | Comparison")
abline(0,0)
#mod_DS_growth_2c within 0.3 AIC

resid_d <- resid(mod_DS_growth_5d)
fitted_d <- fitted(mod_DS_growth_5d)
plot(fitted_d, resid_d, main = "D: smd ~ mg_cm2_d, control = addSlist,
     random =  ~ mg_cm2_d | Comparison")
abline(0,0)

par(op)


# Normal Q-Q Plot
op <- par(mfrow = c(2,2), mar = c(2,2,4,1))
qqnorm(resid_a)
qqline(resid_a)
qqnorm(resid_b)
qqline(resid_b)
qqnorm(resid_c)
qqline(resid_c)
qqnorm(resid_d)
qqline(resid_d)
par(op)


# Patterns:
# 1.  Residuals for all 4 best-fit models are very similar, with moderate magnitudes (<abs(2))
# 2.  Models A and B have one value that is outside others, but its residual value is fine -- this was probably a point with long exposure duration.
# 3.  No obvious need to try out log or non-linear models!
```
  
Now let's check out a log, base 10 transformation of sediment exposure to see if it improves the residuals of the models.

```{r, echo=FALSE, warning=FALSE}
#?mixmeta

# Organizational Note: 
#All models with 'a' or 'b' at the end of its name have cumulative exposure as its predictor. 
#All models with 'c' or 'd' at the end of its name have exposure concentration as its predictor. 
#All models with 'a' or 'c' at the end of its name have manually calculated Hedges' d as its response.
#All models with 'b' or 'd' at the end of its name have covar.smd calculated Hedges' d as its response.

log10_mg_cm2 <- log10(ES_DS_growth2$Sed_exposure_mg_cm2)
log10_mg_cm2_d <- log10(ES_DS_growth2$Sed_level_standardized)

# LINEAR FIXED AND RANDOM EFFECTS NOT ACCOUNTING FOR WITHIN-STUDY CORRELATIONS
log_mod_DS_growth_1a <- mixmeta(Hedges_d ~ log10_mg_cm2, S = Var_d,
                       random =  ~ log10_mg_cm2 | Comparison,
                       data = ES_DS_growth2, method = "ml")
log_mod_DS_growth_1b <- mixmeta(smd ~ log10_mg_cm2, S = vmd,
                      random =  ~ log10_mg_cm2 | Comparison,
                      data = ES_DS_growth2, method = "ml")
log_mod_DS_growth_1c <- mixmeta(Hedges_d ~ log10_mg_cm2_d, S = Var_d,
                       random =  ~ log10_mg_cm2_d | Comparison,
                       data = ES_DS_growth2, method = "ml")
log_mod_DS_growth_1d <- mixmeta(smd ~ log10_mg_cm2_d, S = vmd,
                      random =  ~ log10_mg_cm2_d | Comparison,
                      data = ES_DS_growth2, method = "ml")

# LINEAR FIXED AND RANDOM, NESTED EFFECTS NOT ACCOUNTING FOR WITHIN-STUDY CORRELATIONS
log_mod_DS_growth_2a <- mixmeta(Hedges_d ~ log10_mg_cm2, S = Var_d,
                       random =  ~ log10_mg_cm2 | Ref/Comparison,
                       data = ES_DS_growth2, method = "ml")
log_mod_DS_growth_2b <- mixmeta(smd ~ log10_mg_cm2, S = vmd,
                      random =  ~ log10_mg_cm2 | Ref/Comparison,
                      data = ES_DS_growth2, method = "ml")
log_mod_DS_growth_2c <- mixmeta(Hedges_d ~ log10_mg_cm2_d, S = Var_d,
                       random =  ~ log10_mg_cm2_d | Ref/Comparison,
                       data = ES_DS_growth2, method = "ml")
log_mod_DS_growth_2d <- mixmeta(smd ~ log10_mg_cm2_d, S = vmd,
                      random =  ~ log10_mg_cm2_d | Ref/Comparison,
                      data = ES_DS_growth2, method = "ml")
```

To account for within-comparison correlations in the mixmeta model, I replace the 'S' argument with a 'control' argument, which requires me to provide the covariance matrices, as done with the covar.smd function. 

```{r, echo=FALSE, warning=FALSE}
# COMPUTE THE WITHIN-COMPARISON CORRELATIONS 
# already done above but code repeated here...
#covar_DS_growth <- by(ES_DS_growth, ES_DS_growth$Comparison, function(x) 
#  covar.smd(Tx_mean, Tx_sd, Tx_n, "smd", method="hedges", data = x))

# LINEAR FIXED AND RANDOM EFFECTS ACCOUNTING FOR WITHIN-COMPARISON CORRELATIONS
#Code for covariance matrix to include for addSlist for Comparison random effect
newlist_GRO_log <- list(NA)
for (i in seq(1,length(covar_DS_growth))) {
newlist_GRO_log[i] <- list(covar_DS_growth[[i]]$S)
}
log_mod_DS_growth_5a <- mixmeta(Hedges_d ~ log10_mg_cm2,
                       random = ~ log10_mg_cm2 | Comparison,
                       data=ES_DS_growth2, method="ml", #using dataset without controls
                       control=list(addSlist=newlist_GRO_log))
log_mod_DS_growth_5b <- mixmeta(smd ~ log10_mg_cm2,
                       random = ~ log10_mg_cm2 | Comparison,
                       data=ES_DS_growth2, method="ml", #using dataset without controls
                       control=list(addSlist=newlist_GRO_log))
log_mod_DS_growth_5c <- mixmeta(Hedges_d ~ log10_mg_cm2_d,
                       random = ~ log10_mg_cm2_d | Comparison,
                       data=ES_DS_growth2, method="ml", #using dataset without controls
                       control=list(addSlist=newlist_GRO_log))
log_mod_DS_growth_5d <- mixmeta(smd ~ log10_mg_cm2_d,
                       random = ~ log10_mg_cm2_d | Comparison,
                       data=ES_DS_growth2, method="ml", #using dataset without controls
                       control=list(addSlist=newlist_GRO_log))

# LINEAR FIXED AND RANDOM, NESTED EFFECTS ACCOUNTING FOR WITHIN-COMPARISON & WITHIN-STUDY CORRELATIONS
#Code for covariance matrix to include for addSlist for Ref/Comparison
newlist_GRO_log2 <- list(NA) #collects the list of covariance matrices for the block diag matrix by reference for the nested hierarchial ref/comparison
templist_GRO_log <- list(NA) #holds temp list of covariance matrices associated with each reference
templist_GRO_log2 <-list(NA)
reflista <- ES_DS_growth %>% distinct(Ref,Ref_name,Comparison)
reflist <- reflista[,1]
for (i in seq(1,length(unique(reflista$Ref))))  {
  #pull the elements from covar_DS_growth_b that are all from the same reference [i]
  templist_GRO_log[i] <-list(covar_DS_growth[reflist==unique(reflista$Ref)[i]])
  for (j in seq(1,length(templist_GRO_log[[i]]))) {
  #for each comparison in the reference, pull out the covar matrices (element $S) and put in into templist_GRO_log2
     templist_GRO_log2[j] <- list(templist_GRO_log[[i]][[j]]$S)
  }
  #turn list of covars from all comparison in one reference into block diag matrix
  newlist_GRO_log2[i] <- list(bdiagMat(templist_GRO_log2))
  templist_GRO_log2 <- list(NA)
}
#Error in getSlist(S, nay, groups, m, k, control$addSlist, control$checkPD) : wrong dimensions in 'addSlist'. See help(mixmetaControl)
#log_mod_DS_growth_6a <- mixmeta(Hedges_d ~ log10_mg_cm2,
#                       random = ~ log10_mg_cm2 | Ref/Comparison,
#                       data=ES_DS_growth2, method="ml", #using dataset without controls
#                       control=list(addSlist=newlist_GRO_log2))
#log_mod_DS_growth_6b <- mixmeta(smd ~ log10_mg_cm2,
#                       random = ~ log10_mg_cm2 | Ref/Comparison,
#                       data=ES_DS_growth2, method="ml", #using dataset without controls
#                       control=list(addSlist=newlist_GRO_log2))
#log_mod_DS_growth_6c <- mixmeta(Hedges_d ~ log10_mg_cm2_d,
#                       random = ~ log10_mg_cm2_d | Ref/Comparison,
#                       data=ES_DS_growth2, method="ml", #using dataset without controls
#                       control=list(addSlist=newlist_GRO_log2))
#log_mod_DS_growth_6d <- mixmeta(smd ~ log10_mg_cm2_d,
#                       random = ~ log10_mg_cm2_d | Ref/Comparison,
#                       data=ES_DS_growth2, method="ml", #using dataset without controls
#                       control=list(addSlist=newlist_GRO_log2))
```
    
### 3. Model Comparison and Summary
And now to compare relative fit of logarithmic models...
```{r, echo=FALSE, warning=FALSE}
# Model Comparisons
AIC(mod_DS_growth_5a, log_mod_DS_growth_1a, log_mod_DS_growth_2a, log_mod_DS_growth_5a) 
#of these, model non-log 5a has lowest AIC (but 1a, 5a within 2 AIC)

AIC(mod_DS_growth_5b, log_mod_DS_growth_1b, log_mod_DS_growth_2b, log_mod_DS_growth_5b) 
#of these, model 5b has lowest AIC (but 1a, 5a non-log within 2 AIC)

AIC(mod_DS_growth_5c, log_mod_DS_growth_1c, log_mod_DS_growth_2c, log_mod_DS_growth_5c) 
#of these, model 5c has lowest AIC (but 1c, 5a non-log within 2 AIC)

AIC(mod_DS_growth_5d, log_mod_DS_growth_1d, log_mod_DS_growth_2d, log_mod_DS_growth_5d) 
#of these, model non-log 5d has lowest AIC (but 5a log within 1 AIC)
```
  
And now to evaluate residuals of models...
```{r, echo=FALSE, warning=FALSE}
# Residuals vs. Fitted Plots
op <- par(mfrow = c(2,2), mar = c(2,2,4,1))

resid_loga <- resid(log_mod_DS_growth_5a)
fitted_loga <- fitted(log_mod_DS_growth_5a)
plot(fitted_loga, resid_loga, main = "A: Hedges_d ~ log_mg_cm2, control=addSlist,
     random =  ~ log_mg_cm2 | Comparison")
abline(0,0)
#log_mod_DS_growth_1a and log_mod_DS_growth_3a were within 2 AIC

resid_logb <- resid(log_mod_DS_growth_5b)
fitted_logb <- fitted(log_mod_DS_growth_5b)
plot(fitted_logb, resid_logb, main = "B: smd ~ log_mg_cm2, control=addSlist,
     random =  ~ log_mg_cm2 | Comparison")
abline(0,0)
#log_mod_DS_growth_6b within 1 AIC

resid_logc <- resid(log_mod_DS_growth_5c)
fitted_logc <- fitted(log_mod_DS_growth_5c)
plot(fitted_logc, resid_logc, main = "C: Hedges_d ~ log_mg_cm2_d, Scontrol=addSlist,
     random =  ~ log_mg_cm2_d | Comparison")
abline(0,0)
#log_mod_DS_growth_2c within 0.3 AIC

resid_logd <- resid(log_mod_DS_growth_5d)
fitted_logd <- fitted(log_mod_DS_growth_5d)
plot(fitted_logd, resid_logd, main = "D: smd ~ log_mg_cm2_d, control=addSlist,
     random =  ~ log_mg_cm2_d | Comparison")
abline(0,0)

par(op)


# Normal Q-Q Plot
op <- par(mfrow = c(2,2), mar = c(2,2,4,1))
qqnorm(resid_loga)
qqline(resid_loga)
qqnorm(resid_logb)
qqline(resid_logb)
qqnorm(resid_logc)
qqline(resid_logc)
qqnorm(resid_logd)
qqline(resid_logd)
par(op)


# Patterns:
# 1.  Residuals for all 4 best-fit models are very similar and look great -- even spread along fitted values axis and low magnitude residuals.
# 2.  No need to look at non-linear options
```
  
Residuals for all models are very similar and look great -- even spread along fitted values axis and low magnitude residuals. However, C and D not much improved, so I will defer to linear models for these.

### 4. Best-Fit Model Summary
  
#### i. MG/CM^2 IS PREDICTOR, Model using manual calculation of Hedges' *d*
Now I will predict the effect size along the exposure range and plot with confidence intervals.  

```{r, echo=FALSE, warning=FALSE}
summary(log_mod_DS_growth_5a)

# PREDICT THE EFFECT SIZE FOR 13.2 mg/cm2 FROM TWO MODELS (LOAEL identified by binary analyses)
#predict(log_mod_DS_growth_5a, newdata=data.frame(Sed_exposure_mg_cm2=13.2), ci=TRUE)

# PREDICT THE EFFECT SIZE AND PLOT WITH CONFIDENCE INTERVALS
pred_log_mod_DS_growth_5a <- predict(log_mod_DS_growth_5a, newdata=ES_DS_growth2, ci=TRUE)
ES_DS_growth2_CI <- cbind(ES_DS_growth2, pred_log_mod_DS_growth_5a)
#head(ES_DS_growth2_CI)
min_exp_a <- ES_DS_growth2_CI %>% 
  mutate(overlap0 = 0 >= ci.lb & 0 <= ci.ub) %>% 
  filter(overlap0==FALSE) %>% 
  summarize(min_exp_a=min(Sed_exposure_mg_cm2))
min_exp_a2 <- as.numeric(min_exp_a)

ggplot(ES_DS_growth2_CI, 
       aes(x = Sed_exposure_mg_cm2,
           y = Hedges_d,
           color = Ref,
           ymin = Hedges_d-Var_d,
           ymax = Hedges_d+Var_d)) + 
  geom_pointrange() +
  labs(x = expression("Sediment cumulative exposure (mg/cm"^"2"*")"),
       y = expression("Effect size (Hedges'"~italic(d)~"+/- variance of"~italic(d)~")"),
       color = "Study") +
  geom_abline(intercept=0, slope=0) +
  geom_vline(xintercept=min_exp_a2, linetype="dashed", color = "red") +
  geom_ribbon(aes(x = Sed_exposure_mg_cm2, y = Hedges_d,
                  ymin = ci.lb, ymax = ci.ub), 
              fill = "grey70", alpha = .15, 
              show.legend=FALSE, inherit.aes=FALSE) +
  geom_line(aes(x = Sed_exposure_mg_cm2, y = fit), inherit.aes=FALSE) +
  theme_classic() +
  scale_x_log10(limits = c(10,40000), breaks=c(10,100,1000,10000,min_exp_a2), 
                label=c("10","100","1000","10000",round(min_exp_a2,digits=1)))
```
  
#### ii. MG/CM^2 IS PREDICTOR, Model using dosresmeta calculation of Hedges' *d*
Now I will predict the effect size along the exposure range and plot with confidence intervals.  

```{r, echo=FALSE, warning=FALSE}
summary(log_mod_DS_growth_5b)

# PREDICT THE EFFECT SIZE FOR 13.2 mg/cm2 FROM TWO MODELS (LOAEL identified by binary analyses)
#predict(log_mod_DS_growth_5b, newdata=data.frame(Sed_exposure_mg_cm2=13.2), ci=TRUE)

# PREDICT THE EFFECT SIZE AND PLOT WITH CONFIDENCE INTERVALS
pred_log_mod_DS_growth_5b <- predict(log_mod_DS_growth_5b, newdata=ES_DS_growth2, ci=TRUE)
ES_DS_growth2_CI2 <- cbind(ES_DS_growth2, pred_log_mod_DS_growth_5b)
#head(ES_DS_growth2_CI2)
min_exp_b <- ES_DS_growth2_CI2 %>% 
  mutate(overlap0 = 0 >= ci.lb & 0 <= ci.ub) %>% 
  filter(overlap0==FALSE) %>% 
  summarize(min_exp_b=min(Sed_exposure_mg_cm2))
min_exp_b2 <- as.numeric(min_exp_b)

ggplot(ES_DS_growth2_CI2, 
       aes(x = Sed_exposure_mg_cm2,
           y = smd,
           color = Ref,
           ymin = smd-vmd,
           ymax = smd+vmd)) + 
  geom_pointrange() +
  labs(x = expression("Sediment cumulative exposure (mg/cm"^"2"*")"),
       y = expression("Effect size (Hedges'"~italic(d)~"+/- variance of"~italic(d)~")"),
       color = "Study") +
  geom_abline(intercept=0, slope=0) +
  geom_vline(xintercept=min_exp_b2, linetype="dashed", color = "red") +
  geom_ribbon(aes(x = Sed_exposure_mg_cm2, y = smd,
                  ymin = ci.lb, ymax = ci.ub), 
              fill = "grey70", alpha = .15, 
              show.legend=FALSE, inherit.aes=FALSE) +
  geom_line(aes(x = Sed_exposure_mg_cm2, y = fit), inherit.aes=FALSE) +
  theme_classic() +
  scale_x_log10(limits = c(10,40000), breaks=c(10,100,1000,10000,min_exp_b2), 
                label=c("10","100","1000","10000",round(min_exp_b2,digits=1)))
```
  
#### iii. MG/CM^2/DAY IS PREDICTOR, Model using manual calculation of Hedges' *d*
Now I will predict the effect size along the exposure range and plot with confidence intervals.  

```{r, echo=FALSE, warning=FALSE}
summary(mod_DS_growth_5c)

# PREDICT THE EFFECT SIZE FOR 13.2 mg/cm2 FROM TWO MODELS (LOAEL identified by binary analyses)
#predict(mod_DS_growth_5c, newdata=data.frame(Sed_exposure_mg_cm2=13.2), ci=TRUE)

# PREDICT THE EFFECT SIZE AND PLOT WITH CONFIDENCE INTERVALS
pred_mod_DS_growth_5c <- predict(mod_DS_growth_5c, newdata=ES_DS_growth2, ci=TRUE)
ES_DS_growth2_CI3 <- cbind(ES_DS_growth2, pred_mod_DS_growth_5c)
#head(ES_DS_growth2_CI)
min_conc_c <- ES_DS_growth2_CI3 %>% 
  mutate(overlap0 = 0 >= ci.lb & 0 <= ci.ub) %>% 
  filter(overlap0==FALSE) %>% 
  summarize(min_conc_c=min(Sed_level_standardized))
min_conc_c2 <- as.numeric(min_conc_c)

ggplot(ES_DS_growth2_CI3, 
       aes(x = Sed_level_standardized,
           y = Hedges_d,
           color = Ref,
           ymin = Hedges_d-Var_d,
           ymax = Hedges_d+Var_d)) + 
  geom_pointrange() +
  labs(x = expression("Sediment exposure concentration (mg/cm"^"2"*"/day)"),
       y = expression("Effect size (Hedges'"~italic(d)~"+/- variance of"~italic(d)~")"),
       color = "Study") +
  geom_abline(intercept=0, slope=0) +
  geom_vline(xintercept=min_conc_c2, linetype="dashed", color = "red") +
  geom_ribbon(aes(x = Sed_level_standardized, y = Hedges_d,
                  ymin = ci.lb, ymax = ci.ub), 
              fill = "grey70", alpha = .15, 
              show.legend=FALSE, inherit.aes=FALSE) +
  geom_line(aes(x = Sed_level_standardized, y = fit), inherit.aes=FALSE) +
  theme_classic()
```
  
#### iv. MG/CM^2/DAY IS PREDICTOR, Model using dosresmeta calculation of Hedges' *d*
Now I will predict the effect size along the exposure range and plot with confidence intervals.  

```{r, echo=FALSE, warning=FALSE}
summary(mod_DS_growth_5d)

# PREDICT THE EFFECT SIZE FOR 13.2 mg/cm2 FROM TWO MODELS (LOAEL identified by binary analyses)
#predict(mod_DS_growth_5d, newdata=data.frame(Sed_exposure_mg_cm2=13.2), ci=TRUE)

# PREDICT THE EFFECT SIZE AND PLOT WITH CONFIDENCE INTERVALS
pred_mod_DS_growth_5d <- predict(mod_DS_growth_5d, newdata=ES_DS_growth2, ci=TRUE)
ES_DS_growth2_CI4 <- cbind(ES_DS_growth2, pred_mod_DS_growth_5d)
#head(ES_DS_growth2_CI2)
min_conc_d <- ES_DS_growth2_CI4 %>% 
  mutate(overlap0 = 0 >= ci.lb & 0 <= ci.ub) %>% 
  filter(overlap0==FALSE) %>% 
  summarize(min_conc_d=min(Sed_level_standardized))
min_conc_d2 <- as.numeric(min_conc_d)

ggplot(ES_DS_growth2_CI2, 
       aes(x = Sed_level_standardized,
           y = smd,
           color = Ref,
           ymin = smd-vmd,
           ymax = smd+vmd)) + 
  geom_pointrange() +
  labs(x = expression("Sediment exposure concentration (mg/cm"^"2"*"/day)"),
       y = expression("Effect size (Hedges'"~italic(d)~"+/- variance of"~italic(d)~")"),
       color = "Study") +
  geom_abline(intercept=0, slope=0) +
  geom_vline(xintercept=min_conc_d2, linetype="dashed", color = "red") +
  geom_ribbon(aes(x = Sed_level_standardized, y = smd,
                  ymin = ci.lb, ymax = ci.ub), 
              fill = "grey70", alpha = .15, 
              show.legend=FALSE, inherit.aes=FALSE) +
  geom_line(aes(x = Sed_level_standardized, y = fit), inherit.aes=FALSE) +
  theme_classic()
```