---
title: "Effects of sediment on the survival of larval corals"
author: "Lillian J. Tuttle"
date: "9/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(mixmeta)
library(dosresmeta)
library(splines)

effectsize <- read.csv("data/effectsize_ALL.csv", header=T)
```
  
```{r, echo=FALSE, warning=FALSE}
#Looking at data
#head(effectsize) 

#a lot of variables (columns) are factors instead of numeric/doubles/integers
effectsize <- effectsize %>% #converting appropriate columns to numeric...
  mutate_at(vars(ES, Hedges_d, Var_d, SE_d, CI_lo_d, CI_up_d, Sed_level_standardized, Sed_exposure, Sed_exposure_mg_d_L), as.character) %>% 
  mutate_at(vars(ES, Hedges_d, Var_d, SE_d, CI_lo_d, CI_up_d, Sed_level_standardized, Sed_exposure, Sed_exposure_mg_d_L), as.numeric)
#head(effectsize) #"double"-check!

#Now create a new variable that has both Genus and species, remove sediment burial experiments, and remove irrelevant rows (non-controls, -treatments, or -baselines)
effectsize <- effectsize %>%
  mutate(Gsp = paste(Updated_Genus, Updated_species, sep = "_")) %>% 
  filter(Binary_sed_burial=="0") %>%
  filter(!Control=="-1")
#head(effectsize)

#Separate into deposited sediment (DS) and suspended sediment (SS) datasets, WITH CONTROLS
ES_DS <- effectsize %>% filter(DS_SS=="DS")
ES_SS <- effectsize %>% filter(DS_SS=="SS")

#Second dataset that that removes controls (Control == c("1","2"))
effectsize2 <- effectsize %>% filter(Control=="0")
#effectsize2 #"double"-check!

#Separate into deposited sediment (DS) and suspended sediment (SS) datasets, WITHOUT CONTROLS
ES_DS2 <- effectsize2 %>% filter(DS_SS=="DS")
ES_SS2 <- effectsize2 %>% filter(DS_SS=="SS")

#Now separate into independent dataset for each coral response for SUSPENDED SEDIMENT
ES_SS_larvalsurvival <- ES_SS %>% filter(Response == "larvalsurvival")
#ES_SS_larvalsurvival %>% tally()
#ES_SS_larvalsurvival %>% count(Ref) #n=4 studies from 3 papers
#ES_SS_larvalsurvival %>% count(Ref_name)
#ES_SS_larvalsurvival %>% count(Gsp)
#ES_SS_larvalsurvival %>% count(Updated_Genus)
``` 
  
These datasets contain information from multiple articles (**Ref_name**) and studies/experiments (**Ref**) within those articles about the effects of SUSPENDED sediment on the **larval survival rate** of pre-settlement corals. I would like to explore the dose-response relationship between sediment 'dose' and the effect size, estimated here as the standardized mean difference between treatment and control, in terms of Hedges' *d*:

- Hedges' *d* is the standardized mean difference between the treatment and control groups (Hedges and Olkin 1985). Each study may have multiple Hedges' *d* calculations, one for each treatment-control comparison at each time-point.   

- 'Dose' can be quantified as sediment exposure concentration (mg/cm^2^/day), exposure duration (days), or as "cumulative exposure", for which I multiply concentration and duration (mg/cm^2^).

Each study/experiment (Ref) may contain multiple sets of controls and treatments (due to using multiple species or sediment types, for instance). Thus, I have added two categorical vectors, '**Comparison**' and 'Control', which group each set of control-treatment comparisons into a unique two-letter code, within which is one control (Control=="1") and at least one treatment (Control=="0"). **Therefore, 'Comparison' is nested within 'Ref', which is further nested within 'Ref_name'.**

```{r, echo=FALSE}
#Using dosresmeta to calculate Hedges' d and variance of d
covar_SS_larvalsurvival <- by(ES_SS_larvalsurvival, ES_SS_larvalsurvival$Comparison, function(x) 
  covar.smd(Tx_mean, Tx_sd, Tx_n, "smd", method="hedges", data = x))
ES_SS_larvalsurvival$smd <- unlist(lapply(covar_SS_larvalsurvival, function(x) x$y))
ES_SS_larvalsurvival$vmd <- unlist(lapply(covar_SS_larvalsurvival, function(x) x$v))

#Looking at data to make sure smd and vmd added as columns
#ES_SS_larvalsurvival

#Removing controls for modeling
ES_SS_larvalsurvival2 <- subset(ES_SS_larvalsurvival, Control=="0")
#ES_SS_larvalsurvival2 %>% tally()

#Looking at data to make sure controls were removed
#ES_SS_larvalsurvival2
``` 
  
# I. Larval Survival Rate  

## A. Suspended Sediment  
  
### 1. Plots  
I calculated the standardized mean difference in terms of Hedges' *d* and plot it here with respect to exposure concentration. I did this first for a Hedges' *d* that I calculated before importing the data to R and again for a Hedges' *d* that is calculated by the dosresmeta::covar.smd function.

```{r, echo=FALSE, warning=FALSE}
#EXPOSURE CONCENTRATION vs. effect size, by study
#Using previously calculated Hedges' d and variance of d
ggplot(ES_SS_larvalsurvival2, 
       aes(x = Sed_level_standardized,
           y = Hedges_d,
           color = Ref,
           ymin = Hedges_d-Var_d,
           ymax = Hedges_d+Var_d)) + 
  geom_pointrange() +
  ggtitle("Larval survival rate") +
  labs(x = "Sediment exposure concentration (mg/L)",
       y = "Effect size (Hedges' d +/- s.d., manual calculation)",
       color = "Study") +
  geom_abline(intercept=0, slope=0) +
  theme_classic() +
  scale_x_log10(limits = c(1,1000), breaks=c(0.1,1,10,100,1000), 
                label=c("0.1","1","10","100","1000"))
#Using dosresmeta's calculation of Hedges' d and variance of d
ggplot(ES_SS_larvalsurvival2, 
       aes(x = Sed_level_standardized,
           y = smd,
           color = Ref,
           ymin = smd-vmd,
           ymax = smd+vmd)) + 
  geom_pointrange() +
  ggtitle("Larval survival rate") +
  labs(x = "Sediment exposure concentration (mg/L)",
       y = "Effect size (Hedges' d +/- s.d., dosresmeta calculation)",
       color = "Study") +
  geom_abline(intercept=0, slope=0) +
  theme_classic() +
  scale_x_log10(limits = c(1,1000), breaks=c(1,10,100,1000), 
                label=c("1","10","100","1000")) 


#by SEDIMENT TYPE
#Using previously calculated Hedges' d and variance of d
ggplot(ES_SS_larvalsurvival2, 
       aes(x = Sed_level_standardized,
           y = smd,
           color = Sed_grain_category,
           shape = Sed_source,
           ymin = smd-vmd,
           ymax = smd+vmd)) + 
  geom_pointrange() +
  ggtitle("Larval survival rate") +
  labs(x = "Sediment exposure concentration (mg/L)",
       y = "Effect size (Hedges' d +/- s.d., dosresmeta calculation)",
       color = "Sediment relative\ngrain size",
       shape = "Sediment source") +
  geom_abline(intercept=0, slope=0) +
  theme_classic() +
  scale_x_log10(limits = c(1,1000), breaks=c(1,10,100,1000), 
                label=c("1","10","100","1000")) 
ggplot(ES_SS_larvalsurvival2, 
       aes(x = Sed_level_standardized,
           y = smd,
           color = Sed_grain_class,
           shape = Sed_mineralogy,
           ymin = smd-vmd,
           ymax = smd+vmd)) + 
  geom_pointrange() +
  ggtitle("Larval survival rate") +
  labs(x = "Sediment exposure concentration (mg/L)",
       y = "Effect size (Hedges' d +/- s.d., dosresmeta calculation)",
       color = "Sediment\ngrain class",
       shape = "Sediment mineralogy") +
  geom_abline(intercept=0, slope=0) +
  theme_classic() +
  scale_x_log10(limits = c(1,1000), breaks=c(1,10,100,1000), 
                label=c("1","10","100","1000")) 

#by PHYLOGENY
#Using previously calculated Hedges' d and variance of d
ggplot(ES_SS_larvalsurvival2, 
       aes(x = Sed_level_standardized,
           y = smd,
           color = Gsp,
           ymin = smd-vmd,
           ymax = smd+vmd)) + 
  geom_pointrange() +
  ggtitle("Larval survival rate") +
  labs(x = "Sediment exposure concentration (mg/L)",
       y = "Effect size (Hedges' d +/- s.d., dosresmeta calculation)",
       color = "Species") +
  geom_abline(intercept=0, slope=0) +
  theme_classic() +
  scale_x_log10(limits = c(1,1000), breaks=c(1,10,100,1000), 
                label=c("1","10","100","1000")) 

#by OCEAN and REGION
#Using previously calculated Hedges' d and variance of d
ggplot(ES_SS_larvalsurvival2, 
       aes(x = Sed_level_standardized,
           y = smd,
           color = Region,
           shape = Ocean,
           ymin = smd-vmd,
           ymax = smd+vmd)) + 
  geom_pointrange() +
  ggtitle("Larval survival rate") +
  labs(x = "Sediment exposure concentration (mg/L)",
       y = "Effect size (Hedges' d +/- s.d., dosresmeta calculation)",
       color = "Region",
       shape = "Ocean") +
  geom_abline(intercept=0, slope=0) +
  theme_classic() +
  scale_x_log10(limits = c(1,1000), breaks=c(1,10,100,1000), 
                label=c("1","10","100","1000")) 
```
  
It is interesting that there is a difference between how I calculated Hedges' *d* and how *dosresmeta* appears to calculate it. But in both cases, there does not appear to be an effect of sediment exposure on effect size, except for *Acropora digitifera*, for which there may a strong effect of sediment on effect size (though there are limited datapoints for this species). There is not a discernible effect of geography (ocean and region) because almost all of the data were from the Great Barrier Reef. I still need to investigate the effects of sediment type (source, origin, grain size).
  
### 2. Model Fitting  
The following set of functions model fixed and random (nested) effects WITHOUT accounting for within-comparison (or within-study) correlations. The output for each is a univariate, random-effects meta-regression.

```{r, echo=FALSE, warning=FALSE}
#?mixmeta

# Organizational Note: 
#All models have exposure concentration as its predictor. 
#All models with 'd' at the end of its name have covar.smd calculated Hedges' d as its response.

# LINEAR FIXED AND RANDOM EFFECTS NOT ACCOUNTING FOR WITHIN-STUDY CORRELATIONS
mod_SS_larvalsurv_1d <- mixmeta(smd ~ Sed_level_standardized, S = vmd,
                      random =  ~ Sed_level_standardized | Comparison,
                      data = ES_SS_larvalsurvival2, method = "ml")

# LINEAR FIXED AND RANDOM, NESTED EFFECTS NOT ACCOUNTING FOR WITHIN-STUDY CORRELATIONS
mod_SS_larvalsurv_2d <- mixmeta(smd ~ Sed_level_standardized, S = vmd,
                      random =  ~ Sed_level_standardized | Ref/Comparison,
                      data = ES_SS_larvalsurvival2, method = "ml")

# LINEAR FIXED AND RANDOM, NESTED EFFECTS NOT ACCOUNTING FOR WITHIN-STUDY CORRELATIONS
mod_SS_larvalsurv_3d <- mixmeta(smd ~ Sed_level_standardized, S = vmd,
                      random =  ~ Sed_level_standardized | Ref_name/Comparison,
                      data = ES_SS_larvalsurvival2, method = "ml")

# LINEAR FIXED AND RANDOM, NESTED EFFECTS NOT ACCOUNTING FOR WITHIN-STUDY CORRELATIONS
mod_SS_larvalsurv_4d <- mixmeta(smd ~ Sed_level_standardized, S = vmd,
                       random =  ~ Sed_level_standardized | Ref_name/Ref/Comparison,
                       data = ES_SS_larvalsurvival2, method = "ml")
```

To account for within-comparison correlations in the mixmeta model, I replace the 'S' argument with a 'control' argument, which requires me to provide the covariance matrices, as done with the covar.smd function. 

```{r, echo=FALSE, warning=FALSE}
# COMPUTE THE WITHIN-COMPARISON CORRELATIONS 
# already done above but code repeated here...
#covar_SS_larvalsurvival <- by(ES_SS_larvalsurvival, ES_SS_larvalsurvival$Comparison, function(x) 
#  covar.smd(Tx_mean, Tx_sd, Tx_n, "smd", method="hedges", data = x))


# LINEAR FIXED AND RANDOM EFFECTS ACCOUNTING FOR WITHIN-COMPARISON CORRELATIONS
#Code for covariance matrix to include for addSlist for Comparison random effect
newlist_LSURV <- list(NA)
for (i in seq(1,length(covar_SS_larvalsurvival))) {
newlist_LSURV[i] <- list(covar_SS_larvalsurvival[[i]]$S)
}
mod_SS_larvalsurv_5d <- mixmeta(smd ~ Sed_level_standardized,
                       random = ~ Sed_level_standardized | Comparison,
                       data=ES_SS_larvalsurvival2, method="ml", #dataset without controls
                       control=list(addSlist=newlist_LSURV))


# LINEAR FIXED AND RANDOM, NESTED EFFECTS ACCOUNTING FOR WITHIN-COMPARISON & WITHIN-STUDY CORRELATIONS
#Code for covariance matrix to include for addSlist for Ref/Comparison
newlist_LSURV2 <- list(NA) #collects the list of covariance matrices for the block diag matrix by reference for the nested hierarchial ref/comparison
templist_LSURV <- list(NA) #holds temp list of covariance matrices associated with each reference
templist_LSURV2 <-list(NA)
reflista <- ES_SS_larvalsurvival %>% distinct(Ref,Ref_name,Comparison)
reflist <- reflista[,1]
for (i in seq(1,length(unique(reflista$Ref))))  {
  #pull the elements from covar_SS_larvalsurvival_b that are all from the same reference [i]
  templist_LSURV[i] <-list(covar_SS_larvalsurvival[reflist==unique(reflista$Ref)[i]])
  for (j in seq(1,length(templist_LSURV[[i]]))) {
  #for each comparison in the reference, pull out the covar matrices (element $S) and put in into templist_LSURV2
     templist_LSURV2[j] <- list(templist_LSURV[[i]][[j]]$S)
  }
  #turn list of covars from all comparison in one reference into block diag matrix
  newlist_LSURV2[i] <- list(bdiagMat(templist_LSURV2))
  templist_LSURV2 <- list(NA)
}
mod_SS_larvalsurv_6d <- mixmeta(smd ~ Sed_level_standardized,
                       random = ~ Sed_level_standardized | Ref/Comparison,
                       data=ES_SS_larvalsurvival2, method="ml", #dataset without controls
                       control=list(addSlist=newlist_LSURV2))


# LINEAR FIXED AND RANDOM, NESTED EFFECTS ACCOUNTING FOR WITHIN-COMPARISON & WITHIN-STUDY CORRELATIONS
#Code for covariance matrix to include for addSlist for Ref_name/Comparison
newlist_LSURV3 <- list(NA) #collects the list of covariance matrices for the block diag matrix by reference for the nested hierarchial Ref_name/Comparison
templist_LSURV3 <- list(NA) #holds temp list of covariance matrices associated with each reference
templist_LSURV4 <-list(NA)
reflista <- ES_SS_larvalsurvival %>% distinct(Ref,Ref_name,Comparison)
reflist <- reflista[,2]
for (i in seq(1,length(unique(reflista$Ref_name))))  {
  #pull the elements from covar_SS_larvalsurvival_b that are all from the same reference [i]
  templist_LSURV3[i] <-list(covar_SS_larvalsurvival[reflist==unique(reflista$Ref_name)[i]])
  for (j in seq(1,length(templist_LSURV3[[i]]))) {
  #for each comparison in the reference, pull out the covar matrices (element $S) and put in into templist_LSURV4
     templist_LSURV4[j] <- list(templist_LSURV3[[i]][[j]]$S)
  }
  #turn list of covars from all comparison in one reference into block diag matrix
  newlist_LSURV3[i] <- list(bdiagMat(templist_LSURV4))
  templist_LSURV4 <- list(NA)
}
mod_SS_larvalsurv_7d <- mixmeta(smd ~ Sed_level_standardized,
                       random = ~ Sed_level_standardized | Ref_name/Comparison,
                       data=ES_SS_larvalsurvival2, method="ml", #dataset without controls
                       control=list(addSlist=newlist_LSURV3))
```
  
And now to compare relative fit of linear models...

```{r, echo=FALSE, warning=FALSE}
# Model Comparisons
AIC(mod_SS_larvalsurv_1d, mod_SS_larvalsurv_2d, mod_SS_larvalsurv_3d, mod_SS_larvalsurv_4d, mod_SS_larvalsurv_5d, mod_SS_larvalsurv_6d, mod_SS_larvalsurv_7d) 
#of these, model 1d has lowest AIC
```
  
And now to evaluate residuals of models...

```{r, echo=FALSE, warning=FALSE}
# Residuals vs. Fitted Plots and Normal Q-Q Plots
op <- par(mfrow = c(1,2), mar = c(2,2,4,1))

resid_d <- resid(mod_SS_larvalsurv_1d)
fitted_d <- fitted(mod_SS_larvalsurv_1d)
plot(fitted_d, resid_d, main = "D: smd ~ mg_L, S = vmd, 
     random =  ~ mg_L | Comparison")
abline(0,0)

qqnorm(resid_d)
qqline(resid_d)

par(op)
```
  
Patterns:
1.  Residuals have relatively low magnitude of residuals (<abs(6)).
2.  Slight outliers in negative residual land, but not extreme.
3.  Might benefit from a log transformation of sediment exposure to even the horizontal spread of fitted values.
  
Now let's check out a log, base 10 transformation of sediment exposure to see if it improves the residuals of the models.  

```{r, echo=FALSE, warning=FALSE}
#?mixmeta

# Organizational Note: 
#All models have exposure concentration as its predictor. 
#All models with 'd' at the end of its name have covar.smd calculated Hedges' d as its response.

log10_mg_L <- log10(ES_SS_larvalsurvival2$Sed_level_standardized)

# LINEAR FIXED AND RANDOM EFFECTS NOT ACCOUNTING FOR WITHIN-STUDY CORRELATIONS
log_mod_SS_larvalsurv_1d <- mixmeta(smd ~ log10_mg_L, S = vmd,
                      random =  ~ log10_mg_L | Comparison,
                      data = ES_SS_larvalsurvival2, method = "ml")

# LINEAR FIXED AND RANDOM, NESTED EFFECTS NOT ACCOUNTING FOR WITHIN-STUDY CORRELATIONS
log_mod_SS_larvalsurv_2d <- mixmeta(smd ~ log10_mg_L, S = vmd,
                      random =  ~ log10_mg_L | Ref/Comparison,
                      data = ES_SS_larvalsurvival2, method = "ml")

# LINEAR FIXED AND RANDOM, NESTED EFFECTS NOT ACCOUNTING FOR WITHIN-STUDY CORRELATIONS
log_mod_SS_larvalsurv_3d <- mixmeta(smd ~ log10_mg_L, S = vmd,
                      random =  ~ log10_mg_L | Ref_name/Comparison,
                      data = ES_SS_larvalsurvival2, method = "ml")

# LINEAR FIXED AND RANDOM, NESTED EFFECTS NOT ACCOUNTING FOR WITHIN-STUDY CORRELATIONS
log_mod_SS_larvalsurv_4d <- mixmeta(smd ~ log10_mg_L, S = vmd,
                       random =  ~ log10_mg_L | Ref_name/Ref/Comparison,
                       data = ES_SS_larvalsurvival2, method = "ml")
```

To account for within-comparison correlations in the mixmeta model, I replace the 'S' argument with a 'control' argument, which requires me to provide the covariance matrices, as done with the covar.smd function. 

```{r, echo=FALSE, warning=FALSE}
# COMPUTE THE WITHIN-COMPARISON CORRELATIONS 
# already done above but code repeated here...
#covar_SS_larvalsurvival <- by(ES_SS_larvalsurvival, ES_SS_larvalsurvival$Comparison, function(x) 
#  covar.smd(Tx_mean, Tx_sd, Tx_n, "smd", method="hedges", data = x))


# LINEAR FIXED AND RANDOM EFFECTS ACCOUNTING FOR WITHIN-COMPARISON CORRELATIONS
#Code for covariance matrix to include for addSlist for Comparison random effect
newlist_log_LSURV <- list(NA)
for (i in seq(1,length(covar_SS_larvalsurvival))) {
newlist_log_LSURV[i] <- list(covar_SS_larvalsurvival[[i]]$S)
}
log_mod_SS_larvalsurv_5d <- mixmeta(smd ~ log10_mg_L,
                       random = ~ log10_mg_L | Comparison,
                       data=ES_SS_larvalsurvival2, method="ml", #dataset without controls
                       control=list(addSlist=newlist_log_LSURV))


# LINEAR FIXED AND RANDOM, NESTED EFFECTS ACCOUNTING FOR WITHIN-COMPARISON & WITHIN-STUDY CORRELATIONS
#Code for covariance matrix to include for addSlist for Ref/Comparison
newlist_log_LSURV2 <- list(NA) #collects the list of covariance matrices for the block diag matrix by reference for the nested hierarchial ref/comparison
templist_log_LSURV <- list(NA) #holds temp list of covariance matrices associated with each reference
templist_log_LSURV2 <-list(NA)
reflista <- ES_SS_larvalsurvival %>% distinct(Ref,Ref_name,Comparison)
reflist <- reflista[,1]
for (i in seq(1,length(unique(reflista$Ref))))  {
  #pull the elements from covar_SS_larvalsurvival_b that are all from the same reference [i]
  templist_log_LSURV[i] <-list(covar_SS_larvalsurvival[reflist==unique(reflista$Ref)[i]])
  for (j in seq(1,length(templist_log_LSURV[[i]]))) {
  #for each comparison in the reference, pull out the covar matrices (element $S) and put in into templist_log_LSURV2
     templist_log_LSURV2[j] <- list(templist_log_LSURV[[i]][[j]]$S)
  }
  #turn list of covars from all comparison in one reference into block diag matrix
  newlist_log_LSURV2[i] <- list(bdiagMat(templist_log_LSURV2))
  templist_log_LSURV2 <- list(NA)
}
log_mod_SS_larvalsurv_6d <- mixmeta(smd ~ log10_mg_L,
                       random = ~ log10_mg_L | Ref/Comparison,
                       data=ES_SS_larvalsurvival2, method="ml", #dataset without controls
                       control=list(addSlist=newlist_log_LSURV2))


# LINEAR FIXED AND RANDOM, NESTED EFFECTS ACCOUNTING FOR WITHIN-COMPARISON & WITHIN-STUDY CORRELATIONS
#Code for covariance matrix to include for addSlist for Ref_name/Comparison
newlist_log_LSURV3 <- list(NA) #collects the list of covariance matrices for the block diag matrix by reference for the nested hierarchial Ref_name/Comparison
templist_log_LSURV3 <- list(NA) #holds temp list of covariance matrices associated with each reference
templist_log_LSURV4 <-list(NA)
reflista <- ES_SS_larvalsurvival %>% distinct(Ref,Ref_name,Comparison)
reflist <- reflista[,2]
for (i in seq(1,length(unique(reflista$Ref_name))))  {
  #pull the elements from covar_SS_larvalsurvival_b that are all from the same reference [i]
  templist_log_LSURV3[i] <-list(covar_SS_larvalsurvival[reflist==unique(reflista$Ref_name)[i]])
  for (j in seq(1,length(templist_log_LSURV3[[i]]))) {
  #for each comparison in the reference, pull out the covar matrices (element $S) and put in into templist_log_LSURV4
     templist_log_LSURV4[j] <- list(templist_log_LSURV3[[i]][[j]]$S)
  }
  #turn list of covars from all comparison in one reference into block diag matrix
  newlist_log_LSURV3[i] <- list(bdiagMat(templist_log_LSURV4))
  templist_log_LSURV4 <- list(NA)
}
log_mod_SS_larvalsurv_7d <- mixmeta(smd ~ log10_mg_L,
                       random = ~ log10_mg_L | Ref_name/Comparison,
                       data=ES_SS_larvalsurvival2, method="ml", #dataset without controls
                       control=list(addSlist=newlist_log_LSURV3))
```
  
And now to compare relative fit of linear models...

```{r, echo=FALSE, warning=FALSE}
# Model Comparisons
AIC(log_mod_SS_larvalsurv_1d, log_mod_SS_larvalsurv_2d, log_mod_SS_larvalsurv_3d, log_mod_SS_larvalsurv_4d, log_mod_SS_larvalsurv_5d, log_mod_SS_larvalsurv_6d, log_mod_SS_larvalsurv_7d) 
#of these, model 3d has lowest AIC (but 1d within 2 AIC)
```
  
And now to evaluate residuals of models...

```{r, echo=FALSE, warning=FALSE}
# Residuals vs. Fitted Plots and Normal Q-Q Plots
op <- par(mfrow = c(1,2), mar = c(2,2,4,1))

resid_logd <- resid(log_mod_SS_larvalsurv_3d)
fitted_logd <- fitted(log_mod_SS_larvalsurv_3d)
plot(fitted_logd, resid_logd, main = "D: smd ~ log_mg_L, S = vmd, 
     random =  ~ log_mg_L | Ref_name/Comparison")
abline(0,0)

qqnorm(resid_logd)
qqline(resid_logd)

par(op)
```
  
Patterns:
1.  Looks great -- still slight outliers in negative residual land, but not extreme, and now a more even horizontal spread of fitted values. No need for non-linear models.
  
### 3. Best-Fit Model Summary
  
#### MG/L IS PREDICTOR, Model using dosresmeta calculation of Hedges' *d*
Now I will predict the effect size along the exposure range and plot with confidence intervals.  

```{r, echo=FALSE, warning=FALSE}
summary(log_mod_SS_larvalsurv_3d)

# PREDICT THE EFFECT SIZE FOR 13.2 mg/cm2 FROM TWO MODELS (LOAEL identified by binary analyses)
#predict(log_mod_SS_larvalsurv_3d, newdata=data.frame(Sed_level_standardized=13.2), ci=TRUE)

# PREDICT THE EFFECT SIZE AND PLOT WITH CONFIDENCE INTERVALS
pred_log_mod_SS_larvalsurv_3d <- predict(log_mod_SS_larvalsurv_3d, newdata=ES_SS_larvalsurvival2, ci=TRUE)
ES_SS_larvalsurvival2_CI4 <- cbind(ES_SS_larvalsurvival2, pred_log_mod_SS_larvalsurv_3d)
#head(ES_SS_larvalsurvival2_CI4)
min_exp_d <- ES_SS_larvalsurvival2_CI4 %>% 
  mutate(overlap0 = 0 >= ci.lb & 0 <= ci.ub) %>% 
  filter(overlap0==FALSE) %>% 
  summarize(min_exp_d=min(Sed_level_standardized))
min_exp_d2 <- as.numeric(min_exp_d)

ggplot(ES_SS_larvalsurvival2_CI4, 
       aes(x = Sed_level_standardized,
           y = smd,
           color = Ref,
           ymin = smd-vmd,
           ymax = smd+vmd)) + 
  geom_pointrange() +
  labs(x = "Sediment exposure concentration (mg/L)",
       y = expression("Effect size (Hedges'"~italic(d)~"+/- variance of"~italic(d)~")"),
       color = "Study") +
  geom_abline(intercept=0, slope=0) +
  geom_vline(xintercept=min_exp_d2, linetype="dashed", color = "red") +
  geom_ribbon(aes(x = Sed_level_standardized, y = smd,
                  ymin = ci.lb, ymax = ci.ub), 
              fill = "grey70", alpha = .15, 
              show.legend=FALSE, inherit.aes=FALSE) +
  geom_line(aes(x = Sed_level_standardized, y = fit), inherit.aes=FALSE) +
  theme_classic() +
  scale_x_log10(limits = c(1,1000), breaks=c(1,10,100,1000,min_exp_d2), 
                label=c("1","10","100","1000",round(min_exp_d2,digits=1)))
```
  
There is no significant relationship between exposure concentration to suspended sediment and the Hedges' d effect size for larval survival (DRMA z = -1.4434, p = 0.1489).
