---
title: "Effects of sediment on the mortality of adult corals"
author: "Lillian J. Tuttle"
date: "9/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(mixmeta)
library(dosresmeta)
library(splines)

effectsize <- read.csv("data/effectsize_ALL.csv", header=T)
```
  
```{r, echo=FALSE, warning=FALSE}
#Looking at data
#totalmort

#a lot of variables (columns) are factors instead of numeric/doubles/integers
effectsize <- effectsize %>% #converting appropriate columns to numeric...
  mutate_at(vars(ES, Hedges_d, Var_d, SE_d, CI_lo_d, CI_up_d, Sed_level_standardized, Sed_exposure, Sed_exposure_mg_d_L), as.character) %>% 
  mutate_at(vars(ES, Hedges_d, Var_d, SE_d, CI_lo_d, CI_up_d, Sed_level_standardized, Sed_exposure, Sed_exposure_mg_d_L), as.numeric)
#head(effectsize) #"double"-check!

#Now create a new variable that has both Genus and species, remove sediment burial experiments, and remove irrelevant rows (non-controls, -treatments, or -baselines)
effectsize <- effectsize %>%
  mutate(Gsp = paste(Updated_Genus, Updated_species, sep = "_")) %>% 
  filter(Binary_sed_burial=="0") %>%
  filter(!Control=="-1")
#head(effectsize)

#Separate into deposited sediment (DS) and suspended sediment (SS) datasets, WITH CONTROLS
ES_DS <- effectsize %>% filter(DS_SS=="DS")
ES_SS <- effectsize %>% filter(DS_SS=="SS")

#Second dataset that that removes controls (Control == c("1","2"))
effectsize2 <- effectsize %>% filter(Control=="0")
#effectsize2 #"double"-check!

#Separate into deposited sediment (DS) and suspended sediment (SS) datasets, WITHOUT CONTROLS
ES_DS2 <- effectsize2 %>% filter(DS_SS=="DS")
ES_SS2 <- effectsize2 %>% filter(DS_SS=="SS")

#Now separate into independent datasets for each coral response for DEPOSITED SEDIMENT
ES_DS_totalmort <- ES_DS %>% filter(Response == "total_mortality")
#DS dataset without Hodgson 1989 (DS08a,b):
ES_DS_totalmort_b <- ES_DS_totalmort %>% 
  filter(!Ref=="DS08a") %>% filter(!Ref=="DS08b")

#Now separate into independent datasets for each coral response for SUSPENDED SEDIMENT
ES_SS_totalmort <- ES_SS %>% filter(Response == "total_mortality")

#ES_DS_totalmort_b %>% tally()
#ES_DS_totalmort_b %>% count(Ref)
#ES_DS_totalmort_b %>% count(Ref_name)
#ES_DS_totalmort_b %>% count(Gsp)
#ES_DS_totalmort_b %>% count(Updated_Genus)
#ES_SS_totalmort %>% tally()
#ES_SS_totalmort %>% count(Ref)
#ES_SS_totalmort %>% count(Ref_name)
#ES_SS_totalmort %>% count(Gsp)
#ES_SS_totalmort %>% count(Updated_Genus)
``` 
  
These datasets contain information from multiple articles (**Ref_name**) and studies/experiments (**Ref**) within those articles about the effects of DEPOSITED & SUSPENDED sediment on the **total colony mortality** of adult corals, quantified as the percent of complete colony mortality within an experimental group. I would like to explore the dose-response relationship between sediment 'dose' and the effect size, estimated here as the log risk/rate ratio (**logRR**) or the log odds ratio (**logOR**):

- The log risk ratio is defined as...

- The log odds ratio is defined as...

- 'Dose' can be quantified as sediment exposure concentration (mg/cm^2^/day for deposited sediment and mg/L for suspended sediment), exposure duration (days), or as "cumulative exposure", for which I multiply concentration and duration (mg/cm^2^ for deposited sediment and mg x day/L for suspended sediment).

Each study/experiment (Ref) may contain multiple sets of controls and treatments (due to using multiple species or sediment types, for instance). Thus, I have added two categorical vectors, '**Comparison**' and 'Control', which group each set of control-treatment comparisons into a unique two-letter code, within which is one control (Control=="1") and at least one treatment (Control=="0"). **Therefore, 'Comparison' is nested within 'Ref', which is further nested within 'Ref_name'.**

```{r, echo=FALSE}
#CALCULATE COVARIANCE
#Deposited sediment, without Hodgson 1989 Chapter IV
addS_ES_DS_totalmort_b <- lapply(split(ES_DS_totalmort_b, ES_DS_totalmort_b$Comparison), function(x)
  covar.logrr(y=logRR, v=logRR_var, cases=Tx_n_dead_adj, n=Tx_n_total_adj, type=Type, data=x))

#Suspended sediment
addS_ES_SS_totalmort <- lapply(split(ES_SS_totalmort, ES_SS_totalmort$Comparison), function(x)
  covar.logrr(y=logRR, v=logRR_var, cases=Tx_n_dead_adj, n=Tx_n_total_adj, type=Type, data=x))

#Removing controls for modeling
ES_DS_totalmort2 <- subset(ES_DS_totalmort, Control=="0")
ES_DS_totalmort_b2 <- subset(ES_DS_totalmort_b, Control=="0")
ES_SS_totalmort2 <- subset(ES_SS_totalmort, Control=="0")

#Looking at data to make sure controls were removed
#ES_DS_totalmort2
#ES_DS_totalmort_b2
#ES_SS_totalmort2
#ES_DS_totalmort_b2 %>% tally()
#ES_SS_totalmort2 %>% tally()
``` 
  
# I. Deposited Sediment  
  
## A. Plots  

```{r, echo=FALSE, warning=FALSE}

##WITH HODGSON 1989 PHD CHAPTER IV
#EXPOSURE CONCENTRATION vs. effect size, by study
#Using previously calculated Hedges' d and variance of d
ggplot(ES_DS_totalmort2, 
       aes(x = Sed_level_standardized,
           y = logRR,
           color = Ref,
           ymin = logRR-logRR_var,
           ymax = logRR+logRR_var)) + 
  geom_pointrange() +
  ggtitle("Colony mortality rate") +
  labs(x = "Sediment exposure concentration (mg/cm2/day)",
       y = "Log Relative Risk +/- s.d.",
       color = "Study") +
  geom_abline(intercept=0, slope=0) +
  theme_classic() +
  scale_x_log10(limits = c(1,100), breaks=c(1,10,100), 
                label=c("1","10","100"))


##WITHOUT HODGSON 1989 PHD CHAPTER IV
#EXPOSURE CONCENTRATION vs. effect size, by study
ggplot(ES_DS_totalmort_b2, 
       aes(x = Sed_level_standardized,
           y = logRR,
           color = Ref,
           ymin = logRR-logRR_var,
           ymax = logRR+logRR_var)) + 
  geom_pointrange() +
  ggtitle("Colony mortality rate") +
  labs(x = "Sediment exposure concentration (mg/cm2/day)",
       y = "Log Relative Risk +/- s.d.",
       color = "Study") +
  geom_abline(intercept=0, slope=0) +
  theme_classic() +
  scale_x_log10(limits = c(1,100), breaks=c(1,10,100), 
                label=c("1","10","100"))

#by SEDIMENT TYPE
ggplot(ES_DS_totalmort_b2, 
       aes(x = Sed_level_standardized,
           y = logRR,
           color = Sed_grain_category,
           shape = Sed_source,
           ymin = logRR-logRR_var,
           ymax = logRR+logRR_var)) + 
  geom_pointrange() +
  ggtitle("Colony mortality rate") +
  labs(x = "Sediment cumulative exposure (mg/cm2)",
       y = "Log Relative Risk +/- s.d.",
       color = "Sediment relative\ngrain size",
       shape = "Sediment source") +
  geom_abline(intercept=0, slope=0) +
  theme_classic() +
  scale_x_log10(limits = c(1,100), breaks=c(1,10,100), 
                label=c("1","10","100"))
ggplot(ES_DS_totalmort_b2, 
       aes(x = Sed_level_standardized,
           y = logRR,
           color = Sed_grain_class,
           shape = Sed_mineralogy,
           ymin = logRR-logRR_var,
           ymax = logRR+logRR_var)) + 
  geom_pointrange() +
  ggtitle("Colony mortality rate") +
  labs(x = "Sediment cumulative exposure (mg/cm2)",
       y = "Log Relative Risk +/- s.d.",
       color = "Sediment\ngrain class",
       shape = "Sediment mineralogy") +
  geom_abline(intercept=0, slope=0) +
  theme_classic() +
  scale_x_log10(limits = c(1,100), breaks=c(1,10,100), 
                label=c("1","10","100"))

#by PHYLOGENY
ggplot(ES_DS_totalmort_b2, 
       aes(x = Sed_level_standardized,
           y = logRR,
           color = Gsp,
           ymin = logRR-logRR_var,
           ymax = logRR+logRR_var)) + 
  geom_pointrange() +
  ggtitle("Colony mortality rate") +
  labs(x = "Sediment cumulative exposure (mg/cm2)",
       y = "Log Relative Risk +/- s.d.",
       color = "Species") +
  geom_abline(intercept=0, slope=0) +
  theme_classic() +
  scale_x_log10(limits = c(1,100), breaks=c(1,10,100), 
                label=c("1","10","100"))

#by OCEAN and REGION
ggplot(ES_DS_totalmort_b2, 
       aes(x = Sed_level_standardized,
           y = logRR,
           color = Region,
           shape = Ocean,
           ymin = logRR-logRR_var,
           ymax = logRR+logRR_var)) + 
  geom_pointrange() +
  ggtitle("Colony mortality rate") +
  labs(x = "Sediment cumulative exposure (mg/cm2)",
       y = "Log Relative Risk +/- s.d.",
       color = "Region",
       shape = "Ocean") +
  geom_abline(intercept=0, slope=0) +
  theme_classic() +
  scale_x_log10(limits = c(1,100), breaks=c(1,10,100), 
                label=c("1","10","100"))
```
  
There does not appear to be a relationship between sediment exposure and effect size. Also, there does not appear to be strong effects of phylogeny (genus) or geography (ocean and region). Sediment type (source, origin, grain size) is yet to be analyzed.
  
## B. Linear and Non-Linear Models   
The following set of functions model fixed and random (nested) effects WITHOUT accounting for within-comparison (or within-study) correlations. The output for each is a univariate, random-effects meta-regression.

```{r, echo=FALSE, warning=FALSE}
#?mixmeta

###ALL ANALYSES WITHOUT HODGSON 1989 PHD CHAPTER IV (a or b)

# Organizational Note: 
#All models with 'b' at the end of its name have exposure concentration as its predictor. 

# LINEAR FIXED AND RANDOM EFFECTS NOT ACCOUNTING FOR WITHIN-STUDY CORRELATIONS
mod_DS_totalmort1b <- mixmeta(logRR ~ 0 + Sed_level_standardized, S = logRR_var,
                              random =  ~ 0 + Sed_level_standardized | Comparison,
                              data = ES_DS_totalmort_b2, method = "ml")

# LINEAR FIXED AND RANDOM, NESTED EFFECTS NOT ACCOUNTING FOR WITHIN-STUDY CORRELATIONS
mod_ES_DS_totalmort2b <- mixmeta(logRR ~ 0 + Sed_level_standardized, S = logRR_var,
                              random =  ~ 0 + Sed_level_standardized | Ref/Comparison,
                              data = ES_DS_totalmort_b2, method = "ml")

# LINEAR FIXED AND RANDOM EFFECTS ACCOUNTING FOR WITHIN-STUDY CORRELATIONS
mod_DS_totalmort3b <- mixmeta(logRR ~ 0 + Sed_level_standardized,
                              random =  ~ 0 + Sed_level_standardized | Comparison,
                              data = ES_DS_totalmort_b2, method = "ml",
                              control=list(addSlist=addS_ES_DS_totalmort_b))

# LINEAR FIXED AND RANDOM, NESTED EFFECTS ACCOUNTING FOR WITHIN-STUDY CORRELATIONS
#Error in getSlist(S, nay, groups, m, k, control$addSlist, control$checkPD) : 'addSlist' not consistent with required format and grouping length. See help(mixmeta.control)
#mod_DS_totalmort4b <- mixmeta(logRR ~ 0 + Sed_level_standardized,
#                              random =  ~ 0 + Sed_level_standardized | Ref/Comparison,
#                              data = ES_DS_totalmort_b2, method = "ml",
#                              control=list(addSlist=addS_ES_DS_totalmort_b))
```
  
      
## C. Model Comparison and Summary
And now to compare relative fit of linear models...
```{r, echo=FALSE, warning=FALSE}
AIC(mod_DS_totalmort1b, mod_ES_DS_totalmort2b, mod_DS_totalmort3b)
#of these, model 3b has lowest AIC
```
  
And now to evaluate residuals of models...
  
```{r, echo=FALSE, warning=FALSE}
# Residuals vs. Fitted Plots and Normal Q-Q Plot
op <- par(mfrow = c(1,2), mar = c(2,2,4,1))

resid_b <- resid(mod_DS_totalmort3b)
fitted_b <- fitted(mod_DS_totalmort3b)
plot(fitted_b, resid_b, main = "B: logRR ~ mg_cm2_d, control = addSlist,
     random =  ~ mg_cm2_d | Comparison")
abline(0,0)

qqnorm(resid_b)
qqline(resid_b)

par(op)
```
   
Patterns:
1.  Looks ok despite sparse data.
   
## D. Best-Fit Model Summary
  
### MG/CM^2^/DAY IS PREDICTOR
Now I will predict the effect size along the exposure range and plot with confidence intervals.  

```{r, echo=FALSE, warning=FALSE}
summary(mod_DS_totalmort3b)

# PREDICT THE EFFECT SIZE FOR 13.2 mg/cm2 FROM TWO MODELS
#predict(mod_DS_totalmort3b, newdata=data.frame(Sed_level_standardized=13.2), ci=TRUE)

# PREDICT THE EFFECT SIZE AND PLOT WITH CONFIDENCE INTERVALS
pred_mod_DS_totalmort3b <- predict(mod_DS_totalmort3b, newdata=ES_DS_totalmort_b2, ci=TRUE)
ES_DS_totalmort_b2_CI2 <- cbind(ES_DS_totalmort_b2, pred_mod_DS_totalmort3b)
#head(ES_DS_totalmort_b2_CI2)
min_conc <- ES_DS_totalmort_b2_CI2 %>% 
  mutate(overlap0 = 0 >= ci.lb & 0 <= ci.ub) %>% 
  filter(overlap0==FALSE) %>% 
  summarize(min_conc=min(Sed_exposure_mg_cm2))
min_conc2 <- as.numeric(min_conc)

ggplot(ES_DS_totalmort_b2_CI2, 
       aes(x = Sed_level_standardized,
           y = logRR,
           color = Ref,
           ymin = logRR-logRR_var,
           ymax = logRR+logRR_var)) + 
  geom_pointrange() +
  labs(x = expression("Sediment exposure concentration (mg/cm"^"2"*"/day)"),
       y = "Log Relative Risk +/- s.d.",
       color = "Study") +
  geom_abline(intercept=0, slope=0) +
  geom_vline(xintercept=min_conc2, linetype="dashed", color = "red") +
  geom_ribbon(aes(x = Sed_level_standardized, y = logRR,
                  ymin = ci.lb, ymax = ci.ub), 
              fill = "grey70", alpha = .15, 
              show.legend=FALSE, inherit.aes=FALSE) +
  geom_line(aes(x = Sed_level_standardized, y = fit), inherit.aes=FALSE) +
  theme_classic() +
  scale_x_log10(limits=c(1,100),breaks=c(1,10,100,min_conc2),
                label=c("1","10","100",round(min_conc2,digits=1)))
```
  
  
# II. Suspended Sediment  
  
## A. Plots  

```{r, echo=FALSE, warning=FALSE}
#EXPOSURE CONCENTRATION vs. effect size, by study
ggplot(ES_SS_totalmort2, 
       aes(x = Sed_level_standardized,
           y = logRR,
           color = Ref,
           ymin = logRR-logRR_var,
           ymax = logRR+logRR_var)) + 
  geom_pointrange() +
  ggtitle("Colony mortality rate") +
  labs(x = "Sediment exposure concentration (mg/L)",
       y = "Log Relative Risk +/- s.d.",
       color = "Study") +
  geom_abline(intercept=0, slope=0) +
  theme_classic() +
  scale_x_log10(limits = c(1,200), breaks=c(1,10,100,1000), 
                label=c("1","10","100","1000"))

#by SEDIMENT TYPE
ggplot(ES_SS_totalmort2, 
       aes(x = Sed_level_standardized,
           y = logRR,
           color = Sed_grain_category,
           shape = Sed_source,
           ymin = logRR-logRR_var,
           ymax = logRR+logRR_var)) + 
  geom_pointrange() +
  ggtitle("Colony mortality rate") +
  labs(x = "Sediment cumulative exposure (mg x day/L)",
       y = "Log Relative Risk +/- s.d.",
       color = "Sediment relative\ngrain size",
       shape = "Sediment source") +
  geom_abline(intercept=0, slope=0) +
  theme_classic() +
  scale_x_log10(limits = c(1,200), breaks=c(1,10,100,1000), 
                label=c("1","10","100","1000"))
ggplot(ES_SS_totalmort2, 
       aes(x = Sed_level_standardized,
           y = logRR,
           color = Sed_grain_class,
           shape = Sed_mineralogy,
           ymin = logRR-logRR_var,
           ymax = logRR+logRR_var)) + 
  geom_pointrange() +
  ggtitle("Colony mortality rate") +
  labs(x = "Sediment cumulative exposure (mg x day/L)",
       y = "Log Relative Risk +/- s.d.",
       color = "Sediment\ngrain class",
       shape = "Sediment mineralogy") +
  geom_abline(intercept=0, slope=0) +
  theme_classic() +
  scale_x_log10(limits = c(1,200), breaks=c(1,10,100,1000), 
                label=c("1","10","100","1000"))

#by PHYLOGENY
ggplot(ES_SS_totalmort2, 
       aes(x = Sed_level_standardized,
           y = logRR,
           color = Gsp,
           ymin = logRR-logRR_var,
           ymax = logRR+logRR_var)) + 
  geom_pointrange() +
  ggtitle("Colony mortality rate") +
  labs(x = "Sediment cumulative exposure (mg x day/L)",
       y = "Log Relative Risk +/- s.d.",
       color = "Species") +
  geom_abline(intercept=0, slope=0) +
  theme_classic() +
  scale_x_log10(limits = c(1,200), breaks=c(1,10,100,1000), 
                label=c("1","10","100","1000"))

#by OCEAN and REGION
ggplot(ES_SS_totalmort2, 
       aes(x = Sed_level_standardized,
           y = logRR,
           color = Region,
           shape = Ocean,
           ymin = logRR-logRR_var,
           ymax = logRR+logRR_var)) + 
  geom_pointrange() +
  ggtitle("Colony mortality rate") +
  labs(x = "Sediment cumulative exposure (mg x day/L)",
       y = "Log Relative Risk +/- s.d.",
       color = "Region",
       shape = "Ocean") +
  geom_abline(intercept=0, slope=0) +
  theme_classic() +
  scale_x_log10(limits = c(1,200), breaks=c(1,10,100,1000), 
                label=c("1","10","100","1000"))
```
  
There does not appear to be a relationship between sediment exposure and effect size. Also, there does not appear to be strong effects of phylogeny (genus) or geography (ocean and region). Sediment type (source, origin, grain size) is yet to be analyzed.
  
## B. Model Fitting  
The following set of functions model fixed and random (nested) effects WITHOUT accounting for within-comparison (or within-study) correlations. The output for each is a univariate, random-effects meta-regression.
  
```{r, echo=FALSE, warning=FALSE}
#?mixmeta

# Organizational Note: 
#All models with 'b' at the end of its name have exposure concentration as its predictor. 

# LINEAR FIXED AND RANDOM EFFECTS NOT ACCOUNTING FOR WITHIN-STUDY CORRELATIONS
mod_ES_SS_totalmort1b <- mixmeta(logRR ~ 0 + Sed_level_standardized, S = logRR_var,
                              random =  ~ 0 + Sed_level_standardized | Comparison,
                              data = ES_SS_totalmort2, method = "ml")

# LINEAR FIXED AND RANDOM, NESTED EFFECTS NOT ACCOUNTING FOR WITHIN-STUDY CORRELATIONS
mod_ES_SS_totalmort2b <- mixmeta(logRR ~ 0 + Sed_level_standardized, S = logRR_var,
                              random =  ~ 0 + Sed_level_standardized | Ref/Comparison,
                              data = ES_SS_totalmort2, method = "ml")

# LINEAR FIXED AND RANDOM EFFECTS ACCOUNTING FOR WITHIN-STUDY CORRELATIONS
#Error in getSlist(S, nay, groups, m, k, control$addSlist, control$checkPD) : 'addSlist' not consistent with required format and grouping length. See help(mixmeta.control)
#mod_ES_SS_totalmort3b <- mixmeta(logRR ~ 0 + Sed_level_standardized,
#                              random =  ~ 0 + Sed_level_standardized | Comparison,
#                              data = ES_SS_totalmort2, method = "ml",
#                              control=list(addSlist=addS_ES_DS_totalmort_b))

# LINEAR FIXED AND RANDOM, NESTED EFFECTS ACCOUNTING FOR WITHIN-STUDY CORRELATIONS
#Error in getSlist(S, nay, groups, m, k, control$addSlist, control$checkPD) : 'addSlist' not consistent with required format and grouping length. See help(mixmeta.control)
#mod_ES_SS_totalmort4b <- mixmeta(logRR ~ 0 + Sed_level_standardized,
#                              random =  ~ 0 + Sed_level_standardized | Ref/Comparison,
#                              data = ES_SS_totalmort2, method = "ml",
#                              control=list(addSlist=addS_ES_DS_totalmort_b))
```
  
## C. Model Comparison and Summary
And now to compare relative fit of linear models...
```{r, echo=FALSE, warning=FALSE}
AIC(mod_ES_SS_totalmort1b, mod_ES_SS_totalmort2b)
#of these, model 1b has lowest AIC (but 2b is within 2 AIC)
```
  
And now to evaluate residuals of models...
  
```{r, echo=FALSE, warning=FALSE}
# Residuals vs. Fitted Plots and Normal Q-Q Plot
op <- par(mfrow = c(1,2), mar = c(2,2,4,1))

resid_b <- resid(mod_ES_SS_totalmort1b)
fitted_b <- fitted(mod_ES_SS_totalmort1b)
plot(fitted_b, resid_b, main = "B: logRR ~ mg_L, S = logRR_var,
     random =  ~ mg_L | Comparison")
abline(0,0)

qqnorm(resid_b)
qqline(resid_b)

par(op)
```
   
Patterns:
1.  Sparse data but magnitude of residuals small.
   
## D. Best-Fit Model Summary
  
### MG/L IS PREDICTOR, Model using dosresmeta calculation of Hedges' *d*
Now I will predict the effect size along the exposure range and plot with confidence intervals.  

```{r, echo=FALSE, warning=FALSE}
summary(mod_ES_SS_totalmort1b)

# PREDICT THE EFFECT SIZE FOR 13.2 mg/cm2 FROM TWO MODELS
#predict(mod_ES_SS_totalmort1b, newdata=data.frame(Sed_level_standardized=13.2), ci=TRUE)

# PREDICT THE EFFECT SIZE AND PLOT WITH CONFIDENCE INTERVALS
pred_mod_ES_SS_totalmort1b <- predict(mod_ES_SS_totalmort1b, newdata=ES_SS_totalmort2, ci=TRUE)
ES_SS_totalmort2_CI2 <- cbind(ES_SS_totalmort2, pred_mod_ES_SS_totalmort1b)
#head(ES_SS_totalmort2_CI2)
min_conc_SS <- ES_SS_totalmort2_CI2 %>% 
  mutate(overlap0 = 0 >= ci.lb & 0 <= ci.ub) %>% 
  filter(overlap0==FALSE) %>% 
  summarize(min_conc_SS=min(Sed_level_standardized))
min_conc_SS2 <- as.numeric(min_conc_SS)

ggplot(ES_SS_totalmort2_CI2, 
       aes(x = Sed_level_standardized,
           y = logRR,
           color = Ref,
           ymin = logRR-logRR_var,
           ymax = logRR+logRR_var)) + 
  geom_pointrange() +
  labs(x = "Sediment exposure concentration (mg/L)",
       y = "Log Relative Risk +/- s.d.",
       color = "Study") +
  geom_abline(intercept=0, slope=0) +
  geom_vline(xintercept=min_conc_SS2, linetype="dashed", color = "red") +
  geom_ribbon(aes(x = Sed_level_standardized, y = logRR,
                  ymin = ci.lb, ymax = ci.ub), 
              fill = "grey70", alpha = .15, 
              show.legend=FALSE, inherit.aes=FALSE) +
  geom_line(aes(x = Sed_level_standardized, y = fit), inherit.aes=FALSE) +
  theme_classic() +
  scale_x_log10(limits = c(1,200), breaks=c(1,10,100,1000,min_conc_SS2), 
                label=c("1","10","100","1000",round(min_conc_SS2,digits=1)))
```
  