---
title: "Effects of sediment on the photosynthesis of adult corals"
author: "Lillian J. Tuttle"
date: "9/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(mixmeta)
library(dosresmeta)
library(splines)

effectsize <- read.csv("data/effectsize_ALL.csv", header=T)
```
  
```{r, echo=FALSE, warning=FALSE}
#Looking at data
#head(effectsize) 

#a lot of variables (columns) are factors instead of numeric/doubles/integers
effectsize <- effectsize %>% #converting appropriate columns to numeric...
  mutate_at(vars(ES, Hedges_d, Var_d, SE_d, CI_lo_d, CI_up_d, Sed_level_standardized, Sed_exposure, Sed_exposure_mg_d_L), as.character) %>% 
  mutate_at(vars(ES, Hedges_d, Var_d, SE_d, CI_lo_d, CI_up_d, Sed_level_standardized, Sed_exposure, Sed_exposure_mg_d_L), as.numeric)
#head(effectsize) #"double"-check!

#Now create a new variable that has both Genus and species, remove sediment burial experiments, and remove irrelevant rows (non-controls, -treatments, or -baselines)
effectsize <- effectsize %>%
  mutate(Gsp = paste(Updated_Genus, Updated_species, sep = "_")) %>% 
  filter(Binary_sed_burial=="0") %>%
  filter(!Control=="-1")
#head(effectsize)

#Separate into deposited sediment (DS) and suspended sediment (SS) datasets, WITH CONTROLS
ES_DS <- effectsize %>% filter(DS_SS=="DS")
ES_SS <- effectsize %>% filter(DS_SS=="SS")

#Second dataset that that removes controls (Control == c("1","2"))
effectsize2 <- effectsize %>% filter(Control=="0")
#effectsize2 #"double"-check!

#Separate into deposited sediment (DS) and suspended sediment (SS) datasets, WITHOUT CONTROLS
ES_DS2 <- effectsize2 %>% filter(DS_SS=="DS")
ES_SS2 <- effectsize2 %>% filter(DS_SS=="SS")

#Now separate into independent datasets for each coral response for DEPOSITED SEDIMENT
ES_DS_deltaf_fm <- ES_DS %>% filter(Response == "deltaf_fm")
ES_DS_fv_fm <- ES_DS %>% filter(Response == "fv_fm")
ES_DS_p_r <- ES_DS %>% filter(Response == "p_r_ratio")
#ES_DS_deltaf_fm %>% count(Ref, sort=T) #n=1 study -- not suitable for meta-analysis ****
#ES_DS_fv_fm %>% count(Ref, sort=T) #n=8 studies from 6 papers
#ES_DS_p_r %>% count(Ref, sort=T) #n=3 studies from 3 papers -- suitable for meta-analysis with so few studies?
#ES_DS_fv_fm %>% tally()
#ES_DS_fv_fm %>% count(Ref)
#ES_DS_fv_fm %>% count(Ref_name)
#ES_DS_fv_fm %>% count(Gsp)
#ES_DS_fv_fm %>% count(Updated_Genus)
#ES_DS_p_r %>% tally()
#ES_DS_p_r %>% count(Ref)
#ES_DS_p_r %>% count(Ref_name)
#ES_DS_p_r %>% count(Gsp)
#ES_DS_p_r %>% count(Updated_Genus)

#Now separate into independent datasets for each coral response for SUSPENDED SEDIMENT
ES_SS_fv_fm <- ES_SS %>% filter(Response == "fv_fm")
ES_SS_p_r <- ES_SS %>% filter(Response == "p_r_ratio")
#ES_SS_fv_fm %>% count(Ref, sort=T) #n=5 studies from 4 papers
#ES_SS_p_r %>% count(Ref, sort=T) #n=1 study -- not suitable for meta-analysis ****
#ES_SS_fv_fm %>% tally()
#ES_SS_fv_fm %>% count(Ref)
#ES_SS_fv_fm %>% count(Ref_name)
#ES_SS_fv_fm %>% count(Gsp)
#ES_SS_fv_fm %>% count(Updated_Genus)
``` 
  
These datasets contain information from multiple articles (**Ref_name**) and studies/experiments (**Ref**) within those articles about the effects of DEPOSITED & SUSPENDED sediment on the **photosynthetic yield (Fv/Fm or deltaF/Fm) or the photosynthesis-respiration (P/R) ratio** of symbiotic algae living within coral cells. I would like to explore the dose-response relationship between sediment 'dose' and the effect size, estimated here as the standardized mean difference between treatment and control, in terms of Hedges' *d*:

- Hedges' *d* is the standardized mean difference between the treatment and control groups (Hedges and Olkin 1985). Each study may have multiple Hedges' *d* calculations, one for each treatment-control comparison at each time-point.   

- 'Dose' can be quantified as sediment exposure concentration (mg/cm^2^/day for deposited sediment and mg/L for suspended sediment), exposure duration (days), or as "cumulative exposure", for which I multiply concentration and duration (mg/cm^2^ for deposited sediment and mg x day/L for suspended sediment).

Each study/experiment (Ref) may contain multiple sets of controls and treatments (due to using multiple species or sediment types, for instance). Thus, I have added two categorical vectors, '**Comparison**' and 'Control', which group each set of control-treatment comparisons into a unique two-letter code, within which is one control (Control=="1") and at least one treatment (Control=="0"). **Therefore, 'Comparison' is nested within 'Ref', which is further nested within 'Ref_name'.**

```{r, echo=FALSE}
#Using dosresmeta to calculate Hedges' d and variance of d
covar_DS_photo <- by(ES_DS_fv_fm, ES_DS_fv_fm$Comparison, function(x) 
  covar.smd(Tx_mean, Tx_sd, Tx_n, "smd", method="hedges", data = x))
ES_DS_fv_fm$smd <- unlist(lapply(covar_DS_photo, function(x) x$y))
ES_DS_fv_fm$vmd <- unlist(lapply(covar_DS_photo, function(x) x$v))

covar_DS_p_r <- by(ES_DS_p_r, ES_DS_p_r$Comparison, function(x) 
  covar.smd(Tx_mean, Tx_sd, Tx_n, "smd", method="hedges", data = x))
ES_DS_p_r$smd <- unlist(lapply(covar_DS_p_r, function(x) x$y))
ES_DS_p_r$vmd <- unlist(lapply(covar_DS_p_r, function(x) x$v))

covar_SS_photo <- by(ES_SS_fv_fm, ES_SS_fv_fm$Comparison, function(x) 
  covar.smd(Tx_mean, Tx_sd, Tx_n, "smd", method="hedges", data = x))
ES_SS_fv_fm$smd <- unlist(lapply(covar_SS_photo, function(x) x$y))
ES_SS_fv_fm$vmd <- unlist(lapply(covar_SS_photo, function(x) x$v))

#Looking at data to make sure smd and vmd added as columns
#ES_DS_fv_fm
#ES_DS_p_r
#ES_SS_fv_fm

#Removing controls for modeling
ES_DS_fv_fm2 <- subset(ES_DS_fv_fm, Control=="0")
ES_DS_p_r2 <- subset(ES_DS_p_r, Control=="0")
ES_SS_fv_fm2 <- subset(ES_SS_fv_fm, Control=="0")

#Looking at data to make sure controls were removed
#ES_DS_fv_fm2
#ES_DS_p_r2
#ES_SS_fv_fm2
#ES_DS_fv_fm2 %>% tally()
#ES_DS_p_r2 %>% tally()
#ES_SS_fv_fm2 %>% tally()
``` 
  
# I. Photosynthetic Efficiency  

## A. Deposited Sediment  
  
### 1. Plots  
I calculated the standardized mean difference in terms of Hedges' *d* and plot it here with respect to cumulative exposure. I did this first for a Hedges' *d* that I calculated before importing the data to R and again for a Hedges' *d* that is calculated by the dosresmeta::covar.smd function.

```{r, echo=FALSE, warning=FALSE}
#EXPOSURE CONCENTRATION vs. effect size, by study
#Using previously calculated Hedges' d and variance of d
ggplot(ES_DS_fv_fm2, 
       aes(x = Sed_level_standardized,
           y = Hedges_d,
           color = Ref,
           ymin = Hedges_d-Var_d,
           ymax = Hedges_d+Var_d)) + 
  geom_pointrange() +
  ggtitle("Photosynthetic efficiency, Fv/Fm") +
  labs(x = "Sediment exposure concentration (mg/cm2/day)",
       y = "Effect size (Hedges' d +/- s.d., manual calculation)",
       color = "Study") +
  geom_abline(intercept=0, slope=0) +
  theme_classic() +
  scale_x_log10(limits=c(0.1,1000), breaks=c(0.1,1,10,100,1000), 
                label=c("0.1","1","10","100","1000")) 
#Using dosresmeta's calculation of Hedges' d and variance of d
ggplot(ES_DS_fv_fm2, 
       aes(x = Sed_level_standardized,
           y = smd,
           color = Ref,
           ymin = smd-vmd,
           ymax = smd+vmd)) + 
  geom_pointrange() +
  ggtitle("Photosynthetic efficiency, Fv/Fm") +
  labs(x = "Sediment exposure concentration (mg/cm2/day)",
       y = "Effect size (Hedges' d +/- s.d., dosresmeta calculation)",
       color = "Study") +
  geom_abline(intercept=0, slope=0) +
  theme_classic() +
  scale_x_log10(limits=c(0.1,1000), breaks=c(0.1,1,10,100,1000), 
                label=c("0.1","1","10","100","1000")) 


#by SEDIMENT TYPE
#Using previously calculated Hedges' d and variance of d
ggplot(ES_DS_fv_fm2, 
       aes(x = Sed_level_standardized,
           y = smd,
           color = Sed_grain_category,
           shape = Sed_source,
           ymin = smd-vmd,
           ymax = smd+vmd)) + 
  geom_pointrange() +
  ggtitle("Photosynthetic efficiency, Fv/Fm") +
  labs(x = "Sediment exposure concentration (mg/cm2/day)",
       y = "Effect size (Hedges' d +/- s.d., dosresmeta calculation)",
       color = "Sediment relative\ngrain size",
       shape = "Sediment source") +
  geom_abline(intercept=0, slope=0) +
  theme_classic() +
  scale_x_log10(limits=c(0.1,1000), breaks=c(0.1,1,10,100,1000), 
                label=c("0.1","1","10","100","1000")) 
ggplot(ES_DS_fv_fm2, 
       aes(x = Sed_level_standardized,
           y = smd,
           color = Sed_grain_class,
           shape = Sed_mineralogy,
           ymin = smd-vmd,
           ymax = smd+vmd)) + 
  geom_pointrange() +
  ggtitle("Photosynthetic efficiency, Fv/Fm") +
  labs(x = "Sediment exposure concentration (mg/cm2/day)",
       y = "Effect size (Hedges' d +/- s.d., dosresmeta calculation)",
       color = "Sediment\ngrain class",
       shape = "Sediment mineralogy") +
  geom_abline(intercept=0, slope=0) +
  theme_classic() +
  scale_x_log10(limits=c(0.1,1000), breaks=c(0.1,1,10,100,1000), 
                label=c("0.1","1","10","100","1000")) 

#by PHYLOGENY
#Using previously calculated Hedges' d and variance of d
ggplot(ES_DS_fv_fm2, 
       aes(x = Sed_level_standardized,
           y = smd,
           color = Updated_Genus,
           ymin = smd-vmd,
           ymax = smd+vmd)) + 
  geom_pointrange() +
  ggtitle("Photosynthetic efficiency, Fv/Fm") +
  labs(x = "Sediment exposure concentration (mg/cm2/day)",
       y = "Effect size (Hedges' d +/- s.d., dosresmeta calculation)",
       color = "Genus") +
  geom_abline(intercept=0, slope=0) +
  theme_classic() +
  scale_x_log10(limits=c(0.1,1000), breaks=c(0.1,1,10,100,1000), 
                label=c("0.1","1","10","100","1000")) 

#by OCEAN and REGION
#Using previously calculated Hedges' d and variance of d
ggplot(ES_DS_fv_fm2, 
       aes(x = Sed_level_standardized,
           y = smd,
           color = Region,
           shape = Ocean,
           ymin = smd-vmd,
           ymax = smd+vmd)) + 
  geom_pointrange() +
  ggtitle("Photosynthetic efficiency, Fv/Fm") +
  labs(x = "Sediment exposure concentration (mg/cm2/day)",
       y = "Effect size (Hedges' d +/- s.d., dosresmeta calculation)",
       color = "Region",
       shape = "Ocean") +
  geom_abline(intercept=0, slope=0) +
  theme_classic() +
  scale_x_log10(limits=c(0.1,1000), breaks=c(0.1,1,10,100,1000), 
                label=c("0.1","1","10","100","1000")) 
```
  
It is interesting that there is a difference between how I calculated Hedges' *d* and how *dosresmeta* appears to calculate it. But in both cases, there may be a negative relationship between sediment exposure and effect size. There does not appear to be strong effects of phylogeny (genus) or geography (ocean and region). There may be effects of sediment type (source, origin, grain size), but I do not think these will be discernible given the variance.
  
### 2. Model Fitting
The following set of functions model fixed and random (nested) effects WITHOUT accounting for within-comparison (or within-study) correlations. The output for each is a univariate, random-effects meta-regression.
  
```{r, echo=FALSE, warning=FALSE}
#?mixmeta

# Organizational Note: 
#All models have exposure concentration as its predictor. 
#All models with 'd' at the end of its name have covar.smd calculated Hedges' d as its response.

# LINEAR FIXED AND RANDOM EFFECTS NOT ACCOUNTING FOR WITHIN-STUDY CORRELATIONS
mod_DS_photo_1d <- mixmeta(smd ~ 0 + Sed_level_standardized, S = vmd,
                      random =  ~ 0 + Sed_level_standardized | Comparison,
                      data = ES_DS_fv_fm2, method = "ml")

# LINEAR FIXED AND RANDOM, NESTED EFFECTS NOT ACCOUNTING FOR WITHIN-STUDY CORRELATIONS
mod_DS_photo_2d <- mixmeta(smd ~ 0 + Sed_level_standardized, S = vmd,
                      random =  ~ 0 + Sed_level_standardized | Ref/Comparison,
                      data = ES_DS_fv_fm2, method = "ml")

# LINEAR FIXED AND RANDOM, NESTED EFFECTS NOT ACCOUNTING FOR WITHIN-STUDY CORRELATIONS
mod_DS_photo_3d <- mixmeta(smd ~ 0 + Sed_level_standardized, S = vmd,
                      random =  ~ 0 + Sed_level_standardized | Ref_name/Comparison,
                      data = ES_DS_fv_fm2, method = "ml")

# LINEAR FIXED AND RANDOM, NESTED EFFECTS NOT ACCOUNTING FOR WITHIN-STUDY CORRELATIONS
mod_DS_photo_4d <- mixmeta(smd ~ 0 + Sed_level_standardized, S = vmd,
                       random =  ~ 0 + Sed_level_standardized | Ref_name/Ref/Comparison,
                       data = ES_DS_fv_fm2, method = "ml")
```

To account for within-comparison correlations in the mixmeta model, I replace the 'S' argument with a 'control' argument, which requires me to provide the covariance matrices, as done with the covar.smd function. 

```{r, echo=FALSE, warning=FALSE}
# COMPUTE THE WITHIN-COMPARISON CORRELATIONS 
# already done above but code repeated here...
#covar_DS_photo <- by(ES_DS_fv_fm, ES_DS_fv_fm$Comparison, function(x) 
#  covar.smd(Tx_mean, Tx_sd, Tx_n, "smd", method="hedges", data = x))


# LINEAR FIXED AND RANDOM EFFECTS ACCOUNTING FOR WITHIN-COMPARISON CORRELATIONS
#Code for covariance matrix to include for addSlist for Comparison random effect
newlist_FVFM <- list(NA)
for (i in seq(1,length(covar_DS_photo))) {
newlist_FVFM[i] <- list(covar_DS_photo[[i]]$S)
}
mod_DS_photo_5d <- mixmeta(smd ~ 0 + Sed_level_standardized,
                       random = ~ 0 + Sed_level_standardized | Comparison,
                       data=ES_DS_fv_fm2, method="ml", #using dataset without controls
                       control=list(addSlist=newlist_FVFM))


# LINEAR FIXED AND RANDOM, NESTED EFFECTS ACCOUNTING FOR WITHIN-COMPARISON & WITHIN-STUDY CORRELATIONS
#Code for covariance matrix to include for addSlist for Ref/Comparison
newlist_FVFM2 <- list(NA) #collects the list of covariance matrices for the block diag matrix by reference for the nested hierarchial ref/comparison
templist_FVFM <- list(NA) #holds temp list of covariance matrices associated with each reference
templist_FVFM2 <-list(NA)
reflista <- ES_DS_fv_fm %>% distinct(Ref,Ref_name,Comparison)
reflist <- reflista[,1]
for (i in seq(1,length(unique(reflista$Ref))))  {
  #pull the elements from covar_DS_photo that are all from the same reference [i]
  templist_FVFM[i] <-list(covar_DS_photo[reflist==unique(reflista$Ref)[i]])
  for (j in seq(1,length(templist_FVFM[[i]]))) {
  #for each comparison in the reference, pull out the covar matrices (element $S) and put in into templist_FVFM2
     templist_FVFM2[j] <- list(templist_FVFM[[i]][[j]]$S)
  }
  #turn list of covars from all comparison in one reference into block diag matrix
  newlist_FVFM2[i] <- list(bdiagMat(templist_FVFM2))
  templist_FVFM2 <- list(NA)
}
mod_DS_photo_6d <- mixmeta(smd ~ 0 + Sed_level_standardized,
                       random = ~ 0 + Sed_level_standardized | Ref/Comparison,
                       data=ES_DS_fv_fm2, method="ml", #using dataset without controls
                       control=list(addSlist=newlist_FVFM2))


# LINEAR FIXED AND RANDOM, NESTED EFFECTS ACCOUNTING FOR WITHIN-COMPARISON & WITHIN-STUDY CORRELATIONS
#Code for covariance matrix to include for addSlist for Ref_name/Comparison
newlist_FVFM3 <- list(NA) #collects the list of covariance matrices for the block diag matrix by reference for the nested hierarchial Ref_name/Comparison
templist_FVFM3 <- list(NA) #holds temp list of covariance matrices associated with each reference
templist_FVFM4 <-list(NA)
reflista <- ES_DS_fv_fm %>% distinct(Ref,Ref_name,Comparison)
reflist <- reflista[,2]
for (i in seq(1,length(unique(reflista$Ref_name))))  {
  #pull the elements from covar_DS_photo that are all from the same reference [i]
  templist_FVFM3[i] <-list(covar_DS_photo[reflist==unique(reflista$Ref_name)[i]])
  for (j in seq(1,length(templist_FVFM3[[i]]))) {
  #for each comparison in the reference, pull out the covar matrices (element $S) and put in into templist_FVFM4
     templist_FVFM4[j] <- list(templist_FVFM3[[i]][[j]]$S)
  }
  #turn list of covars from all comparison in one reference into block diag matrix
  newlist_FVFM3[i] <- list(bdiagMat(templist_FVFM4))
  templist_FVFM4 <- list(NA)
}
mod_DS_photo_7d <- mixmeta(smd ~ 0 + Sed_level_standardized,
                       random = ~ 0 + Sed_level_standardized | Ref_name/Comparison,
                       data=ES_DS_fv_fm2, method="ml", #using dataset without controls
                       control=list(addSlist=newlist_FVFM3))
```
  
And now to compare relative fit of linear models...

```{r, echo=FALSE, warning=FALSE}
# Model Comparisons
AIC(mod_DS_photo_1d, mod_DS_photo_2d, mod_DS_photo_3d, mod_DS_photo_4d, mod_DS_photo_5d, mod_DS_photo_6d, mod_DS_photo_7d) 
#of these, model 6d has lowest AIC

#ALTERNATIVE MODELS WITHOUT ZERO INTERCEPT
# LINEAR FIXED AND RANDOM, NESTED EFFECTS NOT ACCOUNTING FOR WITHIN-STUDY CORRELATIONS
alt_mod_DS_photo_6d <- mixmeta(smd ~ Sed_level_standardized,
                       random = ~ Sed_level_standardized | Ref/Comparison,
                       data=ES_DS_fv_fm2, method="ml", #using dataset without controls
                       control=list(addSlist=newlist_FVFM2))

AIC(alt_mod_DS_photo_6d, mod_DS_photo_6d)
#In all cases, the alternative models have lower AIC values, but take a look at the outrageous residuals from these alt models...
```
  
And now to evaluate residuals of models...

```{r, echo=FALSE, warning=FALSE}
#Models WITH zero intercept
# Residuals vs. Fitted Plots and Normal Q-Q Plots
op <- par(mfrow = c(1,2), mar = c(2,2,4,1))

resid_6d <- resid(mod_DS_photo_6d)
fitted_6d <- fitted(mod_DS_photo_6d)
plot(fitted_6d, resid_6d, main = "D: smd ~ 0+mg_cm2_d, S = addSlist,
     random =  ~ 0+mg_cm2_d | Ref/Comparison")
abline(0,0)

qqnorm(resid_6d)
qqline(resid_6d)

par(op)


#Models WITHOUT zero intercept
# Residuals vs. Fitted Plots and Normal Q-Q Plots
op <- par(mfrow = c(1,2), mar = c(2,2,4,1))

alt_resid_6d <- resid(alt_mod_DS_photo_6d)
alt_fitted_6d <- fitted(alt_mod_DS_photo_6d)
plot(alt_fitted_6d, alt_resid_6d, main = "D: smd ~ mg_cm2_d, S = addSlist,
     random =  ~ mg_cm2_d | Ref/Comparison")
abline(0,0)

qqnorm(alt_resid_6d)
qqline(alt_resid_6d)

par(op)
```
  
Patterns:
1.  "Clumping" of residuals that may be improved by log-transformation.
2.  The angle of the residual line does not look good for alternative models.
3.  First let's try a log transformation of sediment exposure for both alt and normal models.  
  
Now let's check out a log, base 10 transformation of sediment exposure to see if it improves the residuals of the models.
  
```{r, echo=FALSE, warning=FALSE}
#?mixmeta

# Organizational Note: 
#All models have exposure concentration as its predictor. 
#All models with 'd' at the end of its name have covar.smd calculated Hedges' d as its response.

log10_mg_cm2_d <- log10(ES_DS_fv_fm2$Sed_level_standardized)

# LINEAR FIXED AND RANDOM EFFECTS NOT ACCOUNTING FOR WITHIN-STUDY CORRELATIONS
log_mod_DS_photo_1d <- mixmeta(smd ~ 0 + log10_mg_cm2_d, S = vmd,
                      random =  ~ 0 + log10_mg_cm2_d | Comparison,
                      data = ES_DS_fv_fm2, method = "ml")

# LINEAR FIXED AND RANDOM, NESTED EFFECTS NOT ACCOUNTING FOR WITHIN-STUDY CORRELATIONS
log_mod_DS_photo_2d <- mixmeta(smd ~ 0 + log10_mg_cm2_d, S = vmd,
                      random =  ~ 0 + log10_mg_cm2_d | Ref/Comparison,
                      data = ES_DS_fv_fm2, method = "ml")

# LINEAR FIXED AND RANDOM, NESTED EFFECTS NOT ACCOUNTING FOR WITHIN-STUDY CORRELATIONS
log_mod_DS_photo_3d <- mixmeta(smd ~ 0 + log10_mg_cm2_d, S = vmd,
                      random =  ~ 0 + log10_mg_cm2_d | Ref_name/Comparison,
                      data = ES_DS_fv_fm2, method = "ml")

# LINEAR FIXED AND RANDOM, NESTED EFFECTS NOT ACCOUNTING FOR WITHIN-STUDY CORRELATIONS
log_mod_DS_photo_4d <- mixmeta(smd ~ 0 + log10_mg_cm2_d, S = vmd,
                       random =  ~ 0 + log10_mg_cm2_d | Ref_name/Ref/Comparison,
                       data = ES_DS_fv_fm2, method = "ml")
```

To account for within-comparison correlations in the mixmeta model, I replace the 'S' argument with a 'control' argument, which requires me to provide the covariance matrices, as done with the covar.smd function. 

```{r, echo=FALSE, warning=FALSE}
# COMPUTE THE WITHIN-COMPARISON CORRELATIONS 
# already done above but code repeated here...
#covar_DS_photo <- by(ES_DS_fv_fm, ES_DS_fv_fm$Comparison, function(x) 
#  covar.smd(Tx_mean, Tx_sd, Tx_n, "smd", method="hedges", data = x))


# LINEAR FIXED AND RANDOM EFFECTS ACCOUNTING FOR WITHIN-COMPARISON CORRELATIONS
#Code for covariance matrix to include for addSlist for Comparison random effect
newlist_log_FVFM <- list(NA)
for (i in seq(1,length(covar_DS_photo))) {
newlist_log_FVFM[i] <- list(covar_DS_photo[[i]]$S)
}
log_mod_DS_photo_5d <- mixmeta(smd ~ 0 + log10_mg_cm2_d,
                       random = ~ 0 + log10_mg_cm2_d | Comparison,
                       data=ES_DS_fv_fm2, method="ml", #using dataset without controls
                       control=list(addSlist=newlist_log_FVFM))


# LINEAR FIXED AND RANDOM, NESTED EFFECTS ACCOUNTING FOR WITHIN-COMPARISON & WITHIN-STUDY CORRELATIONS
#Code for covariance matrix to include for addSlist for Ref/Comparison
newlist_log_FVFM2 <- list(NA) #collects the list of covariance matrices for the block diag matrix by reference for the nested hierarchial ref/comparison
templist_log_FVFM <- list(NA) #holds temp list of covariance matrices associated with each reference
templist_log_FVFM2 <-list(NA)
reflista <- ES_DS_fv_fm %>% distinct(Ref,Ref_name,Comparison)
reflist <- reflista[,1]
for (i in seq(1,length(unique(reflista$Ref))))  {
  #pull the elements from covar_DS_photo that are all from the same reference [i]
  templist_log_FVFM[i] <-list(covar_DS_photo[reflist==unique(reflista$Ref)[i]])
  for (j in seq(1,length(templist_log_FVFM[[i]]))) {
  #for each comparison in the reference, pull out the covar matrices (element $S) and put in into templist_log_FVFM2
     templist_log_FVFM2[j] <- list(templist_log_FVFM[[i]][[j]]$S)
  }
  #turn list of covars from all comparison in one reference into block diag matrix
  newlist_log_FVFM2[i] <- list(bdiagMat(templist_log_FVFM2))
  templist_log_FVFM2 <- list(NA)
}
log_mod_DS_photo_6d <- mixmeta(smd ~ 0 + log10_mg_cm2_d,
                       random = ~ 0 + log10_mg_cm2_d | Ref/Comparison,
                       data=ES_DS_fv_fm2, method="ml", #using dataset without controls
                       control=list(addSlist=newlist_log_FVFM2))


# LINEAR FIXED AND RANDOM, NESTED EFFECTS ACCOUNTING FOR WITHIN-COMPARISON & WITHIN-STUDY CORRELATIONS
#Code for covariance matrix to include for addSlist for Ref_name/Comparison
newlist_log_FVFM3 <- list(NA) #collects the list of covariance matrices for the block diag matrix by reference for the nested hierarchial Ref_name/Comparison
templist_log_FVFM3 <- list(NA) #holds temp list of covariance matrices associated with each reference
templist_log_FVFM4 <-list(NA)
reflista <- ES_DS_fv_fm %>% distinct(Ref,Ref_name,Comparison)
reflist <- reflista[,2]
for (i in seq(1,length(unique(reflista$Ref_name))))  {
  #pull the elements from covar_DS_photo that are all from the same reference [i]
  templist_log_FVFM3[i] <-list(covar_DS_photo[reflist==unique(reflista$Ref_name)[i]])
  for (j in seq(1,length(templist_log_FVFM3[[i]]))) {
  #for each comparison in the reference, pull out the covar matrices (element $S) and put in into templist_log_FVFM4
     templist_log_FVFM4[j] <- list(templist_log_FVFM3[[i]][[j]]$S)
  }
  #turn list of covars from all comparison in one reference into block diag matrix
  newlist_log_FVFM3[i] <- list(bdiagMat(templist_log_FVFM4))
  templist_log_FVFM4 <- list(NA)
}
log_mod_DS_photo_7d <- mixmeta(smd ~ 0 + log10_mg_cm2_d,
                       random = ~ 0 + log10_mg_cm2_d | Ref_name/Comparison,
                       data=ES_DS_fv_fm2, method="ml", #using dataset without controls
                       control=list(addSlist=newlist_log_FVFM3))
```
  
### 3. Model Comparison and Summary
And now to compare relative fit of log-linear models...
  
```{r, echo=FALSE, warning=FALSE}
# Model Comparisons
AIC(log_mod_DS_photo_1d, log_mod_DS_photo_2d, log_mod_DS_photo_3d, log_mod_DS_photo_4d, log_mod_DS_photo_5d, log_mod_DS_photo_6d, log_mod_DS_photo_7d) 
#of these, model 6d has lowest AIC

#ALTERNATIVE MODELS WITHOUT ZERO INTERCEPT
# LINEAR FIXED AND RANDOM, NESTED EFFECTS NOT ACCOUNTING FOR WITHIN-STUDY CORRELATIONS
alt_log_mod_DS_photo_6d <- mixmeta(smd ~ log10_mg_cm2_d,
                       random = ~ log10_mg_cm2_d | Ref/Comparison,
                       data=ES_DS_fv_fm2, method="ml", #using dataset without controls
                       control=list(addSlist=newlist_log_FVFM2))

AIC(alt_log_mod_DS_photo_6d, log_mod_DS_photo_6d)
#alternative model has lower AIC values, but let's take a look at the residuals from these alt models...
```
  
And now to evaluate residuals of models...

```{r, echo=FALSE, warning=FALSE}
#Models WITH zero intercept
# Residuals vs. Fitted Plots and Normal Q-Q Plots
op <- par(mfrow = c(1,2), mar = c(2,2,4,1))

resid_log6d <- resid(log_mod_DS_photo_6d)
fitted_log6d <- fitted(log_mod_DS_photo_6d)
plot(fitted_log6d, resid_log6d, main = "D: smd ~ 0+log_mg_cm2_d, S = addSlist,
     random =  ~ 0+log_mg_cm2_d | Ref/Comparison")
abline(0,0)

qqnorm(resid_log6d)
qqline(resid_log6d)

par(op)


#Models WITHOUT zero intercept
# Residuals vs. Fitted Plots and Normal Q-Q Plots
op <- par(mfrow = c(1,2), mar = c(2,2,4,1))

resid_alt_log6d <- resid(alt_log_mod_DS_photo_6d)
fitted_alt_log6d <- fitted(alt_log_mod_DS_photo_6d)
plot(fitted_alt_log6d, resid_alt_log6d, main = "D: smd ~ log_mg_cm2_d, S = addSlist,
     random =  ~ log_mg_cm2_d | Ref/Comparison")
abline(0,0)

qqnorm(resid_alt_log6d)
qqline(resid_alt_log6d)

par(op)
```
  
Patterns:
1.  Log model much improves residuals -- lower magnitudes and more even spread. 
2.  Alt model looks terrible -- high magnitude residuals and severe misalignment between residual points and model average (zero abline).
  
### 4. Best-Fit Model Summary
  
#### MG/CM2/DAY IS PREDICTOR, Model using dosresmeta calculation of Hedges' *d*
Now I will predict the effect size along the exposure range and plot with confidence intervals.   
```{r, echo=FALSE, warning=FALSE}
summary(log_mod_DS_photo_6d)

# PREDICT THE EFFECT SIZE FOR 13.2 mg/cm2 FROM TWO MODELS (LOAEL identified by binary analyses)
#predict(log_mod_DS_photo_6d, newdata=data.frame(Sed_level_standardized=13.2), ci=TRUE)

# PREDICT THE EFFECT SIZE AND PLOT WITH CONFIDENCE INTERVALS
pred_log_mod_DS_photo_6d <- predict(log_mod_DS_photo_6d, newdata=ES_DS_fv_fm2, ci=TRUE)
ES_DS_fv_fm2_CI4 <- cbind(ES_DS_fv_fm2, pred_log_mod_DS_photo_6d)
#head(ES_DS_fv_fm2_CI4)
min_conc_d <- ES_DS_fv_fm2_CI4 %>% 
  mutate(overlap0 = 0 >= ci.lb & 0 <= ci.ub) %>% 
  filter(overlap0==FALSE) %>% 
  filter(ci.ub<0) %>% 
  summarize(min_conc_d=min(Sed_level_standardized))
min_conc_d2 <- as.numeric(min_conc_d)

ggplot(ES_DS_fv_fm2_CI4, 
       aes(x = Sed_level_standardized,
           y = smd,
           color = Ref,
           ymin = smd-vmd,
           ymax = smd+vmd)) + 
  geom_pointrange() +
  labs(x = expression("Sediment exposure concentration (mg/cm"^"2"*"/day)"),
       y = expression("Effect size (Hedges'"~italic(d)~"+/- variance of"~italic(d)~")"),
       color = "Study") +
  geom_abline(intercept=0, slope=0) +
  geom_vline(xintercept=min_conc_d2, linetype="dashed", color = "red") +
  geom_ribbon(aes(x = Sed_level_standardized, y = smd,
                  ymin = ci.lb, ymax = ci.ub), 
              fill = "grey70", alpha = .15, 
              show.legend=FALSE, inherit.aes=FALSE) +
  geom_line(aes(x = Sed_level_standardized, y = fit), inherit.aes=FALSE) +
  theme_classic() +
  scale_x_log10(limits = c(0.1,1000), breaks=c(0.1,1,10,100,1000,min_conc_d2), 
                label=c("0.1","1","10","100","1000",round(min_conc_d2,digits=1)))
```
  
  
## B. Suspended Sediment  
  
### 1. Plots  
I calculated the standardized mean difference in terms of Hedges' *d* and plot it here with respect to exposure concentration. I did this first for a Hedges' *d* that I calculated before importing the data to R and again for a Hedges' *d* that is calculated by the dosresmeta::covar.smd function.

```{r, echo=FALSE, warning=FALSE}
#EXPOSURE CONCENTRATION vs. effect size, by study
#Using previously calculated Hedges' d and variance of d
ggplot(ES_SS_fv_fm2, 
       aes(x = Sed_level_standardized,
           y = Hedges_d,
           color = Ref,
           ymin = Hedges_d-Var_d,
           ymax = Hedges_d+Var_d)) + 
  geom_pointrange() +
  ggtitle("Photosynthetic efficiency, Fv/Fm") +
  labs(x = "Sediment exposure concentration (mg/L)",
       y = "Effect size (Hedges' d +/- s.d., manual calculation)",
       color = "Study") +
  geom_abline(intercept=0, slope=0) +
  theme_classic() 
#Using dosresmeta's calculation of Hedges' d and variance of d
ggplot(ES_SS_fv_fm2, 
       aes(x = Sed_level_standardized,
           y = smd,
           color = Ref,
           ymin = smd-vmd,
           ymax = smd+vmd)) + 
  geom_pointrange() +
  ggtitle("Photosynthetic efficiency, Fv/Fm") +
  labs(x = "Sediment exposure concentration (mg/L)",
       y = "Effect size (Hedges' d +/- s.d., dosresmeta calculation)",
       color = "Study") +
  geom_abline(intercept=0, slope=0) +
  theme_classic() 


#by SEDIMENT TYPE
#Using previously calculated Hedges' d and variance of d
ggplot(ES_SS_fv_fm2, 
       aes(x = Sed_level_standardized,
           y = smd,
           color = Sed_grain_category,
           shape = Sed_source,
           ymin = smd-vmd,
           ymax = smd+vmd)) + 
  geom_pointrange() +
  ggtitle("Photosynthetic efficiency, Fv/Fm") +
  labs(x = "Sediment exposure concentration (mg/L)",
       y = "Effect size (Hedges' d +/- s.d., dosresmeta calculation)",
       color = "Sediment relative\ngrain size",
       shape = "Sediment source") +
  geom_abline(intercept=0, slope=0) +
  theme_classic()
ggplot(ES_SS_fv_fm2, 
       aes(x = Sed_level_standardized,
           y = smd,
           color = Sed_grain_class,
           shape = Sed_mineralogy,
           ymin = smd-vmd,
           ymax = smd+vmd)) + 
  geom_pointrange() +
  ggtitle("Photosynthetic efficiency, Fv/Fm") +
  labs(x = "Sediment exposure concentration (mg/L)",
       y = "Effect size (Hedges' d +/- s.d., dosresmeta calculation)",
       color = "Sediment\ngrain class",
       shape = "Sediment mineralogy") +
  geom_abline(intercept=0, slope=0) +
  theme_classic()

#by PHYLOGENY
#Using previously calculated Hedges' d and variance of d
ggplot(ES_SS_fv_fm2, 
       aes(x = Sed_level_standardized,
           y = smd,
           color = Updated_Genus,
           ymin = smd-vmd,
           ymax = smd+vmd)) + 
  geom_pointrange() +
  ggtitle("Photosynthetic efficiency, Fv/Fm") +
  labs(x = "Sediment exposure concentration (mg/L)",
       y = "Effect size (Hedges' d +/- s.d., dosresmeta calculation)",
       color = "Genus") +
  geom_abline(intercept=0, slope=0) +
  theme_classic()

#by OCEAN and REGION
#Using previously calculated Hedges' d and variance of d
ggplot(ES_SS_fv_fm2, 
       aes(x = Sed_level_standardized,
           y = smd,
           color = Region,
           shape = Ocean,
           ymin = smd-vmd,
           ymax = smd+vmd)) + 
  geom_pointrange() +
  ggtitle("Photosynthetic efficiency, Fv/Fm") +
  labs(x = "Sediment exposure concentration (mg/L)",
       y = "Effect size (Hedges' d +/- s.d., dosresmeta calculation)",
       color = "Region",
       shape = "Ocean") +
  geom_abline(intercept=0, slope=0) +
  theme_classic()
```
  
It is interesting that there is a difference between how I calculated Hedges' *d* and how *dosresmeta* appears to calculate it. But in both cases, there is a funnel-like relationship between sediment exposure and effect size. There does not appear to be strong effects of phylogeny (genus) or geography (ocean and region). **There may be effects of sediment type (source, origin, grain size), but these will likely be indiscernible, given the variance.**
  
### 2. Model Fitting  
The following set of functions model fixed and random (nested) effects WITHOUT accounting for within-comparison (or within-study) correlations. The output for each is a univariate, random-effects meta-regression.

```{r, echo=FALSE, warning=FALSE}
#?mixmeta

# Organizational Note: 
#All models have exposure concentration as its predictor. 
#All models with 'd' at the end of its name have covar.smd calculated Hedges' d as its response.

# LINEAR FIXED AND RANDOM EFFECTS NOT ACCOUNTING FOR WITHIN-STUDY CORRELATIONS
mod_SS_photo_1d <- mixmeta(smd ~ Sed_level_standardized, S = vmd,
                      random =  ~ Sed_level_standardized | Comparison,
                      data = ES_SS_fv_fm2, method = "ml")

# LINEAR FIXED AND RANDOM, NESTED EFFECTS NOT ACCOUNTING FOR WITHIN-STUDY CORRELATIONS
mod_SS_photo_2d <- mixmeta(smd ~ Sed_level_standardized, S = vmd,
                      random =  ~ Sed_level_standardized | Ref/Comparison,
                      data = ES_SS_fv_fm2, method = "ml")

# LINEAR FIXED AND RANDOM, NESTED EFFECTS NOT ACCOUNTING FOR WITHIN-STUDY CORRELATIONS
mod_SS_photo_3d <- mixmeta(smd ~ Sed_level_standardized, S = vmd,
                      random =  ~ Sed_level_standardized | Ref_name/Comparison,
                      data = ES_SS_fv_fm2, method = "ml")

# LINEAR FIXED AND RANDOM, NESTED EFFECTS NOT ACCOUNTING FOR WITHIN-STUDY CORRELATIONS
mod_SS_photo_4d <- mixmeta(smd ~ Sed_level_standardized, S = vmd,
                       random =  ~ Sed_level_standardized | Ref_name/Ref/Comparison,
                       data = ES_SS_fv_fm2, method = "ml")
```

To account for within-comparison correlations in the mixmeta model, I replace the 'S' argument with a 'control' argument, which requires me to provide the covariance matrices, as done with the covar.smd function. 

```{r, echo=FALSE, warning=FALSE}
# COMPUTE THE WITHIN-COMPARISON CORRELATIONS 
# already done above but code repeated here...
#covar_DS_photo <- by(ES_DS_fv_fm, ES_DS_fv_fm$Comparison, function(x) 
#  covar.smd(Tx_mean, Tx_sd, Tx_n, "smd", method="hedges", data = x))

# LINEAR FIXED AND RANDOM EFFECTS ACCOUNTING FOR WITHIN-COMPARISON CORRELATIONS
#Code for covariance matrix to include for addSlist for Comparison random effect
newlist_FVFM_SS <- list(NA)
for (i in seq(1,length(covar_SS_photo))) {
newlist_FVFM_SS[i] <- list(covar_SS_photo[[i]]$S)
}
mod_SS_photo_5d <- mixmeta(smd ~ Sed_level_standardized,
                       random = ~ Sed_level_standardized | Comparison,
                       data=ES_SS_fv_fm2, method="ml", #using dataset without controls
                       control=list(addSlist=newlist_FVFM_SS))


# LINEAR FIXED AND RANDOM, NESTED EFFECTS ACCOUNTING FOR WITHIN-COMPARISON & WITHIN-STUDY CORRELATIONS
#Code for covariance matrix to include for addSlist for Ref/Comparison
newlist_FVFM_SS2 <- list(NA) #collects the list of covariance matrices for the block diag matrix by reference for the nested hierarchial ref/comparison
templist_FVFM_SS <- list(NA) #holds temp list of covariance matrices associated with each reference
templist_FVFM_SS2 <-list(NA)
reflista <- ES_SS_fv_fm %>% distinct(Ref,Ref_name,Comparison)
reflist <- reflista[,1]
for (i in seq(1,length(unique(reflista$Ref))))  {
  #pull the elements from covar_SS_photo that are all from the same reference [i]
  templist_FVFM_SS[i] <-list(covar_SS_photo[reflist==unique(reflista$Ref)[i]])
  for (j in seq(1,length(templist_FVFM_SS[[i]]))) {
  #for each comparison in the reference, pull out the covar matrices (element $S) and put in into templist_FVFM_SS2
     templist_FVFM_SS2[j] <- list(templist_FVFM_SS[[i]][[j]]$S)
  }
  #turn list of covars from all comparison in one reference into block diag matrix
  newlist_FVFM_SS2[i] <- list(bdiagMat(templist_FVFM_SS2))
  templist_FVFM_SS2 <- list(NA)
}
mod_SS_photo_6d <- mixmeta(smd ~ Sed_level_standardized,
                       random = ~ Sed_level_standardized | Ref/Comparison,
                       data=ES_SS_fv_fm2, method="ml", #using dataset without controls
                       control=list(addSlist=newlist_FVFM_SS2))


# LINEAR FIXED AND RANDOM, NESTED EFFECTS ACCOUNTING FOR WITHIN-COMPARISON & WITHIN-STUDY CORRELATIONS
#Code for covariance matrix to include for addSlist for Ref_name/Comparison
newlist_FVFM_SS3 <- list(NA) #collects the list of covariance matrices for the block diag matrix by reference for the nested hierarchial Ref_name/Comparison
templist_FVFM_SS3 <- list(NA) #holds temp list of covariance matrices associated with each reference
templist_FVFM_SS4 <-list(NA)
reflista <- ES_SS_fv_fm %>% distinct(Ref,Ref_name,Comparison)
reflist <- reflista[,2]
for (i in seq(1,length(unique(reflista$Ref_name))))  {
  #pull the elements from covar_SS_photo that are all from the same reference [i]
  templist_FVFM_SS3[i] <-list(covar_SS_photo[reflist==unique(reflista$Ref_name)[i]])
  for (j in seq(1,length(templist_FVFM_SS3[[i]]))) {
  #for each comparison in the reference, pull out the covar matrices (element $S) and put in into templist_FVFM_SS4
     templist_FVFM_SS4[j] <- list(templist_FVFM_SS3[[i]][[j]]$S)
  }
  #turn list of covars from all comparison in one reference into block diag matrix
  newlist_FVFM_SS3[i] <- list(bdiagMat(templist_FVFM_SS4))
  templist_FVFM_SS4 <- list(NA)
}
mod_SS_photo_7d <- mixmeta(smd ~ Sed_level_standardized,
                       random = ~ Sed_level_standardized | Ref_name/Comparison,
                       data=ES_SS_fv_fm2, method="ml", #using dataset without controls
                       control=list(addSlist=newlist_FVFM_SS3))
```
  
And now to compare relative fit of linear models...

```{r, echo=FALSE, warning=FALSE}
# Model Comparisons
AIC(mod_SS_photo_1d, mod_SS_photo_2d, mod_SS_photo_3d, mod_SS_photo_4d, mod_SS_photo_5d, mod_SS_photo_6d, mod_SS_photo_7d) 
#of these, model 5d has lowest AIC
```
  
And now to evaluate residuals of models...

```{r, echo=FALSE, warning=FALSE}
# Residuals vs. Fitted Plots and Normal Q-Q Plots
op <- par(mfrow = c(1,2), mar = c(2,2,4,1))

resid_d <- resid(mod_SS_photo_5d)
fitted_d <- fitted(mod_SS_photo_5d)
plot(fitted_d, resid_d, main = "D: smd ~ mg_L, control=addSlist,
     random =  ~ mg_L | Comparison")
abline(0,0)

qqnorm(resid_d)
qqline(resid_d)

par(op)
```
  
Patterns:
1.  Not bad -- not high magnitude of residuals and does not look like it requires transformations or non-linear approaches.
  
  
### 4. Best-Fit Model Summary
  
#### MG/L IS PREDICTOR, Model using dosresmeta calculation of Hedges' *d*
Now I will predict the effect size along the exposure range and plot with confidence intervals.
```{r, echo=FALSE, warning=FALSE}
summary(mod_SS_photo_5d)

# PREDICT THE EFFECT SIZE FOR 13.2 mg/cm2 FROM TWO MODELS (LOAEL identified by binary analyses)
#predict(mod_SS_photo_5d, newdata=data.frame(Sed_level_standardized=13.2), ci=TRUE)

# PREDICT THE EFFECT SIZE AND PLOT WITH CONFIDENCE INTERVALS
pred_mod_SS_photo_5d <- predict(mod_SS_photo_5d, newdata=ES_SS_fv_fm2, ci=TRUE)
ES_SS_fv_fm2_CI4 <- cbind(ES_SS_fv_fm2, pred_mod_SS_photo_5d)
#head(ES_SS_fv_fm2_CI4)
min_conc_SSd <- ES_SS_fv_fm2_CI4 %>% 
  mutate(overlap0 = 0 >= ci.lb & 0 <= ci.ub) %>% 
  filter(overlap0==FALSE) %>% 
  summarize(min_conc_SSd=min(Sed_level_standardized))
min_conc_SSd2 <- as.numeric(min_conc_SSd)

ggplot(ES_SS_fv_fm2_CI4, 
       aes(x = Sed_level_standardized,
           y = smd,
           color = Ref,
           ymin = smd-vmd,
           ymax = smd+vmd)) + 
  geom_pointrange() +
  labs(x = "Sediment exposure concentration (mg/L)",
       y = expression("Effect size (Hedges'"~italic(d)~"+/- variance of"~italic(d)~")"),
       color = "Study") +
  geom_abline(intercept=0, slope=0) +
  geom_vline(xintercept=min_conc_SSd2, linetype="dashed", color = "red") +
  geom_ribbon(aes(x = Sed_level_standardized, y = smd,
                  ymin = ci.lb, ymax = ci.ub), 
              fill = "grey70", alpha = .15, 
              show.legend=FALSE, inherit.aes=FALSE) +
  geom_line(aes(x = Sed_level_standardized, y = fit), inherit.aes=FALSE) +
  theme_classic() +
  scale_x_log10(limits = c(1,250), breaks=c(1,10,100,min_conc_SSd2), 
                label=c("1","10","100",round(min_conc_SSd2,digits=1))) +
  scale_y_continuous(limits=c(-2.5,3))
```
  
  
# II. Photosynthesis-Respiration Ratio 

## A. Deposited Sediment  
  
### 1. Plots  
I calculated the standardized mean difference in terms of Hedges' *d* and plot it here with respect to exposure concentration. I did this first for a Hedges' *d* that I calculated before importing the data to R and again for a Hedges' *d* that is calculated by the dosresmeta::covar.smd function.

```{r, echo=FALSE, warning=FALSE}
#EXPOSURE CONCENTRATION vs. effect size, by study
#Using previously calculated Hedges' d and variance of d
ggplot(ES_DS_p_r2, 
       aes(x = Sed_level_standardized,
           y = Hedges_d,
           color = Ref,
           ymin = Hedges_d-Var_d,
           ymax = Hedges_d+Var_d)) + 
  geom_pointrange() +
  ggtitle("Photosynthesis-Respiration Ratio") +
  labs(x = "Sediment exposure concentration (mg/cm2/day)",
       y = "Effect size (Hedges' d +/- s.d., manual calculation)",
       color = "Study") +
  geom_abline(intercept=0, slope=0) +
  theme_classic() 
#Using dosresmeta's calculation of Hedges' d and variance of d
ggplot(ES_DS_p_r2, 
       aes(x = Sed_level_standardized,
           y = smd,
           color = Ref,
           ymin = smd-vmd,
           ymax = smd+vmd)) + 
  geom_pointrange() +
  ggtitle("Photosynthesis-Respiration Ratio") +
  labs(x = "Sediment exposure concentration (mg/cm2/day)",
       y = "Effect size (Hedges' d +/- s.d., dosresmeta calculation)",
       color = "Study") +
  geom_abline(intercept=0, slope=0) +
  theme_classic() 


#by SEDIMENT TYPE
#Using previously calculated Hedges' d and variance of d
ggplot(ES_DS_p_r2, 
       aes(x = Sed_level_standardized,
           y = smd,
           color = Sed_grain_category,
           shape = Sed_source,
           ymin = smd-vmd,
           ymax = smd+vmd)) + 
  geom_pointrange() +
  ggtitle("Photosynthesis-Respiration Ratio") +
  labs(x = "Sediment exposure concentration (mg/cm2/day)",
       y = "Effect size (Hedges' d +/- s.d., dosresmeta calculation)",
       color = "Sediment relative\ngrain size",
       shape = "Sediment source") +
  geom_abline(intercept=0, slope=0) +
  theme_classic() 
ggplot(ES_DS_p_r2, 
       aes(x = Sed_level_standardized,
           y = smd,
           color = Sed_grain_class,
           shape = Sed_mineralogy,
           ymin = smd-vmd,
           ymax = smd+vmd)) + 
  geom_pointrange() +
  ggtitle("Photosynthesis-Respiration Ratio") +
  labs(x = "Sediment exposure concentration (mg/cm2/day)",
       y = "Effect size (Hedges' d +/- s.d., dosresmeta calculation)",
       color = "Sediment\ngrain class",
       shape = "Sediment mineralogy") +
  geom_abline(intercept=0, slope=0) +
  theme_classic() 

#by PHYLOGENY
#Using previously calculated Hedges' d and variance of d
ggplot(ES_DS_p_r2, 
       aes(x = Sed_level_standardized,
           y = smd,
           color = Updated_Genus,
           ymin = smd-vmd,
           ymax = smd+vmd)) + 
  geom_pointrange() +
  ggtitle("Photosynthesis-Respiration Ratio") +
  labs(x = "Sediment exposure concentration (mg/cm2/day)",
       y = "Effect size (Hedges' d +/- s.d., dosresmeta calculation)",
       color = "Genus") +
  geom_abline(intercept=0, slope=0) +
  theme_classic() 

#by OCEAN and REGION
#Using previously calculated Hedges' d and variance of d
ggplot(ES_DS_p_r2, 
       aes(x = Sed_level_standardized,
           y = smd,
           color = Region,
           shape = Ocean,
           ymin = smd-vmd,
           ymax = smd+vmd)) + 
  geom_pointrange() +
  ggtitle("Photosynthesis-Respiration Ratio") +
  labs(x = "Sediment exposure concentration (mg/cm2/day)",
       y = "Effect size (Hedges' d +/- s.d., dosresmeta calculation)",
       color = "Region",
       shape = "Ocean") +
  geom_abline(intercept=0, slope=0) +
  theme_classic() 
```
  
It is interesting that there is a difference between how I calculated Hedges' *d* and how *dosresmeta* appears to calculate it. But in both cases, there may be a negative relationship between sediment exposure and effect size. There does not appear to be strong effects of phylogeny (genus) or geography (ocean and region). There may be effects of sediment type (source, origin, grain size), but I do not think these will be discernible given the variance.
   
### 2. Model Fitting
The following set of functions model fixed and random (nested) effects WITHOUT accounting for within-comparison (or within-study) correlations. The output for each is a univariate, random-effects meta-regression.
  
```{r, echo=FALSE, warning=FALSE}
#?mixmeta

# Organizational Note: 
#All models have exposure concentration as its predictor. 
#All models with 'd' at the end of its name have covar.smd calculated Hedges' d as its response.

# LINEAR FIXED AND RANDOM EFFECTS NOT ACCOUNTING FOR WITHIN-STUDY CORRELATIONS
mod_DS_p_r_1d <- mixmeta(smd ~ Sed_level_standardized, S = vmd,
                      random =  ~ Sed_level_standardized | Comparison,
                      data = ES_DS_p_r2, method = "ml")

# LINEAR FIXED AND RANDOM, NESTED EFFECTS NOT ACCOUNTING FOR WITHIN-STUDY CORRELATIONS
#Error in chol.default(XtVX) : the leading minor of order 6 is not positive definite
#mod_DS_p_r_2d <- mixmeta(smd ~ Sed_level_standardized, S = vmd,
#                      random =  ~ Sed_level_standardized | Ref/Comparison,
#                      data = ES_DS_p_r2, method = "ml")
```

To account for within-comparison correlations in the mixmeta model, I replace the 'S' argument with a 'control' argument, which requires me to provide the covariance matrices, as done with the covar.smd function. 

```{r, echo=FALSE, warning=FALSE}
# COMPUTE THE WITHIN-COMPARISON CORRELATIONS 
# already done above but code repeated here...
#covar_DS_photo <- by(ES_DS_p_r, ES_DS_p_r$Comparison, function(x) 
#  covar.smd(Tx_mean, Tx_sd, Tx_n, "smd", method="hedges", data = x))

# LINEAR FIXED AND RANDOM EFFECTS ACCOUNTING FOR WITHIN-COMPARISON CORRELATIONS
#Code for covariance matrix to include for addSlist for Comparison random effect
newlist_PR <- list(NA)
for (i in seq(1,length(covar_DS_p_r))) {
newlist_PR[i] <- list(covar_DS_p_r[[i]]$S)
}
mod_DS_p_r_5d <- mixmeta(smd ~ Sed_level_standardized,
                       random = ~ Sed_level_standardized | Comparison,
                       data=ES_DS_p_r2, method="ml", #using dataset without controls
                       control=list(addSlist=newlist_PR))

# LINEAR FIXED AND RANDOM, NESTED EFFECTS ACCOUNTING FOR WITHIN-COMPARISON & WITHIN-STUDY CORRELATIONS
#Code for covariance matrix to include for addSlist for Ref/Comparison
newlist_PR2 <- list(NA) #collects the list of covariance matrices for the block diag matrix by reference for the nested hierarchial ref/comparison
templist_PR <- list(NA) #holds temp list of covariance matrices associated with each reference
templist_PR2 <-list(NA)
reflista <- ES_DS_p_r %>% distinct(Ref,Ref_name,Comparison)
reflist <- reflista[,1]
for (i in seq(1,length(unique(reflista$Ref))))  {
  #pull the elements from covar_DS_p_r that are all from the same reference [i]
  templist_PR[i] <-list(covar_DS_p_r[reflist==unique(reflista$Ref)[i]])
  for (j in seq(1,length(templist_PR[[i]]))) {
  #for each comparison in the reference, pull out the covar matrices (element $S) and put in into templist_PR2
     templist_PR2[j] <- list(templist_PR[[i]][[j]]$S)
  }
  #turn list of covars from all comparison in one reference into block diag matrix
  newlist_PR2[i] <- list(bdiagMat(templist_PR2))
  templist_PR2 <- list(NA)
}
#Error in chol.default(XtVX) : the leading minor of order 6 is not positive definite
#mod_DS_p_r_6d <- mixmeta(smd ~ Sed_level_standardized,
#                       random = ~ Sed_level_standardized | Ref/Comparison,
#                       data=ES_DS_p_r2, method="ml", #using dataset without controls
#                       control=list(addSlist=newlist_PR2))
```
  
### 3. Model Comparison and Summary
And now to compare relative fit of linear models...

```{r, echo=FALSE, warning=FALSE}
# Model Comparisons
AIC(mod_DS_p_r_1d, mod_DS_p_r_5d) 
#of these, model 1d and 5d have lowest AIC
```
  
And now to evaluate residuals of models...

```{r, echo=FALSE, warning=FALSE}
# Residuals vs. Fitted Plots and Normal Q-Q Plots
op <- par(mfrow = c(2,2), mar = c(2,2,4,1))

resid_PRd <- resid(mod_DS_p_r_5d)
fitted_PRd <- fitted(mod_DS_p_r_5d)
plot(fitted_PRd, resid_PRd, main = "D: smd ~ mg_cm2_d, control=addSlist,
     random =  ~ mg_cm2_d | Comparison")
abline(0,0)

qqnorm(resid_PRd)
qqline(resid_PRd)

par(op)
```
  
Patterns:
1.  Sparse data, but they all seem to fit a linear trend with minimal residual magnitudes.
  
### 4. Best-Fit Model Summary  
  
#### MG/CM2/DAY IS PREDICTOR, Model using dosresmeta calculation of Hedges' *d*
Now I will predict the effect size along the exposure range and plot with confidence intervals.
```{r, echo=FALSE, warning=FALSE}
summary(mod_DS_p_r_5d)

# PREDICT THE EFFECT SIZE FOR 13.2 mg/cm2 FROM TWO MODELS (LOAEL identified by binary analyses)
#predict(mod_DS_p_r_5d, newdata=data.frame(Sed_level_standardized=13.2), ci=TRUE)

# PREDICT THE EFFECT SIZE AND PLOT WITH CONFIDENCE INTERVALS
pred_mod_DS_p_r_5d <- predict(mod_DS_p_r_5d, newdata=ES_DS_p_r2, ci=TRUE)
ES_DS_p_r2_CI4 <- cbind(ES_DS_p_r2, pred_mod_DS_p_r_5d)
min_conc_d <- ES_DS_p_r2_CI4 %>% 
  mutate(overlap0 = 0 >= ci.lb & 0 <= ci.ub) %>% 
  filter(overlap0==FALSE) %>% 
  summarize(min_conc_d=min(Sed_level_standardized))
min_conc_d2 <- as.numeric(min_conc_d)

ggplot(ES_DS_p_r2_CI4, 
       aes(x = Sed_level_standardized,
           y = Hedges_d,
           color = Ref,
           ymin = Hedges_d-Var_d,
           ymax = Hedges_d+Var_d)) + 
  geom_pointrange() +
  labs(x = expression("Sediment exposure concentration (mg/cm"^"2"*"/day)"),
       y = expression("Effect size (Hedges'"~italic(d)~"+/- variance of"~italic(d)~")"),
       color = "Study") +
  geom_abline(intercept=0, slope=0) +
  geom_vline(xintercept=min_conc_d2, linetype="dashed", color = "red") +
  geom_ribbon(aes(x = Sed_level_standardized, y = Hedges_d,
                  ymin = ci.lb, ymax = ci.ub), 
              fill = "grey70", alpha = .15, 
              show.legend=FALSE, inherit.aes=FALSE) +
  geom_line(aes(x = Sed_level_standardized, y = fit), inherit.aes=FALSE) +
  theme_classic() +
  scale_y_continuous(limits=c(-8,1)) +
  scale_x_continuous(limits=c(20,65), breaks=c(20,40,60,min_conc_d2), 
                label=c("20","40","60",round(min_conc_d2,digits=0)))
```
